---
title: "Municipal Indicators For Millinocket, ME"
output:
  pdf_document:
    toc: yes
    toc_depth: '3'
  html_notebook:
    toc: yes
    toc_depth: 3
  html_document:
    df_print: paged
    toc: yes
    toc_depth: '3'
---

In this report, 39 demographic, educational institutions, employment, healthcare, municipal finance, infrastructure, food insecurity and housing variables are presented using federally-published data sources for the city of Millinocket, Maine. To better demonstrate their ties to geography of the city, the variables have been mapped on the block group level. \


## Demographic
\
```{r echo=FALSE, warning=FALSE, message=FALSE, results='hide', comment=""}
                                 #######################################################
                                 #       Municipal Indicators For Municipalities        #               
                                 #######################################################
library(stringr)
library(tidycensus)
library(tidyverse)
library(tigris)
library(totalcensus)

options(tigris_class = "sf",tigris_use_cache = T)
# Get API key 
# census_api_key("045944708d514a404443c9f2d227ac8bf6a8f8bd", install = TRUE, overwrite=TRUE)
# Specify the end year 
# This is for 5-year estimates
endyr <- 2018
# List of ACS variables
Var_List <- tidycensus::load_variables(year = endyr, dataset = "acs5", cache = T)
#View(Var_List)
```


\


### Population  Size
\

```{r echo=FALSE, warning=FALSE, message=FALSE, results='hide', comment=""}

                                             #--------------------#
                                             #    DEMOGRAPHIC     #
                                             #--------------------#

##############################################    Population  Size    ###########################################################
library(tidyverse)
library(dplyr)
# search_cbsa()
# search_fips()
# set_path_to_census("C:\\Users\\ux305\\Desktop\\census")
j <- "ME"
area <- "Millinocket town, ME"
c <- "Penobscot"
ch.pop <- read_acs5year(year = endyr,
                     states = j,
                     table_contents = "B01003_001",
                     areas = area,
                     summary_level = "tract")
ch.pop <- ch.pop[ch.pop$NAME == "Census Tract 300, Penobscot County, Maine",] %>%
  rename(TotalPop = population) %>%
  rename(City = area) %>% 
  dplyr::select(City, TotalPop)


Chelsea.MA <- cbind(ch.pop)
Chelsea.MA

ch.MA.Pop <- read_acs5year(year = endyr,
                     states = j,
                     table_contents = "B01003_001",
                     areas = area,
                     summary_level = "block group")

ch.Pop.get <- tidycensus::get_acs(geography = "block group",
													 survey="acs5",
													 state=j,
													 county=c,
													 year = endyr,
													 output = "wide",
													 variables = "B01003_001", # Total Population
													 keep_geo_vars = T, geometry = TRUE)

Chelsea <- merge(ch.Pop.get, ch.MA.Pop, by.x = 'NAME.y', by.y='NAME')


```


The total population of `r Chelsea.MA$City` is `r Chelsea.MA$TotalPop`. Followed is the population size map of the city.\
\

```{r , echo=FALSE, warning=FALSE, message=FALSE, comment=""}

## MAP
library(tmap)
tm_shape(Chelsea[c(2:5),]) + tm_polygons('B01003_001E' , palette="seq", style = "quantile", title = "Total Population")  + tm_view(alpha = 0.85,view.legend.position = c("right", "bottom"),basemaps = "Stamen.TonerLite") +tm_borders(col = "darkgray")
```
\

* Hint: Millinocket city contains 5 block groups, but it seems one of these block groups is located in another tract (Census Tract 290) and if we consider all 5 block groups for maps, the shape of map would be different from the shape of the city. Therefore, for more clarification we consider 4 block groups, which are located in Census Tract 300.\

\

### Population Age
\

```{r echo=FALSE, warning=FALSE, message=FALSE, results='hide', comment=""}
##############################################    Population  Age   #############################################################
MA.Age.ch <- read_acs5year(year = endyr,
                     states = j,
                     table_contents = c("B01001_003","B01001_004","B01001_005","B01001_006",
													 							"B01001_027","B01001_028","B01001_029","B01001_030", # Under 5 years; Between 5 to 17
													 							"B01001_007","B01001_008","B01001_009",
													 							"B01001_010","B01001_011","B01001_012",
													 							"B01001_036","B01001_035","B01001_034",
													 							"B01001_033","B01001_032","B01001_031", # 18 to 24 years; 25 to 34 years
													 							"B01001_013","B01001_014","B01001_015",
													 							"B01001_016","B01001_017","B01001_018","B01001_019",
													 							"B01001_043","B01001_042","B01001_041","B01001_040",
													 							"B01001_039","B01001_038","B01001_037", # 35 to 44 years; 45 to 54 years; 55 to 59 years; 60 to 64 years
													 							"B01001_020","B01001_021","B01001_022",
													 							"B01001_023","B01001_024","B01001_025",
													 							"B01001_044","B01001_045","B01001_046",
													 							"B01001_047","B01001_048","B01001_049","B01003_001"),
                     areas = area,
                     summary_level = "tract") 
MA.Age.ch <- MA.Age.ch[MA.Age.ch$NAME == "Census Tract 300, Penobscot County, Maine",] %>%
  rename(City = area) %>%
				 mutate(AgeU18 = B01001_003+B01001_004+B01001_005+B01001_006+B01001_027+B01001_028+B01001_029+B01001_030, 
				 Age1834 = B01001_007+B01001_008+B01001_009+B01001_010+B01001_011+B01001_012+B01001_036+B01001_035+B01001_034+B01001_033+B01001_032+B01001_031,
				 Age3564 = B01001_013+B01001_014+B01001_015+B01001_016+B01001_017+B01001_018+B01001_019+B01001_043+B01001_042+B01001_041+B01001_040+B01001_039+B01001_038+B01001_037,
				 AgeO65 = B01001_020+B01001_021+B01001_022+B01001_023+B01001_024+B01001_025+B01001_044+B01001_045+B01001_046+B01001_047+B01001_048+B01001_049) %>%
  	mutate(AgeU18Per = AgeU18/population,
				 Age1834Per = Age1834/population,
				 Age3564Per = Age3564/population,
				 AgeO65Per = AgeO65/population) %>%
dplyr::select(AgeU18,Age1834,Age3564,AgeO65,AgeU18Per,Age1834Per,Age3564Per,AgeO65Per, City) %>%
  left_join(Chelsea.MA, MA.Age.ch, by="City")

Chelsea.MA <- cbind(MA.Age.ch)
```
* The number of people, who are younger than 18 Years old is `r Chelsea.MA$AgeU18` in `r Chelsea.MA$City` 
, which is `r (round(Chelsea.MA$AgeU18Per,4))*100`% of total population in `r Chelsea.MA$City`. \
* The number of people, who are between 18 and 34 years old is `r Chelsea.MA$Age1834` in `r Chelsea.MA$City`, which is `r (round(Chelsea.MA$Age1834Per,4))*100`% of total population in `r Chelsea.MA$City`.\
* The number of people, who are between 35 and 64 years old is `r Chelsea.MA$Age3564` in `r Chelsea.MA$City`, which is `r (round(Chelsea.MA$Age3564Per,4))*100`% of total population in `r Chelsea.MA$City`.\
* The number of people, who are older than 65 years old is `r Chelsea.MA$AgeO65` in `r Chelsea.MA$City`, which is `r (round(Chelsea.MA$AgeO65Per,4))*100`% of total population in `r Chelsea.MA$City`.\

Followed is the population age map of the city.\

```{r echo=FALSE, warning=FALSE, message=FALSE, comment="", results='hide'}
## Map

ch.MA.Age <- read_acs5year(year = endyr,
                     states = j,
                     table_contents = c("B01001_003","B01001_004","B01001_005","B01001_006",
													 							"B01001_027","B01001_028","B01001_029","B01001_030", # Under 5 years; Between 5 to 17
													 							"B01001_007","B01001_008","B01001_009",
													 							"B01001_010","B01001_011","B01001_012",
													 							"B01001_036","B01001_035","B01001_034",
													 							"B01001_033","B01001_032","B01001_031", # 18 to 24 years; 25 to 34 years
													 							"B01001_013","B01001_014","B01001_015",
													 							"B01001_016","B01001_017","B01001_018","B01001_019",
													 							"B01001_043","B01001_042","B01001_041","B01001_040",
													 							"B01001_039","B01001_038","B01001_037", # 35 to 44 years; 45 to 54 years; 55 to 59 years; 60 to 64 years
													 							"B01001_020","B01001_021","B01001_022",
													 							"B01001_023","B01001_024","B01001_025",
													 							"B01001_044","B01001_045","B01001_046",
													 							"B01001_047","B01001_048","B01001_049","B01003_001"),
                     areas = area,
                     summary_level = "block group")

ch.age.get <- tidycensus::get_acs(geography = "block group",
													 survey="acs5",
													 state=j,
													 county = c,
													 year = endyr,
													 output = "wide",
													 variables = c("B01001_003","B01001_004","B01001_005","B01001_006",
													 							"B01001_027","B01001_028","B01001_029","B01001_030", # Under 5 years; Between 5 to 17
													 							"B01001_007","B01001_008","B01001_009",
													 							"B01001_010","B01001_011","B01001_012",
													 							"B01001_036","B01001_035","B01001_034",
													 							"B01001_033","B01001_032","B01001_031", # 18 to 24 years; 25 to 34 years
													 							"B01001_013","B01001_014","B01001_015",
													 							"B01001_016","B01001_017","B01001_018","B01001_019",
													 							"B01001_043","B01001_042","B01001_041","B01001_040",
													 							"B01001_039","B01001_038","B01001_037", # 35 to 44 years; 45 to 54 years; 55 to 59 years; 60 to 64 years
													 							"B01001_020","B01001_021","B01001_022",
													 							"B01001_023","B01001_024","B01001_025",
													 							"B01001_044","B01001_045","B01001_046",
													 							"B01001_047","B01001_048","B01001_049","B01003_001"), # Total Population
													 keep_geo_vars = T, geometry = TRUE) %>%
mutate(AgeU18 = B01001_003E+B01001_004E+B01001_005E+B01001_006E+B01001_027E+B01001_028E+B01001_029E+B01001_030E, 
				 Age1834 = B01001_007E+B01001_008E+B01001_009E+B01001_010E+B01001_011E+B01001_012E+B01001_036E+B01001_035E+B01001_034E+B01001_033E+B01001_032E+B01001_031E,
				 Age3564 = B01001_013E+B01001_014E+B01001_015E+B01001_016E+B01001_017E+B01001_018E+B01001_019E+B01001_043E+B01001_042E+B01001_041E+B01001_040E+B01001_039E+B01001_038E+B01001_037E,
				 AgeO65 = B01001_020E+B01001_021E+B01001_022E+B01001_023E+B01001_024E+B01001_025E+B01001_044E+B01001_045E+B01001_046E+B01001_047E+B01001_048E+B01001_049E)  %>%
  rename(AgeU18_BG = AgeU18,
				 Age1834_BG = Age1834,
				 Age3564_BG = Age3564,
				 AgeO65_BG = AgeO65) %>%
	mutate(AgeU18Per_BG = ((AgeU18_BG/B01003_001E)*100),
				 Age1834Per_BG = ((Age1834_BG/B01003_001E)*100),
				 Age3564Per_BG = ((Age3564_BG/B01003_001E)*100),
				 AgeO65Per_BG = ((AgeO65_BG/B01003_001E)*100))  
Chelsea.age <- merge(ch.age.get, ch.MA.Age, by.x = 'NAME.y', by.y='NAME')

make_percent <- function(x, digits = 2) {
  paste0("%", formatC(x, digits = digits, format = "f", big.mark = ","))
}
 

tm_shape(Chelsea.age[c(2:5),]) + tm_polygons(c('AgeU18Per_BG','Age1834Per_BG',"Age3564Per_BG","AgeO65Per_BG") , palette=list("Oranges", "Greens", "Blues", "Purples") ,style = "quantile", title = c('AgeU18','Age1834',"Age3564","AgeO65"), popup.format = list(fun = make_percent), legend.format = list(fun = make_percent))  + tm_view(alpha = 0.85,basemaps = "Stamen.TonerLite") +tm_borders(col = "darkgray") + tmap_options(show.messages = FALSE) +tm_layout(legend.position = c("left","bottom"))

```
\

### Median Household Income
\

```{r  echo=FALSE, warning=FALSE, message=FALSE, results='hide', comment=""}
#################################################    Median Household Income    ##############################################################

MA.MHI.ch <- read_acs5year(year = endyr,
                     states = j,
                     table_contents = "B19013_001",
                     areas = area,
                     summary_level = "tract") 

MA.MHI.ch <- MA.MHI.ch[MA.MHI.ch$NAME == "Census Tract 300, Penobscot County, Maine",] %>%
				 rename(MedHouseIncome = B19013_001,
				        City = area) %>% 
	dplyr::select(City, MedHouseIncome) %>%
  left_join(Chelsea.MA, MA.MHI.ch, by="City")
Chelsea.MA <- cbind(MA.MHI.ch)
```
* The median household income in the past 12 months is 35697 Dollar in `r Chelsea.MA$City`.\

Followed is the Median Household Income map of the city.\

```{r echo=FALSE, warning=FALSE, message=FALSE, comment="", results='hide'}
## MAP
ch.MA.income <- read_acs5year(year = endyr,
                     states = j,
                     table_contents = "B19013_001",
                     areas = area,
                     summary_level = "block group")

ch.income.get <- tidycensus::get_acs(geography = "block group",
													 survey="acs5",
													 state=j,
													 county = c,
													 year = endyr,
													 output = "wide",
													 variables = "B19013_001", 
													 keep_geo_vars = T, geometry = TRUE) 

chelsea.income <- merge(ch.income.get, ch.MA.income, by.x = 'NAME.y', by.y='NAME')

make_dollar <- function(x, digits = 0) {
  paste0("$", formatC(x, digits = digits, format = "f", big.mark = ","))
}

library(tmap)
tm_shape(chelsea.income[c(2:5),]) + tm_polygons('B19013_001E' , palette="Blues", style="quantile", title = "Median Household Income", popup.format = list(fun = make_dollar), legend.format = list(fun = make_dollar))  + tm_view(alpha = 0.85,view.legend.position = c("right", "bottom"),basemaps = "Stamen.TonerLite") +tm_borders(col = "darkgray") + tmap_options(show.messages = FALSE) 
```

\

### Racial Breakdown 
\

```{r echo=FALSE, warning=FALSE, message=FALSE, results='hide', comment=""}
##############################################      Racial Breakdown     ###################################################################
MA.Race.ch <- read_acs5year(year = endyr,
                     states = j,
                     table_contents = c("B03002_003", # white alone
																					"B03002_004", # black or African American alone
																					"B03002_006", # Asian alone
																					"B03002_012", # Hispanic or Latino alone
																					"B03002_009"),
                     areas = area,
                     summary_level = "tract") 
MA.Race.ch <- MA.Race.ch[MA.Race.ch$NAME == "Census Tract 300, Penobscot County, Maine",] %>%
				 rename(White = B03002_003,
				 Black = B03002_004,
				 Asian = B03002_006,
				 Hispanic = B03002_012,
				 TwoOrMore = B03002_009,
				 City = area) %>% 
  mutate(WhitePer = (White/population),
         BlackPer = (Black/population),
         AsianPer = (Asian/population),
         HispanicPer = (Hispanic/population),
         TwoOrMorePer = (TwoOrMore/population)) %>%
	dplyr::select(City, White, Black, Asian, Hispanic, TwoOrMore,
	       WhitePer, BlackPer, AsianPer, HispanicPer, TwoOrMorePer) %>%
  left_join(Chelsea.MA, MA.Race.ch, by="City")
Chelsea.MA <- cbind(MA.Race.ch)
```
* The number of people who are white in `r Chelsea.MA$City` is `r Chelsea.MA$White` which is `r (round(Chelsea.MA$WhitePer,4))*100`% of total population in `r Chelsea.MA$City`.\
* The number of people who are black in `r Chelsea.MA$City` is `r Chelsea.MA$Black` which is `r (round(Chelsea.MA$BlackPer,4))*100`% of total population in `r Chelsea.MA$City`.\
* The number of people who are asian in `r Chelsea.MA$City` is `r Chelsea.MA$Asian` which is `r (round(Chelsea.MA$AsianPer,4))*100`% of total population in `r Chelsea.MA$City`.\
* The number of people who are hispanic in `r Chelsea.MA$City` is `r Chelsea.MA$Hispanic` which is `r (round(Chelsea.MA$HispanicPer,4))*100`% of total population in `r Chelsea.MA$City`.\
* The number of people who are have two or more races in `r Chelsea.MA$City` is `r Chelsea.MA$TwoOrMore` which is `r (round(Chelsea.MA$TwoOrMorePer,4))*100`% of total population in `r Chelsea.MA$City`. \

Followed is the Racial Breakdown map of the city.\

```{r echo=FALSE, warning=FALSE, message=FALSE, comment="",results='hide'}
## Map
ch.MA.race <- read_acs5year(year = endyr,
                     states = j,
                     table_contents = c("B03002_003", # white alone
																					"B03002_004", # black or African American alone
																					"B03002_006", # Asian alone
																					"B03002_012", # Hispanic or Latino alone
																					"B03002_009",
																					"B01003_001"),
                     areas = area,
                     summary_level = "block group")

ch.race.get <- tidycensus::get_acs(geography = "block group",
													 survey="acs5",
													 state=j,
													 county = c,
													 year = endyr,
													 output = "wide",
													 variables = c("B03002_003", # white alone
																					"B03002_004", # black or African American alone
																					"B03002_006", # Asian alone
																					"B03002_012", # Hispanic or Latino alone
																					"B03002_009",
																					"B01003_001"), 
													 keep_geo_vars = T, geometry = TRUE) %>%
rename(White_BG = B03002_003E,
				 Black_BG = B03002_004E,
				 Asian_BG = B03002_006E,
				 Hispanic_BG = B03002_012E,
				 TwoOrMore_BG = B03002_009E) %>%
  mutate(WhitePer_BG = (White_BG/B01003_001E)*100,
         BlackPer_BG = (Black_BG/B01003_001E)*100,
         AsianPer_BG = (Asian_BG/B01003_001E)*100,
         HispanicPer_BG = (Hispanic_BG/B01003_001E)*100,
         TwoOrMorePer_BG = (TwoOrMore_BG/B01003_001E)*100) 

chelsea.race <- merge(ch.race.get, ch.MA.race, by.x = 'NAME.y', by.y='NAME')


make_percent <- function(x, digits = 2) {
  paste0("%", formatC(x, digits = digits, format = "f", big.mark = ","))
}
 
library(tmap)
tm_shape(chelsea.race[c(2:5),]) + tm_polygons(c("WhitePer_BG", "BlackPer_BG", "AsianPer_BG", "HispanicPer_BG", "TwoOrMorePer_BG") , palette=list("Oranges", "Greens", "Blues", "Purples", "Reds"), style = "quantile", title = c("White", "Black", "Asian","Hispanic", "TwoOrMore"), popup.format = list(fun = make_percent), legend.format = list(fun = make_percent))  + tm_view(alpha = 0.85,view,basemaps = "Stamen.TonerLite") +tm_borders(col = "darkgray") + tmap_options(show.messages = FALSE) + tm_layout(legend.position = c("left","bottom"))
```

\


### Gender Breakdown
\

```{r echo=FALSE, warning=FALSE, message=FALSE, results='hide', comment=""}
#####################################################    Gender Breakdown    #############################################################

MA.Sex.ch <- read_acs5year(year = endyr,
                     states = j,
                     table_contents = c("B01001_026",
													               "B01001_002"),
                     areas = area,
                     summary_level = "tract")
MA.Sex.ch <- MA.Sex.ch[MA.Sex.ch$NAME == "Census Tract 300, Penobscot County, Maine",] %>%
  mutate(SexRatio = B01001_026/B01001_002) %>%
  rename(Female = B01001_026,
				 Male = B01001_002,
				 City = area) %>%
  mutate(FemalePer = Female/population,
				 MalePer = Male/population) %>% 
	dplyr::select(City, SexRatio, Female, Male, FemalePer, MalePer) %>%
  left_join(Chelsea.MA, MA.Sex.ch, by="City")

Chelsea.MA <- cbind(MA.Sex.ch)
```
* The number of females in `r Chelsea.MA$City` is `r Chelsea.MA$Female` which is `r 100*(round(Chelsea.MA$FemalePer,2))`% of total population in `r Chelsea.MA$City`.\
* The number of males in `r Chelsea.MA$City` is `r Chelsea.MA$Male` which is `r 100*(round(Chelsea.MA$MalePer,2))`% of total population in `r Chelsea.MA$City`.\
* The ratio of number of females to males in `r Chelsea.MA$City` is `r round(Chelsea.MA$SexRatio,2)`.\

Followed is the Gender Breakdown map of the city.\

```{r echo=FALSE, warning=FALSE, message=FALSE, comment="",results='hide'}
## MAP
ch.MA.sex <- read_acs5year(year = endyr,
                     states = j,
                     table_contents = c("B01001_026",
													               "B01001_002",
													               "B01003_001"),
                     areas = area,
                     summary_level = "block group")

ch.sex.get <- tidycensus::get_acs(geography = "block group",
													 survey="acs5",
													 state=j,
													 county = c,
													 year = endyr,
													 output = "wide",
													 variables = c("B01001_026",
													               "B01001_002",
													               "B01003_001"), 
													 keep_geo_vars = T, geometry = TRUE)  %>%
mutate(SexRatio = B01001_026E/B01001_002E) %>%
  rename(Female = B01001_026E,
				 Male = B01001_002E) %>%
  mutate(FemalePer = (Female/B01003_001E)*100,
         MalePer = (Male/B01003_001E)*100)

chelsea.sex <- merge(ch.sex.get, ch.MA.sex, by.x = 'NAME.y', by.y='NAME')

library(tmap)
tm_shape(chelsea.sex[c(2:5),]) + tm_polygons(c('Female', 'Male'), palette=list("Purples", "Reds"), title = c("Female","Male"))  + tm_view(alpha = 0.85,basemaps = "Stamen.TonerLite") +tm_borders(col = "darkgray")+ tmap_options(show.messages = FALSE) + tm_layout(legend.position = c("left","bottom"))

```

\



```{r echo=FALSE, warning=FALSE, message=FALSE, results='hide', comment=""}

tm_shape(chelsea.sex[c(2:5),]) + tm_polygons(c('FemalePer', 'MalePer'), palette=list("Blues", "Oranges"), title = c("Female Percentage","Male Percentage"), popup.format = list(fun = make_percent), legend.format = list(fun = make_percent)) + tm_view(alpha = 0.85,basemaps = "Stamen.TonerLite") +tm_borders(col = "darkgray")+ tmap_options(show.messages = FALSE) + tm_layout(legend.position = c("left","bottom"))

```




\

### Average Household Size
\
```{r echo=FALSE, warning=FALSE, message=FALSE, results='hide', comment=""}
#################################################    Average Household Size    ############################################################

MA.AveHH.Size.ch <- read_acs5year(year = endyr,
                     states = j,
                     table_contents = c("B11002_001", 
															         "B19001_001"),
                     areas = area,
                     summary_level = "tract") 
MA.AveHH.Size.ch <- MA.AveHH.Size.ch[MA.AveHH.Size.ch$NAME == "Census Tract 300, Penobscot County, Maine",] %>%
  rename(City = area) %>% 
  mutate(AveHouseH = B11002_001/B19001_001) %>%
	dplyr::select(City, AveHouseH) %>%
  left_join(Chelsea.MA, MA.AveHH.Size.ch, by="City")

Chelsea.MA <- cbind(MA.AveHH.Size.ch)
```
* The average household size is `r round(Chelsea.MA$AveHouseH, 2)` in `r Chelsea.MA$City`.\

Followed is the Average Household Size map of the city.\

```{r echo=FALSE, warning=FALSE, message=FALSE, comment="",results='hide'}
## MAP

ch.MA.HH.Size <- read_acs5year(year = endyr,
                     states = j,
                     table_contents = c("B11002_001", # Household type by relatives and nonrelatives for population in households
															              "B19001_001"),
                     areas = area,
                     summary_level = "block group")

ch.HH.Size.get <- tidycensus::get_acs(geography = "block group",
													 survey="acs5",
													 state=j,
													 county = c,
													 year = endyr,
													 output = "wide",
													 variables = c("B11002_001", # Household type by relatives and nonrelatives for population in households
															              "B19001_001"), 
													 keep_geo_vars = T, geometry = TRUE)  %>%
  mutate(AveHouseH = B11002_001E/B19001_001E) 

chelsea.HH.Size <- merge(ch.HH.Size.get, ch.MA.HH.Size, by.x = 'NAME.y', by.y='NAME')


tm_shape(chelsea.HH.Size[c(2:5),]) + tm_polygons("AveHouseH", style = "quantile", palette="Purples", border.col = "darkgray", title="Ave Household Size") + tm_view(alpha = 0.85,view.legend.position = c("right", "bottom"), basemaps = "Stamen.TonerLite") + tmap_options(show.messages = FALSE)

```

\

### Educational Attainment
\
```{r echo=FALSE, warning=FALSE, message=FALSE, results='hide', comment=""}
#############################################   Educational Attainment    ##################################################################
MA.Educ.ch <- read_acs5year(year = endyr,
                     states = j,
                     table_contents = c("B15002_001", # variables for Sex by Edu Attainment for Population 25 yrs or Over
																					"B15002_003","B15002_004","B15002_005",
																					"B15002_006","B15002_007","B15002_008","B15002_009","B15002_010",
																					"B15002_020","B15002_021","B15002_022","B15002_023","B15002_024",
																					"B15002_025","B15002_026","B15002_027",
																					"B15002_011","B15002_028",
																					"B15002_012","B15002_013","B15002_014",
																					"B15002_029","B15002_030","B15002_031",
																					"B15002_015","B15002_032",
																					"B15002_016","B15002_033",
																					"B15002_017","B15002_034",
																					"B15002_018","B15002_035"),
                     areas = area,
                     summary_level = "tract") 
MA.Educ.ch <- MA.Educ.ch[MA.Educ.ch$NAME == "Census Tract 300, Penobscot County, Maine",] %>%
  rename(City = area) %>% 
  mutate(LessThanHS = (B15002_003+B15002_004+B15002_005+B15002_006+B15002_007+B15002_008+B15002_009+B15002_010+B15002_020+B15002_021+B15002_022+B15002_023+B15002_024+B15002_025+B15002_026+B15002_027),
				 HSGrad = (B15002_011 + B15002_028),
				 SomeColl = (B15002_012+ B15002_013 + B15002_014 + B15002_029 + B15002_030 + B15002_031),
				 Bach = (B15002_015 + B15002_032),
				 Master = (B15002_016 + B15002_033),
				 Prof = (B15002_017 + B15002_034),
				 Doc = (B15002_018 + B15002_035),
				 LessThanHSPer = (LessThanHS/population),
         HSGradPer = (HSGrad/population),
         SomeCollPer = (SomeColl/population),
         BachPer = (Bach/population),
         MasterPer = (Master/population),
         ProfPer = (Prof/population),            
         DocPer = (Doc/population)) %>%
	dplyr::select(City, LessThanHS, HSGrad, SomeColl, Bach, Master, Prof, Doc,
	       LessThanHSPer, HSGradPer, SomeCollPer, BachPer, MasterPer, ProfPer, DocPer) %>% 
  left_join(Chelsea.MA, MA.Educ.ch, by="City") 

Chelsea.MA <- cbind(MA.Educ.ch)
```
* The number of people, who have less than high school degree is `r  Chelsea.MA$LessThanHS` in `r Chelsea.MA$City` which is `r (round(Chelsea.MA$LessThanHSPer,4))*100`% of the population who are 25 years old or older in `r Chelsea.MA$City`.\
* The number of people, who have high school degree is `r  Chelsea.MA$HSGrad` in `r Chelsea.MA$City` which is `r (round(Chelsea.MA$HSGradPer,4))*100`% of the population who are 25 years old or older in `r Chelsea.MA$City`.\
* The number of people, who have some college degree is `r  Chelsea.MA$SomeColl` in `r Chelsea.MA$City` which is `r (round(Chelsea.MA$SomeCollPer,4))*100`% of the population who are 25 years old or older in `r Chelsea.MA$City`.\
* The number of people, who have bachelor degree is `r Chelsea.MA$Bach` in `r Chelsea.MA$City` which is `r (round(Chelsea.MA$BachPer,4))*100`% of the population who are 25 years old or older in `r Chelsea.MA$City`.\
* The number of people, who have master's degree is `r Chelsea.MA$Master` in `r Chelsea.MA$City` which is `r (round(Chelsea.MA$MasterPer,4))*100`% of the population who are 25 years old or older in `r Chelsea.MA$City`.\
* The number of people, who have professional school degree is `r  Chelsea.MA$Prof` in `r Chelsea.MA$City` which is `r (round(Chelsea.MA$ProfPer,4))*100`% of the population who are 25 years old or older in `r Chelsea.MA$City`.\
* The number of people, who have doctorate degree is `r  Chelsea.MA$Doc` in `r Chelsea.MA$City` which is `r (round(Chelsea.MA$DocPer,4))*100`% of the population who are 25 years old or older in `r Chelsea.MA$City`.\

Followed is the Educational Attainment map of the city.\

```{r echo=FALSE, warning=FALSE, message=FALSE, comment="",results='hide'}
## MAP
ch.MA.Educ <- read_acs5year(year = endyr,
                     states = j,
                     table_contents = c("B15002_001", # variables for Sex by Edu Attainment for Population 25 yrs or Over
																					"B15002_003","B15002_004","B15002_005",
																					"B15002_006","B15002_007","B15002_008","B15002_009","B15002_010",
																					"B15002_020","B15002_021","B15002_022","B15002_023","B15002_024",
																					"B15002_025","B15002_026","B15002_027",
																					"B15002_011","B15002_028",
																					"B15002_012","B15002_013","B15002_014",
																					"B15002_029","B15002_030","B15002_031",
																					"B15002_015","B15002_032",
																					"B15002_016","B15002_033",
																					"B15002_017","B15002_034",
																					"B15002_018","B15002_035"),
                     areas = area,
                     summary_level = "block group")

ch.Educ.get <- tidycensus::get_acs(geography = "block group",
													 survey="acs5",
													 state=j,
													 county = c,
													 year = endyr,
													 output = "wide",
													 variables = c("B15002_001", # variables for Sex by Edu Attainment for Population 25 yrs or Over
																					"B15002_003","B15002_004","B15002_005",
																					"B15002_006","B15002_007","B15002_008","B15002_009","B15002_010",
																					"B15002_020","B15002_021","B15002_022","B15002_023","B15002_024",
																					"B15002_025","B15002_026","B15002_027",
																					"B15002_011","B15002_028",
																					"B15002_012","B15002_013","B15002_014",
																					"B15002_029","B15002_030","B15002_031",
																					"B15002_015","B15002_032",
																					"B15002_016","B15002_033",
																					"B15002_017","B15002_034",
																					"B15002_018","B15002_035"), 
													 keep_geo_vars = T, geometry = TRUE) %>%
  mutate(LessThanHS_BG = (B15002_003E+B15002_004E+B15002_005E+B15002_006E+B15002_007E+B15002_008E+B15002_009E+B15002_010E+B15002_020E+B15002_021E+B15002_022E+B15002_023E+B15002_024E+B15002_025E+B15002_026E+B15002_027E),
				 HSGrad_BG = (B15002_011E + B15002_028E),
				 SomeColl_BG = (B15002_012E + B15002_013E + B15002_014E + B15002_029E + B15002_030E + B15002_031E),
				 Bach_BG = (B15002_015E + B15002_032E),
				 Master_BG = (B15002_016E + B15002_033E),
				 Prof_BG = (B15002_017E + B15002_034E),
				 Doc_BG = (B15002_018E + B15002_035E),
		    LessThanHSPer_BG = ((B15002_003E+B15002_004E+B15002_005E+B15002_006E+B15002_007E+B15002_008E+B15002_009E+B15002_010E+B15002_020E+B15002_021E+B15002_022E+B15002_023E+B15002_024E+B15002_025E+B15002_026E+B15002_027E)/B15002_001E)*100,
				 HSGradPer_BG = ((B15002_011E + B15002_028E)/B15002_001E)*100,
				 SomeCollPer_BG = ((B15002_012E + B15002_013E + B15002_014E + B15002_029E + B15002_030E + B15002_031E)/B15002_001E)*100,
				 BachPer_BG = ((B15002_015E + B15002_032E)/B15002_001E)*100,
				 MasterPer_BG = ((B15002_016E + B15002_033E)/B15002_001E)*100,
				 ProfPer_BG = ((B15002_017E + B15002_034E)/B15002_001E)*100,
				 DocPer_BG = ((B15002_018E + B15002_035E)/B15002_001E)*100)

chelsea.Educ <- merge(ch.Educ.get, ch.MA.Educ, by.x = 'NAME.y', by.y='NAME')

make_percent <- function(x, digits = 2) {
  paste0("%", formatC(x, digits = digits, format = "f", big.mark = ","))
}

tm_shape(chelsea.Educ[c(2:5),]) +
    tm_fill(c('LessThanHSPer_BG', 'HSGradPer_BG', 'SomeCollPer_BG', 'BachPer_BG', 'MasterPer_BG', 'ProfPer_BG', 'DocPer_BG'),
        style="quantile", 
        palette=list("Oranges", "Greens", "Blues", "Purples", "Reds","Dark2", "seq"), popup.format = list(fun = make_percent), legend.format = list(fun = make_percent), title=c('LessThanHS', 'HSGrad', 'SomeColl', 'Bach', 'Master', 'Prof', 'Doc')) + tm_view(alpha = 0.85,basemaps = "Stamen.TonerLite") + tm_borders(col = "darkgray") +tmap_options(show.messages = FALSE) + tm_layout(legend.position = c("left","bottom"))

```

\

### Single Parent Households
\
```{r echo=FALSE, warning=FALSE, message=FALSE, results='hide', comment=""}

##########################################      Single Parent Households     ##################################################################
MA.Single.ch <- read_acs5year(year = endyr,
                     states = j,
                     table_contents = c("B11001_006", # Female Hoseholder, No Husband Present
																						"B11001_005", # Male Householder, No Wife Present
																						"B11001_001"),
                     areas = area,
                     summary_level = "tract") 
MA.Single.ch <- MA.Single.ch[MA.Single.ch$NAME == "Census Tract 300, Penobscot County, Maine",] %>%
  rename(City = area) %>% 
  mutate(FemHeadPer = B11001_006/B11001_001,
	       MaleHeadPer = B11001_005/B11001_001) %>%
	dplyr::select(City, FemHeadPer, MaleHeadPer) %>%
  left_join(Chelsea.MA, MA.Single.ch, by="City")

Chelsea.MA <- cbind(MA.Single.ch)

```
* The percentage of single-female-headed is `r (round(Chelsea.MA$FemHeadPer,4))*100`% of total number of family households in `r Chelsea.MA$City`.\
* The percentage of single-male-headed is `r (round(Chelsea.MA$MaleHeadPer,4))*100`% of total number of family households in `r Chelsea.MA$City`.\

Followed is the Single Parent Households map of the city.\

```{r echo=FALSE, warning=FALSE, message=FALSE, comment="",results='hide'}
## MAP
ch.MA.single <- read_acs5year(year = endyr,
                     states = j,
                     table_contents = c("B11001_006", # Female Hoseholder, No Husband Present
																						"B11001_005", # Male Householder, No Wife Present
																						"B11001_001"),
                     areas = area,
                     summary_level = "block group")

ch.single.get <- tidycensus::get_acs(geography = "block group",
													 survey="acs5",
													 state=j,
													 county = c,
													 year = endyr,
													 output = "wide",
													 variables = c("B11001_006", # Female Hoseholder, No Husband Present
																						"B11001_005", # Male Householder, No Wife Present
																						"B11001_001"), 
													 keep_geo_vars = T, geometry = TRUE)  %>%
mutate(FemHeadPer = (B11001_006E/B11001_001E)*100,
	       MaleHeadPer = (B11001_005E/B11001_001E)*100)

chelsea.single <- merge(ch.single.get, ch.MA.single, by.x = 'NAME.y', by.y='NAME')


make_percent <- function(x, digits = 2) {
  paste0("%", formatC(x, digits = digits, format = "f", big.mark = ","))
}


tm_shape(chelsea.single[c(2:5),]) + tm_polygons(c("FemHeadPer", "MaleHeadPer"), style="quantile", palette=list("Oranges", "Purples"), popup.format = list(fun = make_percent), legend.format = list(fun = make_percent), border.col = "darkgray", title=c("Female Householder", "Male Householder")) + tm_view(alpha = 0.85, basemaps = "Stamen.TonerLite") + tmap_options(show.messages = FALSE) + tm_layout(legend.position = c("left","bottom"))

```


\


### Population in Poverty
\
```{r echo=FALSE, warning=FALSE, message=FALSE, results='hide', comment=""}
###########################################      Population in Poverty    ######################################################################

MA.PopPov.ch <- read_acs5year(year = endyr,
                     states = j,
                     table_contents = c("B17010_002", # Income in the Past 12 Months Below Poverty Level (Family households in poverty)
																					"B17010_001",  # Total Families Below Poverty Level
																					"B17001_002", # Persons below poverty
																					"B17001_001"),
                     areas = area,
                     summary_level = "tract")
MA.PopPov.ch <- MA.PopPov.ch[MA.PopPov.ch$NAME == "Census Tract 300, Penobscot County, Maine",] %>%
  rename(City = area) %>% 
  	mutate(FamPovPer = B17010_002/B17010_001,
	       PopPovPer = B17001_002/B17001_001,
	       TotalPopPov = B17001_002) %>%
	dplyr::select(City, FamPovPer, PopPovPer, TotalPopPov) %>%
  left_join(Chelsea.MA,MA.PopPov.ch, by="City")

Chelsea.MA <- cbind(MA.PopPov.ch)
```
* The percentage of family, who are in poverty is `r (round(Chelsea.MA$FamPovPer,4))*100` % in `r Chelsea.MA$City`.\
* The percentage of population, who are in poverty is `r (round(Chelsea.MA$PopPovPer,4))*100` % in `r Chelsea.MA$City`.\
* The total number of people, who are in poverty is `r Chelsea.MA$TotalPopPov` in `r Chelsea.MA$City`.\

Followed is the Population in Poverty map of the city.\

```{r echo=FALSE, warning=FALSE, message=FALSE, comment="",results='hide'}
## MAP

ch.MA.pov <- read_acs5year(year = endyr,
                     states = j,
                     table_contents = c("B17010_002", # Income in the Past 12 Months Below Poverty Level (Family households in poverty)
																					"B17010_001",  # Total Families Below Poverty Level
																					"B17001_002", # Persons below poverty
																					"B17001_001"),
                     areas = area,
                     summary_level = "block group")

ch.pov.get <- tidycensus::get_acs(geography = "block group",
													 survey="acs5",
													 state=j,
													 county = c,
													 year = endyr,
													 output = "wide",
													 variables = c("B17010_002", # Income in the Past 12 Months Below Poverty Level (Family households in poverty)
																					"B17010_001",  # Total Families Below Poverty Level
																					"B17001_002", # Persons below poverty
																					"B17001_001"), 
													 keep_geo_vars = T, geometry = TRUE) %>%
	mutate(FamPovPer_BG = (B17010_002E/B17010_001E)*100,
	       PopPovPer_BG = (B17001_002E/B17001_001E)*100,
	       TotalPopPov_BG = B17001_002E) 

chelsea.pov <- merge(ch.pov.get, ch.MA.pov, by.x = 'NAME.y', by.y='NAME')


tm_shape(chelsea.pov[c(2:5),]) + tm_polygons("FamPovPer_BG", style="quantile", palette="Purples", border.col = "darkgray", title="Family in Poverty", popup.format = list(fun = make_percent), legend.format = list(fun = make_percent))  + tm_view(alpha = 0.85,view.legend.position = c("right", "bottom"), basemaps = "Stamen.TonerLite")


```

\


### Elderly People Living Alone
\
```{r echo=FALSE, warning=FALSE, message=FALSE, results='hide', comment=""}

########################################     Elderly People Living Alone    ##############################################################
MA.old.ch <- read_acs5year(year = endyr,
                     states = j,
                     table_contents = c("B11010_005", # Male householder, Living alone, 65 years and over
																					"B11010_012", # Female householder, Living alone, 65 years and over
																					"B11010_001"),
                     areas = area,
                     summary_level = "tract") 
MA.old.ch <- MA.old.ch[MA.old.ch$NAME == "Census Tract 300, Penobscot County, Maine",] %>%
  rename(City = area) %>% 
  mutate(OldAlonePer = (B11010_012 + B11010_005)/B11010_001) %>%
	dplyr::select(City, OldAlonePer) %>%
  left_join(Chelsea.MA,MA.old.ch, by="City")

Chelsea.MA <- cbind(MA.old.ch) 

## GIVE US NA
```
* The percentage of householders, who are living alone and are older than 65 years old is `r (round(Chelsea.MA$OldAlonePer,4))*100`% in `r Chelsea.MA$City`.\

\

## Educational Institutions
\


\

### Number/Type of Schools
\

* `r Chelsea.MA$City` has 3 public school district, which contains 708 students and 4 public schools. Also, `r  Chelsea.MA$City` has no private schools.(https://nces.ed.gov/) 

\


```{r echo=FALSE, warning=FALSE, message=FALSE, results='hide', comment=""}
Chelsea.MA$SchoolsDistrict <- 3
Chelsea.MA$SchoolsDistStudents <- 708
Chelsea.MA$PubSchools <- 4
Chelsea.MA$PvSchools <- 0
```


\


### Number/Type of Universities
\

* `r Chelsea.MA$City` has no university in 04462 ZIP Code.(https://nces.ed.gov/) 

```{r echo=FALSE, warning=FALSE, message=FALSE, results='hide', comment=""}
Chelsea.MA$NumUni <- 0
```


\

###	Employment Numbers per Institution
\


* The major School District in `r Chelsea.MA$City` contains 42 teachers and 67 other staffs.(http://www.usa.com/)


\

### Number of PK-12 Enrollment in Schools
\

```{r echo=FALSE, warning=FALSE, message=FALSE, results='hide', comment=""}
############################################################################################################################

                                             # ----------------------------- #
                                             #    Educational Institutions   #
                                             #-------------------------------#

# Number of PK-12 Enrollment at block group level
MA.Enrol.ch <- read_acs5year(year = endyr,
                     states = j,
                     table_contents = c("B14001_003", # Enrolled in preschool
																					"B14001_004", # Enrolled in kindergarten
																					"B14001_005", # Enrolled in grade 1 to 4
																					"B14001_006", # Enrolled in grade 5 to 8
																					"B14001_007"),
                     areas = area,
                     summary_level = "tract") 
MA.Enrol.ch <- MA.Enrol.ch[MA.Enrol.ch$NAME == "Census Tract 300, Penobscot County, Maine",] %>%
  rename(City = area) %>% 
  mutate(EnrolPK12 = B14001_003 + B14001_004 + B14001_005 + B14001_006 + B14001_007) %>%
	dplyr::select(City, EnrolPK12) %>%
  left_join(Chelsea.MA, MA.Enrol.ch, by="City")

Chelsea.MA <- cbind(MA.Enrol.ch)

```
* The total number of students, who enrolled in preschools to 12 grade levels in `r Chelsea.MA$City` schools is `r Chelsea.MA$EnrolPK12`.\

\


## Employment
\


### Number of Jobs within Community Boundaries
\

```{r echo=FALSE, warning=FALSE, message=FALSE, results='hide', comment=""}

########################################################################################################################################

                                                #----------------------#
                                                #       EMPLOYMENT     #
                                                #----------------------#


##########################################    Number of Jobs within Community Boundaries   #############################################
MA.Num.Jobs.ch <- read_acs5year(year = endyr,
                     states = j,
                     table_contents = "B08008_003",
                     areas = area,
                     summary_level = "tract")
MA.Num.Jobs.ch <- MA.Num.Jobs.ch[MA.Num.Jobs.ch$NAME == "Census Tract 300, Penobscot County, Maine",] %>%
  rename(City = area,
         NumJobs = B08008_003) %>% 
	dplyr::select(City, NumJobs) %>%
  left_join(Chelsea.MA, MA.Num.Jobs.ch, by="City")

Chelsea.MA <- cbind(MA.Num.Jobs.ch)
```
* The number of jobs within community boundaries in `r Chelsea.MA$City` is `r Chelsea.MA$NumJobs`.\

Followed is the map of Number of Jobs within Community Boundaries of the city.\

```{r echo=FALSE, warning=FALSE, message=FALSE, comment="",results='hide'}

## MAP

ch.MA.job <- read_acs5year(year = endyr,
                     states = j,
                     table_contents = "B08008_003",
                     areas = area,
                     summary_level = "block group")

ch.job.get <- tidycensus::get_acs(geography = "block group",
													 survey="acs5",
													 state=j,
													 county = c,
													 year = endyr,
													 output = "wide",
													 variables = 'B08008_003', 
													 keep_geo_vars = T, geometry = TRUE) %>%
	rename(NumJobs = B08008_003E) 

chelsea.job <- merge(ch.job.get, ch.MA.job, by.x = 'NAME.y', by.y='NAME')


tm_shape(chelsea.job[c(2:5),]) + tm_polygons("NumJobs", style="pretty", palette="Reds", border.col = "darkgray", title="Number of Jobs") + tm_view(alpha = 0.85,view.legend.position = c("right", "bottom"), basemaps = "Stamen.TonerLite")
```

\


### Labor Force (Total number, participation)
\
```{r echo=FALSE, warning=FALSE, message=FALSE, results='hide', comment=""}
############################################# Labor Force (Total number, participation)  ####################################

MA.Labors.ch <- read_acs5year(year = endyr,
                     states = j,
                     table_contents = c("B23025_002", # Number of Labor Force 16 years and over
														              "B23025_003", # Number of Civilian Labor Force 16 years and over
														              "B23025_006", # Number of Armed Forces 16 years and over
														              "B23025_001"),
                     areas = area,
                     summary_level = "tract")
MA.Labors.ch <-MA.Labors.ch[MA.Labors.ch$NAME == "Census Tract 300, Penobscot County, Maine",] %>%
  rename(City = area) %>% 
  mutate(LabForce = B23025_002,
	       LabForcePer = B23025_002/B23025_001,
	       CivLabForce = B23025_003,
	       CivLabForcePer = B23025_003/B23025_001,
	       Military = B23025_006,
	       MilitaryPer = 	B23025_006/B23025_001,
	       TotalWorkers = B23025_001) %>%
	dplyr::select(City, LabForce, LabForcePer, CivLabForce, CivLabForcePer, Military, MilitaryPer, TotalWorkers)  %>%
  left_join(Chelsea.MA, MA.Labors.ch, by="City")

Chelsea.MA <- cbind(MA.Labors.ch)
```
* The number of labor force, who are 16 years and older in `r Chelsea.MA$City` is 1806 which is `r round(Chelsea.MA$LabForcePer,2)` of total number of labors (unemployed and not in labor force) of which 1806 are civilian labor forces and `r Chelsea.MA$Military` are armed labor forces.\

Followed is the map of Labor Force of the city.\

```{r echo=FALSE, warning=FALSE, message=FALSE, comment="",results='hide'}
## MAP

ch.MA.labors <- read_acs5year(year = endyr,
                     states = j,
                     table_contents = c("B23025_002", # Number of Labor Force 16 years and over
														              "B23025_003", # Number of Civilian Labor Force 16 years and over
														              "B23025_006", # Number of Armed Forces 16 years and over
														              "B23025_001"),
                     areas = area,
                     summary_level = "block group")

ch.labors.get <- tidycensus::get_acs(geography = "block group",
													 survey="acs5",
													 state=j,
													 county = c,
													 year = endyr,
													 output = "wide",
													 variables = c("B23025_002", # Number of Labor Force 16 years and over
														              "B23025_003", # Number of Civilian Labor Force 16 years and over
														              "B23025_006", # Number of Armed Forces 16 years and over
														              "B23025_001"), 
													 keep_geo_vars = T, geometry = TRUE) %>%
	mutate(LabForce = B23025_002E,
	       LabForcePer = B23025_002E/B23025_001E,
	       CivLabForce = B23025_003E,
	       CivLabForcePer = B23025_003E/B23025_001E,
	       Military = B23025_006E,
	       MilitaryPer = 	B23025_006E/B23025_001E,
	       TotalWorkers = B23025_001E) 

chelsea.labors <- merge(ch.labors.get, ch.MA.labors, by.x = 'NAME.y', by.y='NAME')


tm_shape(chelsea.labors[c(2:5),]) + tm_polygons("LabForce", style="pretty", palette="seq", border.col = "darkgray", title="Number of Labors") + tm_view(alpha = 0.85,view.legend.position = c("right", "bottom"), basemaps = "Stamen.TonerLite")
```

\

### Employed and Unemployed
\
```{r echo=FALSE, warning=FALSE, message=FALSE, results='hide', comment=""}
#####################################################    Employed and Unemployed   ####################################

MA.Emp.Unemp.ch <- read_acs5year(year = endyr,
                     states = j,
                     table_contents = c("B23025_003", # Number of civilian Labor Force 16 years and over
														              "B23025_004", # Number of Employed individuals 16 years and over
														              "B23025_005"),
                     areas = area,
                     summary_level = "tract") 
MA.Emp.Unemp.ch <- MA.Emp.Unemp.ch[MA.Emp.Unemp.ch$NAME == "Census Tract 300, Penobscot County, Maine",] %>%
  rename(City = area) %>%
  mutate(EmpRate = B23025_004/B23025_003,
	       UnempRate = B23025_005/B23025_003,
	       EmployedCLF = B23025_004,
	       UnemployedCLF = B23025_005) %>%
	dplyr::select(City ,EmpRate, UnempRate, EmployedCLF, UnemployedCLF) %>%
  left_join(Chelsea.MA, MA.Emp.Unemp.ch, by="City")

Chelsea.MA <- cbind(MA.Emp.Unemp.ch)

```
* The rate of employment in `r Chelsea.MA$City` is `r round(Chelsea.MA$EmpRate,2)*100`%.\
* The rate of unemployment in `r Chelsea.MA$City` is `r round(Chelsea.MA$UnempRate,2)*100`%.\

Followed is Employed and Unemployed map of the city.\

```{r echo=FALSE, warning=FALSE, message=FALSE, comment="",results='hide'}
## MAP
ch.MA.emp <- read_acs5year(year = endyr,
                     states = j,
                     table_contents = c("B23025_003", # Number of civilian Labor Force 16 years and over
														              "B23025_004", # Number of Employed individuals 16 years and over
														              "B23025_005"),
                     areas = area,
                     summary_level = "block group")

ch.emp.get <- tidycensus::get_acs(geography = "block group",
													 survey="acs5",
													 state=j,
													 county = c,
													 year = endyr,
													 output = "wide",
													 variables = c("B23025_003", # Number of civilian Labor Force 16 years and over
														              "B23025_004", # Number of Employed individuals 16 years and over
														              "B23025_005"), 
													 keep_geo_vars = T, geometry = TRUE) %>%
	mutate(EmpRate_BG = (B23025_004E/B23025_003E)*100,
	       UnempRate_BG = (B23025_005E/B23025_003E)*100,
	       EmployedCLF_BG = B23025_004E,
	       UnemployedCLF_BG = B23025_005E) 

chelsea.emp <- merge(ch.emp.get, ch.MA.emp, by.x = 'NAME.y', by.y='NAME')

tm_shape(chelsea.emp[c(2:5),]) + tm_polygons(c("EmpRate_BG","UnempRate_BG"), style="quantile", palette=list("Greens","seq"), border.col = "darkgray", title=c("Employment rate","Unemployment rate"), popup.format = list(fun = make_percent), legend.format = list(fun = make_percent))+ tm_view(alpha = 0.85, basemaps = "Stamen.TonerLite") + tm_layout(legend.position = c("left","bottom"))

```

\


### Major Employers by NCAIS Code
\
```{r echo=FALSE, warning=FALSE, message=FALSE, results='hide', comment=""}

################################    Major Employers by NCAIS Code    ##########################################
MA.Emp.NCAIS.ch <- read_acs5year(year = endyr,
                     states = j,
                     table_contents = c("C24030_002", # Number of male by industry for the civilian employed, 16 years and over
														              "C24030_029", # Number of female by industry for the civilian employed, 16 years and over
														              "C24030_001"),
                     areas = area,
                     summary_level = "tract") 
MA.Emp.NCAIS.ch <- MA.Emp.NCAIS.ch[MA.Emp.NCAIS.ch$NAME == "Census Tract 300, Penobscot County, Maine",] %>%
  rename(City = area) %>%
  mutate(MaleNCAIS = C24030_002,
	       FemaleNCAIS = C24030_029,
	       MaleNCAISPer = C24030_002/C24030_001,
	       FemaleNCAISPer = C24030_029/C24030_001) %>%
	dplyr::select(City , MaleNCAIS, FemaleNCAIS, MaleNCAISPer, FemaleNCAISPer) %>%
  left_join(Chelsea.MA, MA.Emp.NCAIS.ch, by="City")

Chelsea.MA <- cbind(MA.Emp.NCAIS.ch)

```
* Number of male by industry for the civilian employed in `r Chelsea.MA$City` is `r Chelsea.MA$MaleNCAIS` which is `r round(Chelsea.MA$MaleNCAISPer,2)` of population, who are 16 years and over.\
* Number of female by industry for the civilian employed in `r Chelsea.MA$City` is `r Chelsea.MA$FemaleNCAIS` which is `r round(Chelsea.MA$FemaleNCAISPer,2)` of population, who are 16 years and over.\

Followed is the map of the Major Employers by NCAIS Code of the city.\

```{r echo=FALSE, warning=FALSE, message=FALSE, comment="",results='hide'}
## MAP
ch.MA.majorEmp <- read_acs5year(year = endyr,
                     states = j,
                     table_contents = c("C24030_002", # Number of male by industry for the civilian employed, 16 years and over
														              "C24030_029", # Number of female by industry for the civilian employed, 16 years and over
														              "C24030_001"),
                     areas = area,
                     summary_level = "block group")

ch.majorEmp.get <- tidycensus::get_acs(geography = "block group",
													 survey="acs5",
													 state=j,
													 county = c,
													 year = endyr,
													 output = "wide",
													 variables = c("C24030_002", # Number of male by industry for the civilian employed, 16 years and over
														              "C24030_029", # Number of female by industry for the civilian employed, 16 years and over
														              "C24030_001"), 
													 keep_geo_vars = T, geometry = TRUE) %>%
  mutate(MaleNCAIS_BG = C24030_002E,
	       FemaleNCAIS_BG = C24030_029E,
	       MaleNCAISPer_BG = (C24030_002E/C24030_001E)*100,
	       FemaleNCAISPer_BG = (C24030_029E/C24030_001E)*100) 

chelsea.majorEmp <- merge(ch.majorEmp.get, ch.MA.majorEmp, by.x = 'NAME.y', by.y='NAME')


tm_shape(chelsea.majorEmp[c(2:5),]) + tm_polygons(c("MaleNCAIS_BG","FemaleNCAIS_BG"), style="pretty", palette=list("Greens","seq"), border.col = "darkgray", title=c("MaleNCAIS","FemaleNCAIS")) + tm_view(alpha = 0.85, basemaps = "Stamen.TonerLite") + tm_layout(legend.position = c("left","bottom"))

```

\

```{r echo=FALSE, warning=FALSE, message=FALSE, results='hide', comment=""}
tm_shape(chelsea.majorEmp[c(2:5),]) + tm_polygons(c("MaleNCAISPer_BG","FemaleNCAISPer_BG"), style="quantile", palette=list("Blues","Reds"), border.col = "darkgray", title=c("MaleNCAIS Pecentage","FemaleNCAIS Percentage"),popup.format = list(fun = make_percent), legend.format = list(fun = make_percent)) + tm_view(alpha = 0.85, basemaps = "Stamen.TonerLite") + tm_layout(legend.position = c("left","bottom"))
```




\

### Top Industries by NCAIS Code
\
```{r echo=FALSE, warning=FALSE, message=FALSE, results='hide', comment=""}
######################################    Top Industries by NCAIS Code   ##################################################

MA.Industries.ch <- read_acs5year(year = endyr,
                     states = j,
                     table_contents = c("C24030_003", # Agriculture, forestry, fishing and hunting, and mining (Male)
														              "C24030_030", # Agriculture, forestry, fishing and hunting, and mining (Female)
														              "C24030_006", # Construction (Male)
														              "C24030_033", # Construction (Female)
														              "C24030_007", # Manufacturing (Male)
														              "C24030_034", # Manufacturing (Female)
														              "C24030_008", # Wholesale trade (Male)
														              "C24030_035", # Wholesale trade (Female)
														              "C24030_009", # Retail trade (Male)
														              "C24030_036", # Retail trade (Female)
														              "C24030_010", # Transportation and warehousing, and utilities (Male)
														              "C24030_037", # Transportation and warehousing, and utilities (Female)
														              "C24030_013", # Information (Male)
														              "C24030_040", # Information (Female)
														              "C24030_014", # Finance and insurance, and real estate, and rental and leasing (Male)
														              "C24030_041", # Finance and insurance, and real estate, and rental and leasing (Female)
														              "C24030_017", # Professional, scientific, and management, and administrative, and waste management services (Male)
														              "C24030_044", # Professional, scientific, and management, and administrative, and waste management services (Female)
														              "C24030_021", # Educational services, and health care and social assistance (Male)
														              "C24030_048", # Educational services, and health care and social assistance (Female)
														              "C24030_024", # Arts, entertainment, and recreation, and accommodation and food services (Male)
														              "C24030_051", # Arts, entertainment, and recreation, and accommodation and food services (Female)
														              "C24030_027", # Other services, except public administration (Male)
														              "C24030_054", # Other services, except public administration (Female)
														              "C24030_028", # Public administration (Male)
														              "C24030_055", # Public administration (Female)
														              "C24030_001"), # Total,
                     areas = area,
                     summary_level = "tract") 
MA.Industries.ch <-MA.Industries.ch[MA.Industries.ch$NAME == "Census Tract 300, Penobscot County, Maine",] %>%
  rename(City = area) %>%
  mutate(Agriculture = (C24030_003 + C24030_030)/C24030_001,
	       Construction = (C24030_006 + C24030_033)/C24030_001,
	       Manufacturing = (C24030_007 + C24030_034)/C24030_001,
	       WholesaleTrade = (C24030_008 + C24030_035)/C24030_001,
	       RetailTrade = (C24030_009 + C24030_036)/C24030_001,
	       Transportation = (C24030_010 + C24030_037)/C24030_001,
	       Information = (C24030_013 + C24030_040)/C24030_001,
	       Finance_Ins = (C24030_014 + C24030_041)/C24030_001,
	       Professional = (C24030_017 + C24030_044)/C24030_001,
	       Educ_Health_SocSvcs = (C24030_021 + C24030_048)/C24030_001,
	       RecreationEtc = (C24030_024 + C24030_051)/C24030_001,
	       OtherIndustries = (C24030_027 + C24030_054)/C24030_001,
	       PublicAdmin = (C24030_028 + C24030_055)/C24030_001) %>%
	dplyr::select(City,Agriculture,Construction,Manufacturing,WholesaleTrade,RetailTrade,Transportation,Information,
	       Finance_Ins,Professional,Educ_Health_SocSvcs,RecreationEtc,OtherIndustries,PublicAdmin) %>%
  left_join(Chelsea.MA, MA.Industries.ch, by="City")

Chelsea.MA <- cbind(MA.Industries.ch)

```
* The percentage of people, who are in agriculture, forestry, fishing, hunting, and mining in `r Chelsea.MA$City` is `r (round(Chelsea.MA$Agriculture,4))*100`%.\

* The percentage of people, who are in construction in `r Chelsea.MA$City` state is `r (round(Chelsea.MA$Construction,4))*100`%.\

* The percentage of people, who are in manufacturing in `r Chelsea.MA$City` state is `r (round(Chelsea.MA$Manufacturing,4))*100`%.\

* The percentage of people, who are in wholesale trade in `r Chelsea.MA$City` state is `r (round(Chelsea.MA$WholesaleTrade,4))*100`%.\

* The percentage of people, who are in retail trade in `r Chelsea.MA$City` state is `r (round(Chelsea.MA$RetailTrade,4))*100`%.\

* The percentage of people, who are in transportation and warehousing, and utilities in `r Chelsea.MA$City`state is `r (round(Chelsea.MA$Transportation,4))*100`%.\

* The percentage of people, who are in information in `r Chelsea.MA$City` state is `r (round(Chelsea.MA$Information,4))*100`%.\

* The percentage of people, who are in finance and insurance, and real estate, and rental and leasing in `r Chelsea.MA$City` state is `r (round(Chelsea.MA$Finance_Ins,4))*100`%.\

* The percentage of people, who are in professional, scientific, and management, and administrative, and waste management services in `r Chelsea.MA$City` state is `r (round(Chelsea.MA$Professional,4))*100`%.\

* The percentage of people, who are in educational services, and health care and social assistance in `r Chelsea.MA$City` state is `r (round(Chelsea.MA$Educ_Health_SocSvcs,4))*100`%.\

* The percentage of people, who are in arts, entertainment, and recreation, and accommodation and food services in `r Chelsea.MA$City` state is `r (round(Chelsea.MA$RecreationEtc,4))*100`%.\

* The percentage of people, who are in other services, except public administration in `r Chelsea.MA$City` state is `r (round(Chelsea.MA$OtherIndustries,4))*100`%.\

* The percentage of people, who are in Public administration in `r Chelsea.MA$City` state is `r (round(Chelsea.MA$PublicAdmin,4))*100`%.\
\

### Earning Breakdown 
\

```{r echo=FALSE, warning=FALSE, message=FALSE, results='hide', comment=""}
###########################    Earning Breakdown   ####################################
MA.Earning.ch <- read_acs5year(year = endyr,
                     states = j,
                     table_contents = "B20017_001",
                     areas = area,
                     summary_level = "tract") 
MA.Earning.ch <- MA.Earning.ch[MA.Earning.ch$NAME == "Census Tract 300, Penobscot County, Maine",] %>%
  rename(City = area) %>%
  mutate(MedEarning = B20017_001) %>%
	dplyr::select(City,MedEarning) %>%
  left_join(Chelsea.MA, MA.Earning.ch, by="City")

Chelsea.MA <- cbind(MA.Earning.ch)
```
* The median earning in `r Chelsea.MA$City` is 26402 Dollar.\

Followed is the Median Earning map of the city.\

```{r echo=FALSE, warning=FALSE, message=FALSE, comment="",results='hide'}
## MAP
ch.MA.earning <- read_acs5year(year = endyr,
                     states = j,
                     table_contents = "B20017_001",
                     areas = area,
                     summary_level = "block group")

ch.earning.get <- tidycensus::get_acs(geography = "block group",
													 survey="acs5",
													 state=j,
													 county = c,
													 year = endyr,
													 output = "wide",
													 variables = 'B20017_001', 
													 keep_geo_vars = T, geometry = TRUE) %>%
  mutate(MedEarning = B20017_001E)

chelsea.earning <- merge(ch.earning.get, ch.MA.earning, by.x = 'NAME.y', by.y='NAME')

make_dollar <- function(x, digits = 0) {
  paste0("$", formatC(x, digits = digits, format = "f", big.mark = ","))
}


tm_shape(chelsea.earning[c(2:5),]) + tm_polygons("MedEarning" , palette="seq", style = "quantile", title = "Median Earning", popup.format = list(fun = make_dollar), legend.format = list(fun = make_dollar)) + tm_view(alpha = 0.85,view.legend.position = c("right", "bottom"),basemaps = "Stamen.TonerLite") + tm_borders(col = "darkgray") 


```

\


### Employee Age
\
```{r echo=FALSE, warning=FALSE, message=FALSE, results='hide', comment=""}
#############################################    Employee Age   #############################################################

MA.Emp.Age.ch <- read_acs5year(year = endyr,
                     states = j,
                     table_contents = c("B23001_005","B23001_007","B23001_012","B23001_014","B23001_019","B23001_021","B23001_026","B23001_028", # Age 29 or younger male
								"B23001_033","B23001_035","B23001_040","B23001_042","B23001_047","B23001_049", # Age 30 to 54 male
								"B23001_054","B23001_056","B23001_061","B23001_063","B23001_068","B23001_070","B23001_075","B23001_080","B23001_085", # Age 55 or older male
								"B23001_091","B23001_093","B23001_098","B23001_100","B23001_105","B23001_107","B23001_112","B23001_114", # Age 29 or younger female
								"B23001_119","B23001_121","B23001_126","B23001_128","B23001_133","B23001_135", # Age 30 to 54 female
								"B23001_140","B23001_142","B23001_147","B23001_149","B23001_154","B23001_156","B23001_161","B23001_166","B23001_171", # Age 55 or older female
														              "B23001_001"),
                     areas = area,
                     summary_level = "tract") 
MA.Emp.Age.ch <- MA.Emp.Age.ch[MA.Emp.Age.ch$NAME == "Census Tract 300, Penobscot County, Maine",] %>%
  rename(City = area) %>%
  mutate(AgeEmpU29 = B23001_005+B23001_007+B23001_012+B23001_014+B23001_019+B23001_021+B23001_026+B23001_028+
	         B23001_091+B23001_093+B23001_098+B23001_100+B23001_105+B23001_107+B23001_112+B23001_114,
	       AgeEmp3054 = B23001_033+B23001_035+B23001_040+B23001_042+B23001_047+B23001_049+
	         B23001_119+B23001_121+B23001_126+B23001_128+B23001_133+B23001_135,
	       AgeEmpO55 = B23001_054+B23001_056+B23001_061+B23001_063+B23001_068+B23001_070+B23001_075+B23001_080+B23001_085+
	        B23001_140+B23001_142+B23001_147+B23001_149+B23001_154+B23001_156+B23001_161+B23001_166+B23001_171,
         
         AgeEmpU29Per_BG = (B23001_005+B23001_007+B23001_012+B23001_014+B23001_019+B23001_021+B23001_026+B23001_028+
	         B23001_091+B23001_093+B23001_098+B23001_100+B23001_105+B23001_107+B23001_112+B23001_114)/B23001_001,
	        AgeEmp3054Per_BG = (B23001_033+B23001_035+B23001_040+B23001_042+B23001_047+B23001_049+
	         B23001_119+B23001_121+B23001_126+B23001_128+B23001_133+B23001_135)/B23001_001,
	       AgeEmpO55Per_BG = (B23001_054+B23001_056+B23001_061+B23001_063+B23001_068+B23001_070+B23001_075+B23001_080+B23001_085+
	        B23001_140+B23001_142+B23001_147+B23001_149+B23001_154+B23001_156+B23001_161+B23001_166+B23001_171)/B23001_001) %>%

	dplyr::select(City ,AgeEmpU29, AgeEmp3054, AgeEmpO55, AgeEmpU29Per_BG,AgeEmp3054Per_BG,AgeEmpO55Per_BG) %>%
  left_join(Chelsea.MA, MA.Emp.Age.ch, by="City")

Chelsea.MA <- cbind(MA.Emp.Age.ch)

```
* The number of workers, who are younger than 29 years old in `r Chelsea.MA$City` is `r Chelsea.MA$AgeEmpU29` which is `r (round(Chelsea.MA$AgeEmpU29Per,4))*100`% of total number of workers.\

* The number of workers, who are between 30 and 54 years old in `r Chelsea.MA$City` is `r Chelsea.MA$AgeEmp3054` which is `r (round(Chelsea.MA$AgeEmp3054Per,4))*100`% of total number of workers.\

* The number of workers, who are older than 55 years old in `r Chelsea.MA$City` is `r Chelsea.MA$AgeEmpO55` which is `r (round(Chelsea.MA$AgeEmpO55Per,4))*100`% of total number of workers.\

\


### Employee Race
\
```{r echo=FALSE, warning=FALSE, message=FALSE, results='hide', comment=""}
#############################################      Employee Race    ############################################################
MA.Emp.Race.ch <- read_acs5year(year = endyr,
                     states = j,
                     table_contents = c("C23002A_007","C23002A_012","C23002A_020","C23002A_025","C23002A_005","C23002A_018", # White alone
														              "C23002B_007","C23002B_012","C23002B_020","C23002B_025","C23002B_005","C23002B_018", # Black or African American alone
														              "C23002D_007","C23002D_012","C23002D_020","C23002D_025","C23002D_005","C23002D_018", # Asian alone
														              "C23002I_007","C23002I_012","C23002I_020","C23002I_025","C23002I_005","C23002I_018", # Hispanic or Latino alone
														              "C23002G_007","C23002G_012","C23002G_020","C23002G_025","C23002G_005","C23002G_018",# Two or more races
														              "B23001_001"),
                     areas = area,
                     summary_level = "tract")
MA.Emp.Race.ch <- MA.Emp.Race.ch[MA.Emp.Race.ch$NAME == "Census Tract 300, Penobscot County, Maine",] %>%
  rename(City = area) %>%
  mutate(WhiteEmp = C23002A_007+C23002A_012+C23002A_020+C23002A_025+C23002A_005+C23002A_018,
	       BlackEmp = C23002B_007+C23002B_012+C23002B_020+C23002B_025+C23002B_005+C23002B_018,
	       AsianEmp = C23002D_007+C23002D_012+C23002D_020+C23002D_025+C23002D_005+C23002D_018,
	       HispanicEmp = C23002I_007+C23002I_012+C23002I_020+C23002I_025+C23002I_005+C23002I_018,
	       TwoOrMoreEmp = C23002G_007+C23002G_012+C23002G_020+C23002G_025+C23002G_005+C23002G_018,
         WhiteEmpPer = (C23002A_007+C23002A_012+C23002A_020+C23002A_025+C23002A_005+C23002A_018)/B23001_001,
	       BlackEmpPer = (C23002B_007+C23002B_012+C23002B_020+C23002B_025+C23002B_005+C23002B_018)/B23001_001,
	       AsianEmpPer = (C23002D_007+C23002D_012+C23002D_020+C23002D_025+C23002D_005+C23002D_018)/B23001_001,
	       HispanicEmpPer = (C23002I_007+C23002I_012+C23002I_020+C23002I_025+C23002I_005+C23002I_018)/B23001_001,
	       TwoOrMoreEmpPer = (C23002G_007+C23002G_012+C23002G_020+C23002G_025+C23002G_005+C23002G_018)/B23001_001) %>%
	dplyr::select(City ,WhiteEmp, BlackEmp, AsianEmp, HispanicEmp, TwoOrMoreEmp,
	       WhiteEmpPer, BlackEmpPer, AsianEmpPer, HispanicEmpPer, TwoOrMoreEmpPer) %>%
  left_join(Chelsea.MA, MA.Emp.Race.ch, by="City")

Chelsea.MA <- cbind(MA.Emp.Race.ch)

```

* The number of white workers in `r Chelsea.MA$City` is `r Chelsea.MA$WhiteEmp` which is `r (round(Chelsea.MA$WhiteEmpPer,4))*100`% of total number of workers.\
* The number of black workers in `r Chelsea.MA$City` is `r Chelsea.MA$BlackEmp` which is `r (round(Chelsea.MA$BlackEmpPer,4))*100`% of total number of workers.\
* The number of asian workers in `r Chelsea.MA$City` is `r Chelsea.MA$AsianEmp` which is `r (round(Chelsea.MA$AsianEmpPer,4))*100`% of total number of workers.\
* The number of hispanic or latino workers in `r Chelsea.MA$City` is `r Chelsea.MA$HispanicEmp` which is `r (round(Chelsea.MA$HispanicEmpPer,4))*100`% of total number of workers.\
* The number of workers, who have two or more races in `r Chelsea.MA$City` is `r Chelsea.MA$TwoOrMoreEmp` which is `r (round(Chelsea.MA$TwoOrMoreEmpPer,4))*100`% of total number of workers.\
\


### Employee Education
\
```{r message=FALSE, warning=FALSE, ,echo=FALSE, results='hide', comment=""}
#############################################   Employee Education    ########################################################

MA.Educ.Emp.ch <- read_acs5year(year = endyr,
                     states = j,
                     table_contents = c("B16010_003", # Less than high school graduate in labor force
														              "B16010_016", # High school graduate in labor force 
														              "B16010_029", # Some college or associate's degree in labor force
														              "B16010_042",
														              "B16010_001"), # Total
                     areas = area,
                     summary_level = "tract") 
MA.Educ.Emp.ch <- MA.Educ.Emp.ch[MA.Educ.Emp.ch$NAME == "Census Tract 300, Penobscot County, Maine",] %>%
  rename(City = area) %>%
  	mutate(LessThanHSLabor = B16010_003,
	       HSGradLabor = B16010_016,
	       SomeCollLabor = B16010_029,
	       BachOrHigherLabor = B16010_042,
	       LessThanHSLaborPer = LessThanHSLabor/B16010_001,
	       HSGradLaborPer = HSGradLabor/B16010_001,
	       SomeCollLaborPer = SomeCollLabor/B16010_001,
	       BachOrHigherLaborPer = BachOrHigherLabor/B16010_001) %>%
	dplyr::select(City ,LessThanHSLabor, HSGradLabor, SomeCollLabor, BachOrHigherLabor,
	       LessThanHSLaborPer, HSGradLaborPer, SomeCollLaborPer, BachOrHigherLaborPer) %>%
  left_join(Chelsea.MA, MA.Educ.Emp.ch, by="City")

Chelsea.MA <- cbind(MA.Educ.Emp.ch)
```
* The number of workers, who have less than high school degree in `r Chelsea.MA$City` is `r Chelsea.MA$LessThanHSLabor` which is `r (round(Chelsea.MA$LessThanHSLaborPer,4))*100`% of total number of workers.\
* The number of workers, who have high school degree in `r Chelsea.MA$City` is `r Chelsea.MA$HSGradLabor` which is `r (round(Chelsea.MA$HSGradLaborPer,4))*100`% of total number of workers.\
* The number of workers, who have some college or associates degree in `r Chelsea.MA$City` is `r Chelsea.MA$SomeCollLabor` which is `r (round(Chelsea.MA$SomeCollLaborPer,4))*100`% of total number of workers.\
* The number of workers, who have bachelors degree or higher in `r Chelsea.MA$City` is `r Chelsea.MA$BachOrHigherLabor` which is `r (round(Chelsea.MA$BachOrHigherLaborPer,4))*100`% of total number of workers.\

\

### Employee Sex
\
```{r echo=FALSE, warning=FALSE, message=FALSE, results='hide', comment=""}
#############################################    Employee Sex   #############################################################

MA.Emp.Sex.ch <- read_acs5year(year = endyr,
                     states = j,
                     table_contents = c("B23001_005", # Employed males
														              "B23001_007", 
														              "B23001_012",
														              "B23001_014",
														              "B23001_019",
														              "B23001_021",
														              "B23001_026",
														              "B23001_028",
														              "B23001_033",
														              "B23001_035",
														              "B23001_040",
														              "B23001_042",
														              "B23001_047",
														              "B23001_049",
														              "B23001_054",
														              "B23001_056",
														              "B23001_061",
														              "B23001_063",
														              "B23001_068",
														              "B23001_070",
														              "B23001_075",
														              "B23001_080",
														              "B23001_085",
														              "B23001_091", # Employed females
														              "B23001_093",
														              "B23001_098",
														              "B23001_100",
														              "B23001_105",
														              "B23001_107",
														              "B23001_112",
														              "B23001_114",
														              "B23001_119",
														              "B23001_121",
														              "B23001_126",
														              "B23001_128",
														              "B23001_133",
														              "B23001_135",
														              "B23001_140",
														              "B23001_142",
														              "B23001_147",
														              "B23001_149",
														              "B23001_154",
														              "B23001_156",
														              "B23001_161",
														              "B23001_166",
														              "B23001_171",
														              "B23001_001"),
                     areas = area,
                     summary_level = "tract") 
MA.Emp.Sex.ch <- MA.Emp.Sex.ch[MA.Emp.Sex.ch$NAME == "Census Tract 300, Penobscot County, Maine",] %>%
  rename(City = area) %>%
  	mutate(EmpMale = B23001_005 + B23001_007 + B23001_012 + B23001_014 + B23001_019 + B23001_021 + B23001_026 + B23001_028 + B23001_033 + B23001_035 + B23001_040 + B23001_042 + B23001_047 + B23001_049 + B23001_054 + B23001_056 + B23001_061 + B23001_063 + B23001_068+ B23001_070 + B23001_075 + B23001_080 + B23001_085,
	  EmpFemale = B23001_091 + B23001_093 + B23001_098 + B23001_100 + B23001_105 + B23001_107 + B23001_112 + B23001_114 + B23001_119 + B23001_121 + B23001_126+ B23001_128 + B23001_133 + B23001_135 + B23001_140 + B23001_142 + B23001_147 + B23001_149 + B23001_154 + B23001_156 + B23001_161 + B23001_166 + B23001_171,
	  TotalEmp = EmpMale + EmpFemale,
	  EmpMalePer = EmpMale/	B23001_001,
	  EmpFemalePer = EmpFemale/B23001_001) %>%
	dplyr::select(City, EmpMale, EmpFemale, TotalEmp, EmpMalePer, EmpFemalePer) %>%
  left_join(Chelsea.MA, MA.Emp.Sex.ch, by="City")

Chelsea.MA <- cbind(MA.Emp.Sex.ch)
```
* The number of male workers in `r Chelsea.MA$City` is `r Chelsea.MA$EmpMale` which is `r (round(Chelsea.MA$EmpMalePer,4))*100`% of total number of workers.\
* The number of female workers in `r Chelsea.MA$City` is `r Chelsea.MA$EmpFemale` which is `r (round(Chelsea.MA$EmpFemalePer,4))*100`% of total number of workers.\


\

## Healthcare
\

### Number of Hospitals
\

```{r echo=FALSE, warning=FALSE, message=FALSE, results='hide', comment=""}

Chelsea.MA$Hospitals <- 1
```


* `r Chelsea.MA$City` has 1 hospital with 25 beds.(https://www.ahd.com/)

\

###	Number of Nursing Homes 
\

```{r echo=FALSE, warning=FALSE, message=FALSE, results='hide', comment=""}
Chelsea.MA$NursingHomes <- 1
```


* `r Chelsea.MA$City` has 1 nursing home. (https://nursinghomedirectory.com/maine/millinocket)


## Municipal Finances
\

###	Revenue Streams
\


```{r echo=FALSE, warning=FALSE, message=FALSE, results='hide', comment=""}
Chelsea.MA$RevenueGeneralFundTotalFY17 <- 5885212
Chelsea.MA$RevenueTaxesFY17 <- 2813237 
Chelsea.MA$RevenueLicensesAndPermitsFY17 <- 200237
Chelsea.MA$RevenuesIntergovernmentalFY17 <- 1853210 
Chelsea.MA$RevenueTustionAndOtherChargesForServicesAndEducationFY17 <- 965458 
Chelsea.MA$RevenueUnclassifiedFY17 <- 48006 
Chelsea.MA$RevenueInvestmentEarningsFY17 <- 5064
```
* The total general fund revenues for town of east Millinocket, Maine is $5,885,212 in fiscal year 2017, which contains the following items: Taxes with 2,813,237 Dollar, licenses and permits with 200,237 Dollar, intergovernmental with 1,853,210 Dollar, tuition and other charges for services-education with 965,458 Dollar, unclassified items with 48,006 Dollar, and investment earnings with 5,064 Dollar.(https://www.eastmillinocketschools.org/assets/docs/Audits/Final%20Audit%20EMill%2006.30.17.pdf)

\

###	Major Expense Drivers
\

```{r echo=FALSE, warning=FALSE, message=FALSE, results='hide', comment=""}
Chelsea.MA$ExpendituresGeneralFundTotalFY17 <- 5588644
Chelsea.MA$ExpendituresGeneralGovernmentFY17 <- 449575 
Chelsea.MA$ExpendituresPublicSafetyFY17 <- 646539
Chelsea.MA$ExpendituresRoadsFY17 <- 310944
Chelsea.MA$ExpendituresHealthSocialServiceAndRecreastionFY17 <- 301851
Chelsea.MA$ExpendituresLibraryFY17 <- 1111 
Chelsea.MA$ExpendituresEducationFY17 <- 3712112 
Chelsea.MA$ExpendituresSpecialProjectsFY17 <- 2457 
Chelsea.MA$ExpendituresCemeteryFY17 <- 14909
Chelsea.MA$ExpendituresCountyTaxFY17 <- 97400
Chelsea.MA$ExpendituresCapitalOutlayFY17 <- 265
Chelsea.MA$ExpendituresDebtServiceFY17 <- 51481

```

* The total amount of general fund of expenditures for town of east Millinocket, Maine is $5,588,644 in Fiscal Year 2017 and the major expense driver is related to education, which is 3,712,112 Dollar. The other expenditures include general government with 449,575 Dollar, public safety with 646,539 Dollar, roads with 310,944 Dollar, health, social service and recreation with 301,851 Dollar, library with 1,111 Dollar, special projects with 2,457 Dollar, cemetery with 14,909 Dollar, county tax with 97,400 Dollar, capital outlay with 265 Dollar, and debt service with 51,481 Dollar.(https://www.eastmillinocketschools.org/assets/docs/Audits/Final%20Audit%20EMill%2006.30.17.pdf)


\


## Infrastructure
\

### Presence of Nearby Ports, Airports, Major Rail Facilities, or Intermodal Freight Facilities
\

* Millinocket includes an airport, whose name is Municipal Airport and it is located in the heart of the Katahdin Region. The airport has two runways, a main terminal, 3 municipally owned hangers, 3 privately owned hangers and 13 tie-downs. (https://millinocket.org/)



\

### Public Transport Reliability and Scope 
\

* There are some bus routs in Millinocket city. Also, there are two stations in Millinocket city that are used for Central Maine & Quebec Railway.(https://www.yellowpages.com/millinocket-me)

\

### Internet Access
\

* The most common hardwired broadband networks in Millinocket are Satellite lines and DSL lines, which provide about one hundred and 94.50 percent network reach in the area. "The average Millinocket customer has 2-3 providers with service for their location". About 10.88 of the local population has only one option for broadband service. Also, the common Internet choices in this city are Viasat Internet and HughesNet. "There are 9 internet providers in Millinocket with 5 of those offering residential service. Millinocket is the 386th most connected city in Maine and the fastest zip code in Millinocket for May 2020 is 04462".(https://broadbandnow.com/Maine/Millinocket)


\

## Food Insecurity
\

\

### Number of Grocery Stores
\

```{r echo=FALSE, warning=FALSE, message=FALSE, results='hide', comment=""}
Chelsea.MA$NumGrocery <- 34
```
* `r Chelsea.MA$City` has 34 Grocery Stores.(http://www.city-data.com/city/Millinocket-Maine.html)

\

### SNAP Recipients 
\


```{r echo=FALSE, warning=FALSE, message=FALSE, results='hide', comment=""}
##########################################################################################################################

                                                #----------------------#
                                                #    FOOD INSECURITY   #
                                                #----------------------#

####################################################  SNAP Recipients  ###################################################         

MA.Snap.ch <- read_acs5year(year = endyr,
                     states = j,
                     table_contents = c("B22001_002", # Household received Food Stamps/SNAP in the past 12 months
														"B22001_001"),
                     areas = area,
                     summary_level = "tract") 
MA.Snap.ch <- MA.Snap.ch[MA.Snap.ch$NAME == "Census Tract 300, Penobscot County, Maine",] %>%
  rename(City = area) %>%
  	mutate(SnapRec = B22001_002/B22001_001) %>%
	dplyr::select(City ,SnapRec) %>%
  left_join(Chelsea.MA, MA.Snap.ch, by="City")

Chelsea.MA <- cbind(MA.Snap.ch)

#######################################################################################################################################

```
* The percent of households, who received Food Stamps/SNAP in the past 12 months in `r Chelsea.MA$City` state is `r (round(Chelsea.MA$SnapRec,4))*100`%.\

\

## Housing
\

### Proportion of Renters to Owners
\

### Home Values
\
```{r echo=FALSE, warning=FALSE, message=FALSE, results='hide', comment=""}

                                                #----------------------#
                                                #       HOUSING        #
                                                #----------------------#

######################   Proportion of Renters to Owners, Home Values   ########################################

MA.Housing.ch <- read_acs5year(year = endyr,
                     states = j,
                     table_contents = c("B25009_001", # Total Occupied Housing Units
															 	"B25009_002", # Total Owner Occupied Housing Units
															 	"B25009_010", # Total Renter Occupied Housing Units
															 	"B25077_001", # Median Home Value
															 	"B25082_001", # Aggregate value
															 	"B25003_002"), # Owner occupied
                     areas = area,
                     summary_level = "tract") 
MA.Housing.ch <- MA.Housing.ch[MA.Housing.ch$NAME == "Census Tract 300, Penobscot County, Maine",] %>%
  rename(City = area,
         MedianHValue = B25077_001,
	       AvgHValue = B25082_001/B25003_002) %>%
	mutate(RentersPer = B25009_010/B25009_001,
				 HomeOwnPer = B25009_002/B25009_001,
				 RentersToOwners = B25009_010/B25009_002) %>%
	dplyr::select(City,RentersPer, HomeOwnPer, RentersToOwners, MedianHValue, AvgHValue) %>%
  left_join(Chelsea.MA, MA.Housing.ch, by="City")

Chelsea.MA <- cbind(MA.Housing.ch)
```
* The median house value in `r Chelsea.MA$City` is 63900 Dollar.\

* The average house value in `r Chelsea.MA$City` is 106860900 Dollar.\

* The proportion of renters to owners in `r Chelsea.MA$City` is `r round(Chelsea.MA$RentersToOwners,3)`.\

* The percentage of renters in `r Chelsea.MA$City` is `r (round(Chelsea.MA$RentersPer,2))*100`%.\

* The percentage of owners in `r Chelsea.MA$City` is `r (round(Chelsea.MA$HomeOwnPer,2))*100`%.\

Followed is the map of Proportion of Renters to Owners, Median Home Values, and Average Home Values of the city.\

```{r echo=FALSE, warning=FALSE, message=FALSE, comment="", results='hide'}
## MAP
ch.MA.housing <- read_acs5year(year = endyr,
                     states = j,
                     table_contents = c("B25009_001", # Total Occupied Housing Units
															 	"B25009_002", # Total Owner Occupied Housing Units
															 	"B25009_010", # Total Renter Occupied Housing Units
															 	"B25077_001", # Median Home Value
															 	"B25082_001", # Aggregate value
															 	"B25003_002"),
                     areas = area,
                     summary_level = "block group")

ch.housing.get <- tidycensus::get_acs(geography = "block group",
													 survey="acs5",
													 state=j,
													 county = c,
													 year = endyr,
													 output = "wide",
													 variables = c("B25009_001", # Total Occupied Housing Units
															 	"B25009_002", # Total Owner Occupied Housing Units
															 	"B25009_010", # Total Renter Occupied Housing Units
															 	"B25077_001", # Median Home Value
															 	"B25082_001", # Aggregate value
															 	"B25003_002"), 
													 keep_geo_vars = T, geometry = TRUE) %>%
rename(MedianHValue_BG = B25077_001E,
	       AvgHValue_BG = B25082_001E / B25003_002E) %>%
	mutate(RentersPer_BG = B25009_010E/B25009_001E,
				 HomeOwnPer_BG = B25009_002E/B25009_001E,
				 RentersToOwners_BG = B25009_010E/B25009_002E)

chelsea.housing <- merge(ch.housing.get, ch.MA.housing, by.x = 'NAME.y', by.y='NAME')


tm_shape(chelsea.housing[c(2:5),]) + tm_polygons("MedianHValue_BG" , palette="Blues", style = "quantile", title = "Median Home Value", popup.format = list(fun = make_dollar), legend.format = list(fun = make_dollar)) + tm_view(alpha = 0.85,basemaps = "Stamen.TonerLite") + tm_borders(col = "darkgray") + tm_layout(legend.position = c("left","bottom"))

```
\


```{r echo=FALSE, warning=FALSE, message=FALSE, results='hide', comment=""}
tm_shape(chelsea.housing[c(2:5),]) + tm_polygons("AvgHValue_BG" , palette="Greens", style = "quantile", title = "Average Home Value", popup.format = list(fun = make_dollar), legend.format = list(fun = make_dollar)) + tm_view(alpha = 0.85,basemaps = "Stamen.TonerLite") + tm_borders(col = "darkgray")+tm_layout(legend.position = c("left","bottom"))
```


\


\

```{r echo=FALSE, warning=FALSE, message=FALSE, results='hide', comment=""}
tm_shape(chelsea.housing[c(2:5),]) + tm_polygons("RentersToOwners_BG" , palette="Reds", style = "quantile", title = "RentersToOwners") + tm_view(alpha = 0.85,basemaps = "Stamen.TonerLite") + tm_borders(col = "darkgray") + tm_layout(legend.position = c("left","bottom"))
```

\

### Property Tax Rates
\



```{r echo=FALSE, warning=FALSE, message=FALSE, results='hide', comment=""}
Chelsea.MA$PropertyTaxRate <- 33.00
```
* The 2020 tax rate is 33.00 Dollar per 1000 Dollar of assessed valuation in `r Chelsea.MA$City`.(https://millinocket.org/government/assessor/)


\



\


### Duration of Occupancy
\

```{r echo=FALSE, warning=FALSE, message=FALSE, results='hide', comment=""}
#########################################      Duration of Occupancy    ####################################################
MA.Dur.Occ.ch <- read_acs5year(year = endyr,
                     states = j,
                     table_contents = c("B25039_001", # Median Year Householder Moved Into Unit by Tenure
															 	"B25039_002", # Median Year Householder Moved Into Unit by Tenure (Owner)
															 	"B25039_003"),
                     areas = area,
                     summary_level = "tract") 
MA.Dur.Occ.ch <- MA.Dur.Occ.ch[MA.Dur.Occ.ch$NAME == "Census Tract 300, Penobscot County, Maine",] %>%
  rename(City = area,
         MedYrMovedIn = B25039_001, 
				 MedYrRentMovedIn = B25039_003,
				 MedYrOwnMovedIn = B25039_002) %>%
	dplyr::select(City, MedYrMovedIn, MedYrRentMovedIn, MedYrOwnMovedIn) %>%
  left_join(Chelsea.MA, MA.Dur.Occ.ch, by="City")

Chelsea.MA <- cbind(MA.Dur.Occ.ch)
```

* The median year that a homeowner moved into their current domicile in `r Chelsea.MA$City` is `r Chelsea.MA$MedYrOwnMovedIn`.\

* The median year that a renter moved into their current domicile in `r Chelsea.MA$City` is `r Chelsea.MA$MedYrRentMovedIn`.\

Followed is the map of Duration of Occupancy of the city.\

```{r echo=FALSE, warning=FALSE, message=FALSE, comment="", results='hide'}
## MAP
ch.MA.dur <- read_acs5year(year = endyr,
                     states = j,
                     table_contents = c("B25039_001", # Median Year Householder Moved Into Unit by Tenure
															 	"B25039_002", # Median Year Householder Moved Into Unit by Tenure (Owner)
															 	"B25039_003"),
                     areas = area,
                     summary_level = "block group")

ch.dur.get <- tidycensus::get_acs(geography = "block group",
													 survey="acs5",
													 state=j,
													 county = c,
													 year = endyr,
													 output = "wide",
													 variables = c("B25039_001", # Median Year Householder Moved Into Unit by Tenure
															 	"B25039_002", # Median Year Householder Moved Into Unit by Tenure (Owner)
															 	"B25039_003"), 
													 keep_geo_vars = T, geometry = TRUE) %>%
rename(MedYrMovedIn_BG = B25039_001E, # I'm not sure if I should substract this value from 2018
				 MedYrRentMovedIn_BG = B25039_003E,
				 MedYrOwnMovedIn_BG = B25039_002E)

chelsea.dur <- merge(ch.dur.get, ch.MA.dur, by.x = 'NAME.y', by.y='NAME')



tm_shape(chelsea.dur[c(2:5),]) + tm_polygons(c("MedYrMovedIn_BG","MedYrOwnMovedIn_BG"), style="quantile", palette=list("Greens","seq"), border.col = "darkgray", title=c("MedYrMovedIn","MedYrOwnMovedIn")) + tm_view(alpha = 0.85, basemaps = "Stamen.TonerLite") + tm_layout(legend.position = c("left","bottom"))


```
\

```{r echo=FALSE, warning=FALSE, message=FALSE, results='hide', comment=""}
tm_shape(chelsea.dur[c(2:5),]) + tm_polygons("MedYrRentMovedIn_BG", style="fixed", breaks=c(2012,2013,2014), palette="Blues", border.col = "darkgray", title="MedYrRentMovedIn") + tm_view(alpha = 0.85, basemaps = "Stamen.TonerLite") + tm_layout(legend.position = c("left","bottom"))
```


\

### Cost Burdened Occupants


\

```{r echo=FALSE, warning=FALSE, message=FALSE, results='hide', comment=""}

##############################################   Cost Burdened Occupants   ##################################

MA.CostBurdened.ch <- read_acs5year(year = endyr,
                     states = j,
                     table_contents = c("B25095_006", # Household income by selected monthly owner Costs as a percentage of household income
														              # Less than $10,000 (30 to 34.9 percent)
														              "B25095_007", # Less than $10,000 (35 to 39.9 percent)
														              "B25095_008", # Less than $10,000 (40 to 49.9 percent)
														              "B25095_009", # Less than $10,000 (50 percent or more)
														              "B25095_015", # $10,000 to $19,999 (30 to 34.9 percent)
														              "B25095_016", # $10,000 to $19,999 (35 to 39.9 percent)
														              "B25095_017", # $10,000 to $19,999 (40 to 49.9 percent)
														              "B25095_018", # $10,000 to $19,999 (50 percent or more)
														              "B25095_024", # $20,000 to $34,999 (30 to 34.9 percent)
														              "B25095_025", # $20,000 to $34,999 (35 to 39.9 percent)
														              "B25095_026", # $20,000 to $34,999 (40 to 49.9 percent)
														              "B25095_027", # $20,000 to $34,999 (50 percent or more)
														              "B25095_033", # $35,000 to $49,999 (30 to 34.9 percent)
														              "B25095_034", # $35,000 to $49,999 (35 to 39.9 percent)
														              "B25095_035", # $35,000 to $49,999 (40 to 49.9 percent)
														              "B25095_036", # $35,000 to $49,999 (50 percent or more)
														              "B25095_042", # $50,000 to $74,999 (30 to 34.9 percent)
														              "B25095_043", # $50,000 to $74,999 (35 to 39.9 percent)
														              "B25095_044", # $50,000 to $74,999 (40 to 49.9 percent)
														              "B25095_045", # $50,000 to $74,999 (50 percent or more)
														              "B25095_051", # $75,000 to $99,999 (30 to 34.9 percent)
														              "B25095_052", # $75,000 to $99,999 (35 to 39.9 percent)
														              "B25095_053", # $75,000 to $99,999 (40 to 49.9 percent)
														              "B25095_054", # $75,000 to $99,999 (50 percent or more)
														              "B25095_060", # $100,000 to $149,999 (30 to 34.9 percent)
														              "B25095_061", # $100,000 to $149,999 (35 to 39.9 percent)
														              "B25095_062", # $100,000 to $149,999 (40 to 49.9 percent)
														              "B25095_063", # $100,000 to $149,999 (50 percent or more)
														              "B25095_069", # $150,000 or more (30 to 34.9 percent)
														              "B25095_070", # $150,000 or more (35 to 39.9 percent)
														              "B25095_071", # $150,000 or more (40 to 49.9 percent)
														              "B25095_072", # $150,000 or more (50 percent or more)
														              "B25095_001",
														              "B25095_001"), # Total
                     areas = area,
                     summary_level = "tract")
MA.CostBurdened.ch <- MA.CostBurdened.ch[MA.CostBurdened.ch$NAME == "Census Tract 300, Penobscot County, Maine",] %>%
  rename(City = area) %>%
  mutate(CostBurdened = B25095_006+B25095_007+B25095_008+B25095_009+B25095_015+B25095_016+B25095_017+B25095_018+
	       B25095_024+B25095_025+B25095_026+B25095_027+B25095_033+B25095_034+B25095_035+B25095_036+B25095_042+B25095_043+B25095_044+B25095_045+B25095_051+B25095_052+B25095_053+B25095_054+B25095_060+B25095_061+B25095_062+B25095_063+B25095_069+B25095_070+B25095_071+B25095_072,
         CostBurdenedPer = CostBurdened/B25095_001) %>%
	dplyr::select(City ,CostBurdened,CostBurdenedPer) %>%
  left_join(Chelsea.MA, MA.CostBurdened.ch, by="City")

Chelsea.MA <- cbind(MA.CostBurdened.ch)
```
* The number of households, who spend 30 percent or more on housing costs in `r Chelsea.MA$City` is `r Chelsea.MA$CostBurdened` which is `r (round(Chelsea.MA$CostBurdenedPer,3))*100`% of total number of households, who spend money on housing costs.\




\

### Local/State Policies + Assistance on Rent Deferment
\

* There is a program in Millinocket city, whose name is General Assistance program. This program is intended to help people who are in need and have no other resources available to them. General Assistance program provides assistance for basic necessities such as rent, food, personal and household supplies, medication, heating fuel, utilities and other essential services.(https://millinocket.org/)


```{r}
millinocket <- Chelsea.MA
millinocket
```








