---
title: "Municipal Indicators For Providence, RI"
output:
  pdf_document:
    toc: yes
    toc_depth: '3'
  html_notebook:
    toc: yes
    toc_depth: 3
  html_document:
    df_print: paged
    toc: yes
    toc_depth: '3'
---

In this report, 40 demographic, educational institutions, employment, healthcare, municipal finance, infrastructure, food insecurity and housing variables are presented using federally-published data sources for the city of Providence, Rhode Island. To better demonstrate their ties to geography of the city, the variables have been mapped on the block group level. \


## Demographic
\
```{r echo=FALSE, warning=FALSE, message=FALSE, results='hide', comment=""}
                                 #######################################################
                                 #       Municipal Indicators For Municipalities        #               
                                 #######################################################
library(stringr)
library(tidycensus)
library(tidyverse)
library(tigris)
library(totalcensus)

options(tigris_class = "sf",tigris_use_cache = T)
# Get API key 
# census_api_key("045944708d514a404443c9f2d227ac8bf6a8f8bd", install = TRUE, overwrite=TRUE)
# Specify the end year 
# This is for 5-year estimates
endyr <- 2018
# List of ACS variables
Var_List <- tidycensus::load_variables(year = endyr, dataset = "acs5", cache = T)
#View(Var_List)

```


\


### Population  Size
\

```{r echo=FALSE, warning=FALSE, message=FALSE, results='hide', comment=""}

                                             #--------------------#
                                             #    DEMOGRAPHIC     #
                                             #--------------------#

##############################################    Population  Size    ###########################################################
library(tidyverse)
# search_cbsa()
# search_fips()
# set_path_to_census("C:\\Users\\ux305\\Desktop\\census")
j <- "RI"
area <- "Providence city, RI"
c <- "Providence"
ch.pop <- read_acs5year(year = endyr,
                     states = j,
                     table_contents = "B01003_001",
                     areas = area,
                     summary_level = "place") %>%
				 rename(TotalPop = population,
				        City = NAME) %>% dplyr::select(City, TotalPop)

Chelsea.MA <- cbind(ch.pop)

ch.MA.Pop <- read_acs5year(year = endyr,
                     states = j,
                     table_contents = "B01003_001",
                     areas = area,
                     summary_level = "block group")

ch.Pop.get <- tidycensus::get_acs(geography = "block group",
													 survey="acs5",
													 state=j,
													 county=c,
													 year = endyr,
													 output = "wide",
													 variables = "B01003_001", # Total Population
													 keep_geo_vars = T, geometry = TRUE)

Chelsea <- merge(ch.Pop.get, ch.MA.Pop, by.x = 'NAME.y', by.y='NAME')
Chelsea.MA
```
The total population of `r Chelsea.MA$City` is `r Chelsea.MA$TotalPop`. Followed is the population size map of the city.\
\

```{r , echo=FALSE, warning=FALSE, message=FALSE, comment=""}

## MAP
library(tmap)
tm_shape(Chelsea) + tm_polygons('B01003_001E' , palette="seq", style = "jenks", title = "Total Population")  + tm_view(alpha = 0.85,view.legend.position = c("right", "bottom"),basemaps = "Stamen.TonerLite") +tm_borders(col = "darkgray")
```
\


### Population Age
\

```{r echo=FALSE, warning=FALSE, message=FALSE, results='hide', comment=""}
##############################################    Population  Age   #############################################################
MA.Age.ch <- read_acs5year(year = endyr,
                     states = j,
                     table_contents = c("B01001_003","B01001_004","B01001_005","B01001_006",
													 							"B01001_027","B01001_028","B01001_029","B01001_030", # Under 5 years; Between 5 to 17
													 							"B01001_007","B01001_008","B01001_009",
													 							"B01001_010","B01001_011","B01001_012",
													 							"B01001_036","B01001_035","B01001_034",
													 							"B01001_033","B01001_032","B01001_031", # 18 to 24 years; 25 to 34 years
													 							"B01001_013","B01001_014","B01001_015",
													 							"B01001_016","B01001_017","B01001_018","B01001_019",
													 							"B01001_043","B01001_042","B01001_041","B01001_040",
													 							"B01001_039","B01001_038","B01001_037", # 35 to 44 years; 45 to 54 years; 55 to 59 years; 60 to 64 years
													 							"B01001_020","B01001_021","B01001_022",
													 							"B01001_023","B01001_024","B01001_025",
													 							"B01001_044","B01001_045","B01001_046",
													 							"B01001_047","B01001_048","B01001_049","B01003_001"),
                     areas = area,
                     summary_level = "place") %>%
  rename(City = NAME) %>%
				 mutate(AgeU18 = B01001_003+B01001_004+B01001_005+B01001_006+B01001_027+B01001_028+B01001_029+B01001_030, 
				 Age1834 = B01001_007+B01001_008+B01001_009+B01001_010+B01001_011+B01001_012+B01001_036+B01001_035+B01001_034+B01001_033+B01001_032+B01001_031,
				 Age3564 = B01001_013+B01001_014+B01001_015+B01001_016+B01001_017+B01001_018+B01001_019+B01001_043+B01001_042+B01001_041+B01001_040+B01001_039+B01001_038+B01001_037,
				 AgeO65 = B01001_020+B01001_021+B01001_022+B01001_023+B01001_024+B01001_025+B01001_044+B01001_045+B01001_046+B01001_047+B01001_048+B01001_049) %>%
  	mutate(AgeU18Per = AgeU18/population,
				 Age1834Per = Age1834/population,
				 Age3564Per = Age3564/population,
				 AgeO65Per = AgeO65/population) %>%
dplyr::select(AgeU18,Age1834,Age3564,AgeO65,AgeU18Per,Age1834Per,Age3564Per,AgeO65Per, City) %>%
  left_join(Chelsea.MA, MA.Age.ch, by="City")

Chelsea.MA <- cbind(MA.Age.ch)
```
* The number of people, who are younger than 18 Years old is `r Chelsea.MA$AgeU18` in `r Chelsea.MA$City` 
, which is `r (round(Chelsea.MA$AgeU18Per,4))*100`% of total population in `r Chelsea.MA$City`. \
* The number of people, who are between 18 and 34 years old is `r Chelsea.MA$Age1834` in `r Chelsea.MA$City`, which is `r (round(Chelsea.MA$Age1834Per,4))*100`% of total population in `r Chelsea.MA$City`.\
* The number of people, who are between 35 and 64 years old is `r Chelsea.MA$Age3564` in `r Chelsea.MA$City`, which is `r (round(Chelsea.MA$Age3564Per,4))*100`% of total population in `r Chelsea.MA$City`.\
* The number of people, who are older than 65 years old is `r Chelsea.MA$AgeO65` in `r Chelsea.MA$City`, which is `r (round(Chelsea.MA$AgeO65Per,4))*100`% of total population in `r Chelsea.MA$City`.\

Followed is the population age map of the city.\

```{r echo=FALSE, warning=FALSE, message=FALSE, comment="", results='hide'}
## Map

ch.MA.Age <- read_acs5year(year = endyr,
                     states = j,
                     table_contents = c("B01001_003","B01001_004","B01001_005","B01001_006",
													 							"B01001_027","B01001_028","B01001_029","B01001_030", # Under 5 years; Between 5 to 17
													 							"B01001_007","B01001_008","B01001_009",
													 							"B01001_010","B01001_011","B01001_012",
													 							"B01001_036","B01001_035","B01001_034",
													 							"B01001_033","B01001_032","B01001_031", # 18 to 24 years; 25 to 34 years
													 							"B01001_013","B01001_014","B01001_015",
													 							"B01001_016","B01001_017","B01001_018","B01001_019",
													 							"B01001_043","B01001_042","B01001_041","B01001_040",
													 							"B01001_039","B01001_038","B01001_037", # 35 to 44 years; 45 to 54 years; 55 to 59 years; 60 to 64 years
													 							"B01001_020","B01001_021","B01001_022",
													 							"B01001_023","B01001_024","B01001_025",
													 							"B01001_044","B01001_045","B01001_046",
													 							"B01001_047","B01001_048","B01001_049","B01003_001"),
                     areas = area,
                     summary_level = "block group")

ch.age.get <- tidycensus::get_acs(geography = "block group",
													 survey="acs5",
													 state=j,
													 county = c,
													 year = endyr,
													 output = "wide",
													 variables = c("B01001_003","B01001_004","B01001_005","B01001_006",
													 							"B01001_027","B01001_028","B01001_029","B01001_030", # Under 5 years; Between 5 to 17
													 							"B01001_007","B01001_008","B01001_009",
													 							"B01001_010","B01001_011","B01001_012",
													 							"B01001_036","B01001_035","B01001_034",
													 							"B01001_033","B01001_032","B01001_031", # 18 to 24 years; 25 to 34 years
													 							"B01001_013","B01001_014","B01001_015",
													 							"B01001_016","B01001_017","B01001_018","B01001_019",
													 							"B01001_043","B01001_042","B01001_041","B01001_040",
													 							"B01001_039","B01001_038","B01001_037", # 35 to 44 years; 45 to 54 years; 55 to 59 years; 60 to 64 years
													 							"B01001_020","B01001_021","B01001_022",
													 							"B01001_023","B01001_024","B01001_025",
													 							"B01001_044","B01001_045","B01001_046",
													 							"B01001_047","B01001_048","B01001_049","B01003_001"), # Total Population
													 keep_geo_vars = T, geometry = TRUE) %>%
mutate(AgeU18 = B01001_003E+B01001_004E+B01001_005E+B01001_006E+B01001_027E+B01001_028E+B01001_029E+B01001_030E, 
				 Age1834 = B01001_007E+B01001_008E+B01001_009E+B01001_010E+B01001_011E+B01001_012E+B01001_036E+B01001_035E+B01001_034E+B01001_033E+B01001_032E+B01001_031E,
				 Age3564 = B01001_013E+B01001_014E+B01001_015E+B01001_016E+B01001_017E+B01001_018E+B01001_019E+B01001_043E+B01001_042E+B01001_041E+B01001_040E+B01001_039E+B01001_038E+B01001_037E,
				 AgeO65 = B01001_020E+B01001_021E+B01001_022E+B01001_023E+B01001_024E+B01001_025E+B01001_044E+B01001_045E+B01001_046E+B01001_047E+B01001_048E+B01001_049E)  %>%
  rename(AgeU18_BG = AgeU18,
				 Age1834_BG = Age1834,
				 Age3564_BG = Age3564,
				 AgeO65_BG = AgeO65) %>%
	mutate(AgeU18Per_BG = ((AgeU18_BG/B01003_001E)*100),
				 Age1834Per_BG = ((Age1834_BG/B01003_001E)*100),
				 Age3564Per_BG = ((Age3564_BG/B01003_001E)*100),
				 AgeO65Per_BG = ((AgeO65_BG/B01003_001E)*100))  

Chelsea.age <- merge(ch.age.get, ch.MA.Age, by.x = 'NAME.y', by.y='NAME')

make_percent <- function(x, digits = 2) {
  paste0("%", formatC(x, digits = digits, format = "f", big.mark = ","))
}
 

tm_shape(Chelsea.age) + tm_polygons(c('AgeU18Per_BG','Age1834Per_BG',"Age3564Per_BG","AgeO65Per_BG") , palette=list("Oranges", "Greens", "Blues", "Purples") ,style = "quantile", title = c('AgeU18','Age1834',"Age3564","AgeO65"), popup.format = list(fun = make_percent), legend.format = list(fun = make_percent))  + tm_view(alpha = 0.85,basemaps = "Stamen.TonerLite") +tm_borders(col = "darkgray") + tmap_options(show.messages = FALSE) +tm_layout(legend.position = c("left","bottom"))

```
\

### Median Household Income
\


```{r  echo=FALSE, warning=FALSE, message=FALSE, results='hide', comment=""}
#################################################    Median Household Income    ##############################################################

MA.MHI.ch <- read_acs5year(year = endyr,
                     states = j,
                     table_contents = "B19013_001",
                     areas = area,
                     summary_level = "place") %>%
				 rename(MedHouseIncome = B19013_001,
				        City = NAME) %>% 
	dplyr::select(City, MedHouseIncome) %>%
  left_join(Chelsea.MA, MA.MHI.ch, by="City")
Chelsea.MA <- cbind(MA.MHI.ch)
```
* The median household income in the past 12 months is 42158 Dollar in `r Chelsea.MA$City`.\

Followed is the Median Household Income map of the city.\

```{r echo=FALSE, warning=FALSE, message=FALSE, comment="", results='hide'}
## MAP
ch.MA.income <- read_acs5year(year = endyr,
                     states = j,
                     table_contents = "B19013_001",
                     areas = area,
                     summary_level = "block group")

ch.income.get <- tidycensus::get_acs(geography = "block group",
													 survey="acs5",
													 state=j,
													 county = c,
													 year = endyr,
													 output = "wide",
													 variables = "B19013_001", 
													 keep_geo_vars = T, geometry = TRUE) 

chelsea.income <- merge(ch.income.get, ch.MA.income, by.x = 'NAME.y', by.y='NAME')

make_dollar <- function(x, digits = 0) {
  paste0("$", formatC(x, digits = digits, format = "f", big.mark = ","))
}

library(tmap)
tm_shape(chelsea.income) + tm_polygons('B19013_001E' , palette="Blues", style="jenks", title = "Median Household Income", popup.format = list(fun = make_dollar), legend.format = list(fun = make_dollar))  + tm_view(alpha = 0.85,view.legend.position = c("right", "bottom"),basemaps = "Stamen.TonerLite") +tm_borders(col = "darkgray") + tmap_options(show.messages = FALSE) 
```

\

### Racial Breakdown 
\

```{r echo=FALSE, warning=FALSE, message=FALSE, results='hide', comment=""}
##############################################      Racial Breakdown     ###################################################################
MA.Race.ch <- read_acs5year(year = endyr,
                     states = j,
                     table_contents = c("B03002_003", # white alone
																					"B03002_004", # black or African American alone
																					"B03002_006", # Asian alone
																					"B03002_012", # Hispanic or Latino alone
																					"B03002_009"),
                     areas = area,
                     summary_level = "place") %>%
				 rename(White = B03002_003,
				 Black = B03002_004,
				 Asian = B03002_006,
				 Hispanic = B03002_012,
				 TwoOrMore = B03002_009,
				 City = NAME) %>% 
  mutate(WhitePer = (White/population),
         BlackPer = (Black/population),
         AsianPer = (Asian/population),
         HispanicPer = (Hispanic/population),
         TwoOrMorePer = (TwoOrMore/population)) %>%
	dplyr::select(City, White, Black, Asian, Hispanic, TwoOrMore,
	       WhitePer, BlackPer, AsianPer, HispanicPer, TwoOrMorePer) %>%
  left_join(Chelsea.MA, MA.Race.ch, by="City")
Chelsea.MA <- cbind(MA.Race.ch)

```
* The number of people who are white in `r Chelsea.MA$City` is `r Chelsea.MA$White` which is `r (round(Chelsea.MA$WhitePer,4))*100`% of total population in `r Chelsea.MA$City`.\
* The number of people who are black in `r Chelsea.MA$City` is `r Chelsea.MA$Black` which is `r (round(Chelsea.MA$BlackPer,4))*100`% of total population in `r Chelsea.MA$City`.\
* The number of people who are asian in `r Chelsea.MA$City` is `r Chelsea.MA$Asian` which is `r (round(Chelsea.MA$AsianPer,4))*100`% of total population in `r Chelsea.MA$City`.\
* The number of people who are hispanic in `r Chelsea.MA$City` is `r Chelsea.MA$Hispanic` which is `r (round(Chelsea.MA$HispanicPer,4))*100`% of total population in `r Chelsea.MA$City`.\
* The number of people who are have two or more races in `r Chelsea.MA$City` is `r Chelsea.MA$TwoOrMore` which is `r (round(Chelsea.MA$TwoOrMorePer,4))*100`% of total population in `r Chelsea.MA$City`. \

Followed is the Racial Breakdown map of the city.\

```{r echo=FALSE, warning=FALSE, message=FALSE, comment="",results='hide'}
## Map
ch.MA.race <- read_acs5year(year = endyr,
                     states = j,
                     table_contents = c("B03002_003", # white alone
																					"B03002_004", # black or African American alone
																					"B03002_006", # Asian alone
																					"B03002_012", # Hispanic or Latino alone
																					"B03002_009",
																					"B01003_001"),
                     areas = area,
                     summary_level = "block group")

ch.race.get <- tidycensus::get_acs(geography = "block group",
													 survey="acs5",
													 state=j,
													 county = c,
													 year = endyr,
													 output = "wide",
													 variables = c("B03002_003", # white alone
																					"B03002_004", # black or African American alone
																					"B03002_006", # Asian alone
																					"B03002_012", # Hispanic or Latino alone
																					"B03002_009",
																					"B01003_001"), 
													 keep_geo_vars = T, geometry = TRUE) %>%
rename(White_BG = B03002_003E,
				 Black_BG = B03002_004E,
				 Asian_BG = B03002_006E,
				 Hispanic_BG = B03002_012E,
				 TwoOrMore_BG = B03002_009E) %>%
  mutate(WhitePer_BG = (White_BG/B01003_001E)*100,
         BlackPer_BG = (Black_BG/B01003_001E)*100,
         AsianPer_BG = (Asian_BG/B01003_001E)*100,
         HispanicPer_BG = (Hispanic_BG/B01003_001E)*100,
         TwoOrMorePer_BG = (TwoOrMore_BG/B01003_001E)*100) 

chelsea.race <- merge(ch.race.get, ch.MA.race, by.x = 'NAME.y', by.y='NAME')


make_percent <- function(x, digits = 2) {
  paste0("%", formatC(x, digits = digits, format = "f", big.mark = ","))
}
 
library(tmap)
tm_shape(chelsea.race) + tm_polygons(c("WhitePer_BG", "BlackPer_BG", "AsianPer_BG", "HispanicPer_BG", "TwoOrMorePer_BG") , palette=list("Oranges", "Greens", "Blues", "Purples", "Reds"), style = "quantile", title = c("White", "Black", "Asian","Hispanic", "TwoOrMore"), popup.format = list(fun = make_percent), legend.format = list(fun = make_percent))  + tm_view(alpha = 0.85,view,basemaps = "Stamen.TonerLite") +tm_borders(col = "darkgray") + tmap_options(show.messages = FALSE) + tm_layout(legend.position = c("left","bottom"))
```

\


### Gender Breakdown
\

```{r echo=FALSE, warning=FALSE, message=FALSE, results='hide', comment=""}
#####################################################    Gender Breakdown    #############################################################

MA.Sex.ch <- read_acs5year(year = endyr,
                     states = j,
                     table_contents = c("B01001_026",
													               "B01001_002"),
                     areas = area,
                     summary_level = "place") %>%
  mutate(SexRatio = B01001_026/B01001_002) %>%
  rename(Female = B01001_026,
				 Male = B01001_002,
				 City = NAME) %>%
  mutate(FemalePer = Female/population,
				 MalePer = Male/population) %>% 
	dplyr::select(City, SexRatio, Female, Male, FemalePer, MalePer) %>%
  left_join(Chelsea.MA, MA.Sex.ch, by="City")

Chelsea.MA <- cbind(MA.Sex.ch)

```
* The number of females in `r Chelsea.MA$City` is `r Chelsea.MA$Female` which is `r 100*(round(Chelsea.MA$FemalePer,2))`% of total population in `r Chelsea.MA$City`.\
* The number of males in `r Chelsea.MA$City` is `r Chelsea.MA$Male` which is `r 100*(round(Chelsea.MA$MalePer,2))`% of total population in `r Chelsea.MA$City`.\
* The ratio of number of females to males in `r Chelsea.MA$City` is `r round(Chelsea.MA$SexRatio,2)`.\

Followed is the Gender Breakdown map of the city.\

```{r echo=FALSE, warning=FALSE, message=FALSE, comment="",results='hide'}
## MAP
ch.MA.sex <- read_acs5year(year = endyr,
                     states = j,
                     table_contents = c("B01001_026",
													               "B01001_002",
													               "B01003_001"),
                     areas = area,
                     summary_level = "block group")

ch.sex.get <- tidycensus::get_acs(geography = "block group",
													 survey="acs5",
													 state=j,
													 county = c,
													 year = endyr,
													 output = "wide",
													 variables = c("B01001_026",
													               "B01001_002",
													               "B01003_001"), 
													 keep_geo_vars = T, geometry = TRUE)  %>%
mutate(SexRatio = B01001_026E/B01001_002E) %>%
  rename(Female = B01001_026E,
				 Male = B01001_002E) %>%
  mutate(FemalePer = (Female/B01003_001E)*100,
         MalePer = (Male/B01003_001E)*100)

chelsea.sex <- merge(ch.sex.get, ch.MA.sex, by.x = 'NAME.y', by.y='NAME')

library(tmap)
tm_shape(chelsea.sex) + tm_polygons(c('Female', 'Male'), palette=list("Purples", "Reds"), title = c("Female","Male"))  + tm_view(alpha = 0.85,basemaps = "Stamen.TonerLite") +tm_borders(col = "darkgray")+ tmap_options(show.messages = FALSE) + tm_layout(legend.position = c("left","bottom"))

```

\



```{r echo=FALSE, warning=FALSE, message=FALSE, results='hide', comment=""}

tm_shape(chelsea.sex) + tm_polygons(c('FemalePer', 'MalePer'), palette=list("Blues", "Oranges"), title = c("Female Percentage","Male Percentage"), popup.format = list(fun = make_percent), legend.format = list(fun = make_percent)) + tm_view(alpha = 0.85,basemaps = "Stamen.TonerLite") +tm_borders(col = "darkgray")+ tmap_options(show.messages = FALSE) + tm_layout(legend.position = c("left","bottom"))

```




\

### Average Household Size
\
```{r echo=FALSE, warning=FALSE, message=FALSE, results='hide', comment=""}
#################################################    Average Household Size    ############################################################

MA.AveHH.Size.ch <- read_acs5year(year = endyr,
                     states = j,
                     table_contents = c("B11002_001", 
															         "B19001_001"),
                     areas = area,
                     summary_level = "place") %>%
  rename(City = NAME) %>% 
  mutate(AveHouseH = B11002_001/B19001_001) %>%
	dplyr::select(City, AveHouseH) %>%
  left_join(Chelsea.MA, MA.AveHH.Size.ch, by="City")

Chelsea.MA <- cbind(MA.AveHH.Size.ch)
```
* The average household size is `r round(Chelsea.MA$AveHouseH, 2)` in `r Chelsea.MA$City`.\

Followed is the Average Household Size map of the city.\

```{r echo=FALSE, warning=FALSE, message=FALSE, comment="",results='hide'}
## MAP

ch.MA.HH.Size <- read_acs5year(year = endyr,
                     states = j,
                     table_contents = c("B11002_001", # Household type by relatives and nonrelatives for population in households
															              "B19001_001"),
                     areas = area,
                     summary_level = "block group")

ch.HH.Size.get <- tidycensus::get_acs(geography = "block group",
													 survey="acs5",
													 state=j,
													 county = c,
													 year = endyr,
													 output = "wide",
													 variables = c("B11002_001", # Household type by relatives and nonrelatives for population in households
															              "B19001_001"), 
													 keep_geo_vars = T, geometry = TRUE)  %>%
  mutate(AveHouseH = B11002_001E/B19001_001E) 

chelsea.HH.Size <- merge(ch.HH.Size.get, ch.MA.HH.Size, by.x = 'NAME.y', by.y='NAME')


tm_shape(chelsea.HH.Size) + tm_polygons("AveHouseH", style = "quantile", palette="Purples", border.col = "darkgray", title="Ave Household Size") + tm_view(alpha = 0.85,view.legend.position = c("right", "bottom"), basemaps = "Stamen.TonerLite") + tmap_options(show.messages = FALSE)

```

\

### Educational Attainment
\
```{r echo=FALSE, warning=FALSE, message=FALSE, results='hide', comment=""}
#############################################   Educational Attainment    ##################################################################
MA.Educ.ch <- read_acs5year(year = endyr,
                     states = j,
                     table_contents = c("B15002_001", # variables for Sex by Edu Attainment for Population 25 yrs or Over
																					"B15002_003","B15002_004","B15002_005",
																					"B15002_006","B15002_007","B15002_008","B15002_009","B15002_010",
																					"B15002_020","B15002_021","B15002_022","B15002_023","B15002_024",
																					"B15002_025","B15002_026","B15002_027",
																					"B15002_011","B15002_028",
																					"B15002_012","B15002_013","B15002_014",
																					"B15002_029","B15002_030","B15002_031",
																					"B15002_015","B15002_032",
																					"B15002_016","B15002_033",
																					"B15002_017","B15002_034",
																					"B15002_018","B15002_035"),
                     areas = area,
                     summary_level = "place") %>%
  rename(City = NAME) %>% 
  mutate(LessThanHS = (B15002_003+B15002_004+B15002_005+B15002_006+B15002_007+B15002_008+B15002_009+B15002_010+B15002_020+B15002_021+B15002_022+B15002_023+B15002_024+B15002_025+B15002_026+B15002_027),
				 HSGrad = (B15002_011 + B15002_028),
				 SomeColl = (B15002_012+ B15002_013 + B15002_014 + B15002_029 + B15002_030 + B15002_031),
				 Bach = (B15002_015 + B15002_032),
				 Master = (B15002_016 + B15002_033),
				 Prof = (B15002_017 + B15002_034),
				 Doc = (B15002_018 + B15002_035),
				 LessThanHSPer = (LessThanHS/population),
         HSGradPer = (HSGrad/population),
         SomeCollPer = (SomeColl/population),
         BachPer = (Bach/population),
         MasterPer = (Master/population),
         ProfPer = (Prof/population),            
         DocPer = (Doc/population)) %>%
	dplyr::select(City, LessThanHS, HSGrad, SomeColl, Bach, Master, Prof, Doc,
	       LessThanHSPer, HSGradPer, SomeCollPer, BachPer, MasterPer, ProfPer, DocPer) %>% 
  left_join(Chelsea.MA, MA.Educ.ch, by="City") 

Chelsea.MA <- cbind(MA.Educ.ch)
```
* The number of people, who have less than high school degree is `r  Chelsea.MA$LessThanHS` in `r Chelsea.MA$City` which is `r (round(Chelsea.MA$LessThanHSPer,4))*100`% of the population who are 25 years old or older in `r Chelsea.MA$City`.\
* The number of people, who have high school degree is `r  Chelsea.MA$HSGrad` in `r Chelsea.MA$City` which is `r (round(Chelsea.MA$HSGradPer,4))*100`% of the population who are 25 years old or older in `r Chelsea.MA$City`.\
* The number of people, who have some college degree is `r  Chelsea.MA$SomeColl` in `r Chelsea.MA$City` which is `r (round(Chelsea.MA$SomeCollPer,4))*100`% of the population who are 25 years old or older in `r Chelsea.MA$City`.\
* The number of people, who have bachelor degree is `r Chelsea.MA$Bach` in `r Chelsea.MA$City` which is `r (round(Chelsea.MA$BachPer,4))*100`% of the population who are 25 years old or older in `r Chelsea.MA$City`.\
* The number of people, who have master's degree is `r Chelsea.MA$Master` in `r Chelsea.MA$City` which is `r (round(Chelsea.MA$MasterPer,4))*100`% of the population who are 25 years old or older in `r Chelsea.MA$City`.\
* The number of people, who have professional school degree is `r  Chelsea.MA$Prof` in `r Chelsea.MA$City` which is `r (round(Chelsea.MA$ProfPer,4))*100`% of the population who are 25 years old or older in `r Chelsea.MA$City`.\
* The number of people, who have doctorate degree is `r  Chelsea.MA$Doc` in `r Chelsea.MA$City` which is `r (round(Chelsea.MA$DocPer,4))*100`% of the population who are 25 years old or older in `r Chelsea.MA$City`.\

Followed is the Educational Attainment map of the city.\

```{r echo=FALSE, warning=FALSE, message=FALSE, comment="",results='hide'}
## MAP
ch.MA.Educ <- read_acs5year(year = endyr,
                     states = j,
                     table_contents = c("B15002_001", # variables for Sex by Edu Attainment for Population 25 yrs or Over
																					"B15002_003","B15002_004","B15002_005",
																					"B15002_006","B15002_007","B15002_008","B15002_009","B15002_010",
																					"B15002_020","B15002_021","B15002_022","B15002_023","B15002_024",
																					"B15002_025","B15002_026","B15002_027",
																					"B15002_011","B15002_028",
																					"B15002_012","B15002_013","B15002_014",
																					"B15002_029","B15002_030","B15002_031",
																					"B15002_015","B15002_032",
																					"B15002_016","B15002_033",
																					"B15002_017","B15002_034",
																					"B15002_018","B15002_035"),
                     areas = area,
                     summary_level = "block group")

ch.Educ.get <- tidycensus::get_acs(geography = "block group",
													 survey="acs5",
													 state=j,
													 county = c,
													 year = endyr,
													 output = "wide",
													 variables = c("B15002_001", # variables for Sex by Edu Attainment for Population 25 yrs or Over
																					"B15002_003","B15002_004","B15002_005",
																					"B15002_006","B15002_007","B15002_008","B15002_009","B15002_010",
																					"B15002_020","B15002_021","B15002_022","B15002_023","B15002_024",
																					"B15002_025","B15002_026","B15002_027",
																					"B15002_011","B15002_028",
																					"B15002_012","B15002_013","B15002_014",
																					"B15002_029","B15002_030","B15002_031",
																					"B15002_015","B15002_032",
																					"B15002_016","B15002_033",
																					"B15002_017","B15002_034",
																					"B15002_018","B15002_035"), 
													 keep_geo_vars = T, geometry = TRUE) %>%
  mutate(LessThanHS_BG = (B15002_003E+B15002_004E+B15002_005E+B15002_006E+B15002_007E+B15002_008E+B15002_009E+B15002_010E+B15002_020E+B15002_021E+B15002_022E+B15002_023E+B15002_024E+B15002_025E+B15002_026E+B15002_027E),
				 HSGrad_BG = (B15002_011E + B15002_028E),
				 SomeColl_BG = (B15002_012E + B15002_013E + B15002_014E + B15002_029E + B15002_030E + B15002_031E),
				 Bach_BG = (B15002_015E + B15002_032E),
				 Master_BG = (B15002_016E + B15002_033E),
				 Prof_BG = (B15002_017E + B15002_034E),
				 Doc_BG = (B15002_018E + B15002_035E),
		    LessThanHSPer_BG = ((B15002_003E+B15002_004E+B15002_005E+B15002_006E+B15002_007E+B15002_008E+B15002_009E+B15002_010E+B15002_020E+B15002_021E+B15002_022E+B15002_023E+B15002_024E+B15002_025E+B15002_026E+B15002_027E)/B15002_001E)*100,
				 HSGradPer_BG = ((B15002_011E + B15002_028E)/B15002_001E)*100,
				 SomeCollPer_BG = ((B15002_012E + B15002_013E + B15002_014E + B15002_029E + B15002_030E + B15002_031E)/B15002_001E)*100,
				 BachPer_BG = ((B15002_015E + B15002_032E)/B15002_001E)*100,
				 MasterPer_BG = ((B15002_016E + B15002_033E)/B15002_001E)*100,
				 ProfPer_BG = ((B15002_017E + B15002_034E)/B15002_001E)*100,
				 DocPer_BG = ((B15002_018E + B15002_035E)/B15002_001E)*100)

chelsea.Educ <- merge(ch.Educ.get, ch.MA.Educ, by.x = 'NAME.y', by.y='NAME')

make_percent <- function(x, digits = 2) {
  paste0("%", formatC(x, digits = digits, format = "f", big.mark = ","))
}

tm_shape(chelsea.Educ) +
    tm_fill(c('LessThanHSPer_BG', 'HSGradPer_BG', 'SomeCollPer_BG', 'BachPer_BG', 'MasterPer_BG', 'ProfPer_BG', 'DocPer_BG'),
        style="quantile", 
        palette=list("Oranges", "Greens", "Blues", "Purples", "Reds","Dark2", "seq"), popup.format = list(fun = make_percent), legend.format = list(fun = make_percent), title=c('LessThanHS', 'HSGrad', 'SomeColl', 'Bach', 'Master', 'Prof', 'Doc')) + tm_view(alpha = 0.85,basemaps = "Stamen.TonerLite") + tm_borders(col = "darkgray") +tmap_options(show.messages = FALSE) + tm_layout(legend.position = c("left","bottom"))

```

\

### Single Parent Households
\
```{r echo=FALSE, warning=FALSE, message=FALSE, results='hide', comment=""}

##########################################      Single Parent Households     ##################################################################
MA.Single.ch <- read_acs5year(year = endyr,
                     states = j,
                     table_contents = c("B11001_006", # Female Hoseholder, No Husband Present
																						"B11001_005", # Male Householder, No Wife Present
																						"B11001_001"),
                     areas = area,
                     summary_level = "place") %>%
  rename(City = NAME) %>% 
  mutate(FemHeadPer = B11001_006/B11001_001,
	       MaleHeadPer = B11001_005/B11001_001) %>%
	dplyr::select(City, FemHeadPer, MaleHeadPer) %>%
  left_join(Chelsea.MA, MA.Single.ch, by="City")

Chelsea.MA <- cbind(MA.Single.ch)

```
* The percentage of single-female-headed is `r (round(Chelsea.MA$FemHeadPer,4))*100`% of total number of family households in `r Chelsea.MA$City`.\
* The percentage of single-male-headed is `r (round(Chelsea.MA$MaleHeadPer,4))*100`% of total number of family households in `r Chelsea.MA$City`.\

Followed is the Single Parent Households map of the city.\

```{r echo=FALSE, warning=FALSE, message=FALSE, comment="",results='hide'}
## MAP
ch.MA.single <- read_acs5year(year = endyr,
                     states = j,
                     table_contents = c("B11001_006", # Female Hoseholder, No Husband Present
																						"B11001_005", # Male Householder, No Wife Present
																						"B11001_001"),
                     areas = area,
                     summary_level = "block group")

ch.single.get <- tidycensus::get_acs(geography = "block group",
													 survey="acs5",
													 state=j,
													 county = c,
													 year = endyr,
													 output = "wide",
													 variables = c("B11001_006", # Female Hoseholder, No Husband Present
																						"B11001_005", # Male Householder, No Wife Present
																						"B11001_001"), 
													 keep_geo_vars = T, geometry = TRUE)  %>%
mutate(FemHeadPer = (B11001_006E/B11001_001E)*100,
	       MaleHeadPer = (B11001_005E/B11001_001E)*100)

chelsea.single <- merge(ch.single.get, ch.MA.single, by.x = 'NAME.y', by.y='NAME')


make_percent <- function(x, digits = 2) {
  paste0("%", formatC(x, digits = digits, format = "f", big.mark = ","))
}


tm_shape(chelsea.single) + tm_polygons(c("FemHeadPer", "MaleHeadPer"), style="quantile", palette=list("Oranges", "Purples"), popup.format = list(fun = make_percent), legend.format = list(fun = make_percent), border.col = "darkgray", title=c("Female Householder", "Male Householder")) + tm_view(alpha = 0.85, basemaps = "Stamen.TonerLite") + tmap_options(show.messages = FALSE) + tm_layout(legend.position = c("left","bottom"))

```


\


### Population in Poverty
\
```{r echo=FALSE, warning=FALSE, message=FALSE, results='hide', comment=""}
###########################################      Population in Poverty    ######################################################################

MA.PopPov.ch <- read_acs5year(year = endyr,
                     states = j,
                     table_contents = c("B17010_002", # Income in the Past 12 Months Below Poverty Level (Family households in poverty)
																					"B17010_001",  # Total Families Below Poverty Level
																					"B17001_002", # Persons below poverty
																					"B17001_001"),
                     areas = area,
                     summary_level = "place") %>%
  rename(City = NAME) %>% 
  	mutate(FamPovPer = B17010_002/B17010_001,
	       PopPovPer = B17001_002/B17001_001,
	       TotalPopPov = B17001_002) %>%
	dplyr::select(City, FamPovPer, PopPovPer, TotalPopPov) %>%
  left_join(Chelsea.MA,MA.PopPov.ch, by="City")

Chelsea.MA <- cbind(MA.PopPov.ch)
```
* The percentage of family, who are in poverty is `r (round(Chelsea.MA$FamPovPer,4))*100` % in `r Chelsea.MA$City`.\
* The percentage of population, who are in poverty is `r (round(Chelsea.MA$PopPovPer,4))*100` % in `r Chelsea.MA$City`.\
* The total number of people, who are in poverty is `r Chelsea.MA$TotalPopPov` in `r Chelsea.MA$City`.\

Followed is the Population in Poverty map of the city.\

```{r echo=FALSE, warning=FALSE, message=FALSE, comment="",results='hide'}
## MAP

ch.MA.pov <- read_acs5year(year = endyr,
                     states = j,
                     table_contents = c("B17010_002", # Income in the Past 12 Months Below Poverty Level (Family households in poverty)
																					"B17010_001",  # Total Families Below Poverty Level
																					"B17001_002", # Persons below poverty
																					"B17001_001"),
                     areas = area,
                     summary_level = "block group")

ch.pov.get <- tidycensus::get_acs(geography = "block group",
													 survey="acs5",
													 state=j,
													 county = c,
													 year = endyr,
													 output = "wide",
													 variables = c("B17010_002", # Income in the Past 12 Months Below Poverty Level (Family households in poverty)
																					"B17010_001",  # Total Families Below Poverty Level
																					"B17001_002", # Persons below poverty
																					"B17001_001"), 
													 keep_geo_vars = T, geometry = TRUE) %>%
	mutate(FamPovPer_BG = (B17010_002E/B17010_001E)*100,
	       PopPovPer_BG = (B17001_002E/B17001_001E)*100,
	       TotalPopPov_BG = B17001_002E) 

chelsea.pov <- merge(ch.pov.get, ch.MA.pov, by.x = 'NAME.y', by.y='NAME')


tm_shape(chelsea.pov) + tm_polygons("FamPovPer_BG", style="quantile", palette="Purples", border.col = "darkgray", title="Family in Poverty", popup.format = list(fun = make_percent), legend.format = list(fun = make_percent))  + tm_view(alpha = 0.85,view.legend.position = c("left", "bottom"), basemaps = "Stamen.TonerLite")


```

\


### Elderly People Living Alone
\
```{r echo=FALSE, warning=FALSE, message=FALSE, results='hide', comment=""}

########################################     Elderly People Living Alone    ##############################################################
MA.old.ch <- read_acs5year(year = endyr,
                     states = j,
                     table_contents = c("B11010_005", # Male householder, Living alone, 65 years and over
																					"B11010_012", # Female householder, Living alone, 65 years and over
																					"B11010_001"),
                     areas = area,
                     summary_level = "place") %>%
  rename(City = NAME) %>% 
  mutate(OldAlonePer = (B11010_012 + B11010_005)/B11010_001) %>%
	dplyr::select(City, OldAlonePer) %>%
  left_join(Chelsea.MA,MA.old.ch, by="City")

Chelsea.MA <- cbind(MA.old.ch) 

## GIVE US NA
```

* The percentage of householders, who are living alone and are older than 65 years old is `r (round(Chelsea.MA$OldAlonePer,4))*100`% in `r Chelsea.MA$City`.\

\

## Educational Institutions
\


\

### Number/Type of Schools
\

* `r Chelsea.MA$City` has 15 public school district, which contains 37,390 students and 76 public schools. Also, `r  Chelsea.MA$City` has 28 private schools with 5,705 students.(https://nces.ed.gov/) 

\


```{r echo=FALSE, warning=FALSE, message=FALSE, results='hide', comment=""}
Chelsea.MA$SchoolsDistrict <- 15
Chelsea.MA$SchoolsDistStudents <- 37390
Chelsea.MA$PubSchools <- 76
Chelsea.MA$PvSchools <- 28
```


\


### Number/Type of Universities
\

* `r Chelsea.MA$City` has 9 university in 02903, 02905, 02908, 02912, 02918, and 02940 ZIP Codes.(https://nces.ed.gov/) 

```{r echo=FALSE, warning=FALSE, message=FALSE, results='hide', comment=""}
Chelsea.MA$NumUni <- 9
```


\

###	Employment Numbers per Institution
\


* The largest School District in `r Chelsea.MA$City` contains 1,393 teachers and 1,484 other staffs. The total number of teachers in all private schools in `r Chelsea.MA$City` is 763.(http://www.usa.com/)


\

### Number of PK-12 Enrollment in Schools
\

```{r echo=FALSE, warning=FALSE, message=FALSE, results='hide', comment=""}
############################################################################################################################

                                             # ----------------------------- #
                                             #    Educational Institutions   #
                                             #-------------------------------#

# Number of PK-12 Enrollment at block group level
MA.Enrol.ch <- read_acs5year(year = endyr,
                     states = j,
                     table_contents = c("B14001_003", # Enrolled in preschool
																					"B14001_004", # Enrolled in kindergarten
																					"B14001_005", # Enrolled in grade 1 to 4
																					"B14001_006", # Enrolled in grade 5 to 8
																					"B14001_007"),
                     areas = area,
                     summary_level = "place") %>%
  rename(City = NAME) %>% 
  mutate(EnrolPK12 = B14001_003 + B14001_004 + B14001_005 + B14001_006 + B14001_007) %>%
	dplyr::select(City, EnrolPK12) %>%
  left_join(Chelsea.MA, MA.Enrol.ch, by="City")

Chelsea.MA <- cbind(MA.Enrol.ch)

```
* The total number of students, who enrolled in preschools to 12 grade levels in `r Chelsea.MA$City` schools is `r Chelsea.MA$EnrolPK12`.\

\


## Employment
\


### Number of Jobs within Community Boundaries
\

```{r echo=FALSE, warning=FALSE, message=FALSE, results='hide', comment=""}

########################################################################################################################################

                                                #----------------------#
                                                #       EMPLOYMENT     #
                                                #----------------------#


##########################################    Number of Jobs within Community Boundaries   #############################################
MA.Num.Jobs.ch <- read_acs5year(year = endyr,
                     states = j,
                     table_contents = "B08008_003",
                     areas = area,
                     summary_level = "place") %>%
  rename(City = NAME,
         NumJobs = B08008_003) %>% 
	dplyr::select(City, NumJobs) %>%
  left_join(Chelsea.MA, MA.Num.Jobs.ch, by="City")

Chelsea.MA <- cbind(MA.Num.Jobs.ch)
```
* The number of jobs within community boundaries in `r Chelsea.MA$City` is `r Chelsea.MA$NumJobs`.\

Followed is the map of Number of Jobs within Community Boundaries of the city.\

```{r echo=FALSE, warning=FALSE, message=FALSE, comment="",results='hide'}

## MAP

ch.MA.job <- read_acs5year(year = endyr,
                     states = j,
                     table_contents = "B08008_003",
                     areas = area,
                     summary_level = "block group")

ch.job.get <- tidycensus::get_acs(geography = "block group",
													 survey="acs5",
													 state=j,
													 county = c,
													 year = endyr,
													 output = "wide",
													 variables = 'B08008_003', 
													 keep_geo_vars = T, geometry = TRUE) %>%
	rename(NumJobs = B08008_003E) 

chelsea.job <- merge(ch.job.get, ch.MA.job, by.x = 'NAME.y', by.y='NAME')


tm_shape(chelsea.job) + tm_polygons("NumJobs", style="jenks", palette="Reds", border.col = "darkgray", title="Number of Jobs") + tm_view(alpha = 0.85,view.legend.position = c("left", "bottom"), basemaps = "Stamen.TonerLite")
```

\


### Labor Force (Total number, participation)
\
```{r echo=FALSE, warning=FALSE, message=FALSE, results='hide', comment=""}
############################################# Labor Force (Total number, participation)  ####################################

MA.Labors.ch <- read_acs5year(year = endyr,
                     states = j,
                     table_contents = c("B23025_002", # Number of Labor Force 16 years and over
														              "B23025_003", # Number of Civilian Labor Force 16 years and over
														              "B23025_006", # Number of Armed Forces 16 years and over
														              "B23025_001"),
                     areas = area,
                     summary_level = "place") %>%
  rename(City = NAME) %>% 
  mutate(LabForce = B23025_002,
	       LabForcePer = B23025_002/B23025_001,
	       CivLabForce = B23025_003,
	       CivLabForcePer = B23025_003/B23025_001,
	       Military = B23025_006,
	       MilitaryPer = 	B23025_006/B23025_001,
	       TotalWorkers = B23025_001) %>%
	dplyr::select(City, LabForce, LabForcePer, CivLabForce, CivLabForcePer, Military, MilitaryPer, TotalWorkers)  %>%
  left_join(Chelsea.MA, MA.Labors.ch, by="City")

Chelsea.MA <- cbind(MA.Labors.ch)
```
* The number of labor force, who are 16 years and older in `r Chelsea.MA$City` is 89123 which is `r round(Chelsea.MA$LabForcePer,2)` of total number of labors (unemployed and not in labor force) of which 89052 are civilian labor forces and `r Chelsea.MA$Military` are armed labor forces.\

Followed is the map of Labor Force of the city.\

```{r echo=FALSE, warning=FALSE, message=FALSE, comment="",results='hide'}
## MAP

ch.MA.labors <- read_acs5year(year = endyr,
                     states = j,
                     table_contents = c("B23025_002", # Number of Labor Force 16 years and over
														              "B23025_003", # Number of Civilian Labor Force 16 years and over
														              "B23025_006", # Number of Armed Forces 16 years and over
														              "B23025_001"),
                     areas = area,
                     summary_level = "block group")

ch.labors.get <- tidycensus::get_acs(geography = "block group",
													 survey="acs5",
													 state=j,
													 county = c,
													 year = endyr,
													 output = "wide",
													 variables = c("B23025_002", # Number of Labor Force 16 years and over
														              "B23025_003", # Number of Civilian Labor Force 16 years and over
														              "B23025_006", # Number of Armed Forces 16 years and over
														              "B23025_001"), 
													 keep_geo_vars = T, geometry = TRUE) %>%
	mutate(LabForce = B23025_002E,
	       LabForcePer = B23025_002E/B23025_001E,
	       CivLabForce = B23025_003E,
	       CivLabForcePer = B23025_003E/B23025_001E,
	       Military = B23025_006E,
	       MilitaryPer = 	B23025_006E/B23025_001E,
	       TotalWorkers = B23025_001E) 

chelsea.labors <- merge(ch.labors.get, ch.MA.labors, by.x = 'NAME.y', by.y='NAME')


tm_shape(chelsea.labors) + tm_polygons("LabForce", style="jenks", palette="seq", border.col = "darkgray", title="Number of Labors") + tm_view(alpha = 0.85,view.legend.position = c("left", "bottom"), basemaps = "Stamen.TonerLite")
```

\

### Employed and Unemployed
\
```{r echo=FALSE, warning=FALSE, message=FALSE, results='hide', comment=""}
#####################################################    Employed and Unemployed   ####################################

MA.Emp.Unemp.ch <- read_acs5year(year = endyr,
                     states = j,
                     table_contents = c("B23025_003", # Number of civilian Labor Force 16 years and over
														              "B23025_004", # Number of Employed individuals 16 years and over
														              "B23025_005"),
                     areas = area,
                     summary_level = "place") %>%
  rename(City = NAME) %>%
  mutate(EmpRate = B23025_004/B23025_003,
	       UnempRate = B23025_005/B23025_003,
	       EmployedCLF = B23025_004,
	       UnemployedCLF = B23025_005) %>%
	dplyr::select(City ,EmpRate, UnempRate, EmployedCLF, UnemployedCLF) %>%
  left_join(Chelsea.MA, MA.Emp.Unemp.ch, by="City")

Chelsea.MA <- cbind(MA.Emp.Unemp.ch)

```
* The rate of employment in `r Chelsea.MA$City` is `r round(Chelsea.MA$EmpRate,2)*100`%.\
* The rate of unemployment in `r Chelsea.MA$City` is `r round(Chelsea.MA$UnempRate,2)*100`%.\

Followed is Employed and Unemployed map of the city.\

```{r echo=FALSE, warning=FALSE, message=FALSE, comment="",results='hide'}
## MAP
ch.MA.emp <- read_acs5year(year = endyr,
                     states = j,
                     table_contents = c("B23025_003", # Number of civilian Labor Force 16 years and over
														              "B23025_004", # Number of Employed individuals 16 years and over
														              "B23025_005"),
                     areas = area,
                     summary_level = "block group")

ch.emp.get <- tidycensus::get_acs(geography = "block group",
													 survey="acs5",
													 state=j,
													 county = c,
													 year = endyr,
													 output = "wide",
													 variables = c("B23025_003", # Number of civilian Labor Force 16 years and over
														              "B23025_004", # Number of Employed individuals 16 years and over
														              "B23025_005"), 
													 keep_geo_vars = T, geometry = TRUE) %>%
	mutate(EmpRate_BG = (B23025_004E/B23025_003E)*100,
	       UnempRate_BG = (B23025_005E/B23025_003E)*100,
	       EmployedCLF_BG = B23025_004E,
	       UnemployedCLF_BG = B23025_005E) 

chelsea.emp <- merge(ch.emp.get, ch.MA.emp, by.x = 'NAME.y', by.y='NAME')

tm_shape(chelsea.emp) + tm_polygons(c("EmpRate_BG","UnempRate_BG"), style="quantile", palette=list("Greens","seq"), border.col = "darkgray", title=c("Employment rate","Unemployment rate"), popup.format = list(fun = make_percent), legend.format = list(fun = make_percent))+ tm_view(alpha = 0.85, basemaps = "Stamen.TonerLite") + tm_layout(legend.position = c("left","bottom"))

```

\


### Major Employers by NCAIS Code
\
```{r echo=FALSE, warning=FALSE, message=FALSE, results='hide', comment=""}

################################    Major Employers by NCAIS Code    ##########################################
MA.Emp.NCAIS.ch <- read_acs5year(year = endyr,
                     states = j,
                     table_contents = c("C24030_002", # Number of male by industry for the civilian employed, 16 years and over
														              "C24030_029", # Number of female by industry for the civilian employed, 16 years and over
														              "C24030_001"),
                     areas = area,
                     summary_level = "place") %>%
  rename(City = NAME) %>%
  mutate(MaleNCAIS = C24030_002,
	       FemaleNCAIS = C24030_029,
	       MaleNCAISPer = C24030_002/C24030_001,
	       FemaleNCAISPer = C24030_029/C24030_001) %>%
	dplyr::select(City , MaleNCAIS, FemaleNCAIS, MaleNCAISPer, FemaleNCAISPer) %>%
  left_join(Chelsea.MA, MA.Emp.NCAIS.ch, by="City")

Chelsea.MA <- cbind(MA.Emp.NCAIS.ch)

```
* Number of male by industry for the civilian employed in `r Chelsea.MA$City` is `r Chelsea.MA$MaleNCAIS` which is `r round(Chelsea.MA$MaleNCAISPer,4)` of population, who are 16 years and over.\
* Number of female by industry for the civilian employed in `r Chelsea.MA$City` is `r Chelsea.MA$FemaleNCAIS` which is `r round(Chelsea.MA$FemaleNCAISPer,4)` of population, who are 16 years and over.\

Followed is the map of the Major Employers by NCAIS Code of the city.\

```{r echo=FALSE, warning=FALSE, message=FALSE, comment="",results='hide'}
## MAP
ch.MA.majorEmp <- read_acs5year(year = endyr,
                     states = j,
                     table_contents = c("C24030_002", # Number of male by industry for the civilian employed, 16 years and over
														              "C24030_029", # Number of female by industry for the civilian employed, 16 years and over
														              "C24030_001"),
                     areas = area,
                     summary_level = "block group")

ch.majorEmp.get <- tidycensus::get_acs(geography = "block group",
													 survey="acs5",
													 state=j,
													 county = c,
													 year = endyr,
													 output = "wide",
													 variables = c("C24030_002", # Number of male by industry for the civilian employed, 16 years and over
														              "C24030_029", # Number of female by industry for the civilian employed, 16 years and over
														              "C24030_001"), 
													 keep_geo_vars = T, geometry = TRUE) %>%
  mutate(MaleNCAIS_BG = C24030_002E,
	       FemaleNCAIS_BG = C24030_029E,
	       MaleNCAISPer_BG = (C24030_002E/C24030_001E)*100,
	       FemaleNCAISPer_BG = (C24030_029E/C24030_001E)*100) 

chelsea.majorEmp <- merge(ch.majorEmp.get, ch.MA.majorEmp, by.x = 'NAME.y', by.y='NAME')


tm_shape(chelsea.majorEmp) + tm_polygons(c("MaleNCAIS_BG","FemaleNCAIS_BG"), style="jenks", palette=list("Greens","seq"), border.col = "darkgray", title=c("MaleNCAIS","FemaleNCAIS")) + tm_view(alpha = 0.85, basemaps = "Stamen.TonerLite") + tm_layout(legend.position = c("left","bottom"))

```

\

```{r echo=FALSE, warning=FALSE, message=FALSE, results='hide', comment=""}
tm_shape(chelsea.majorEmp) + tm_polygons(c("MaleNCAISPer_BG","FemaleNCAISPer_BG"), style="quantile", palette=list("Blues","Reds"), border.col = "darkgray", title=c("MaleNCAIS Pecentage","FemaleNCAIS Percentage"),popup.format = list(fun = make_percent), legend.format = list(fun = make_percent)) + tm_view(alpha = 0.85, basemaps = "Stamen.TonerLite") + tm_layout(legend.position = c("left","bottom"))
```




\

### Top Industries by NCAIS Code
\
```{r echo=FALSE, warning=FALSE, message=FALSE, results='hide', comment=""}
######################################    Top Industries by NCAIS Code   ##################################################

MA.Industries.ch <- read_acs5year(year = endyr,
                     states = j,
                     table_contents = c("C24030_003", # Agriculture, forestry, fishing and hunting, and mining (Male)
														              "C24030_030", # Agriculture, forestry, fishing and hunting, and mining (Female)
														              "C24030_006", # Construction (Male)
														              "C24030_033", # Construction (Female)
														              "C24030_007", # Manufacturing (Male)
														              "C24030_034", # Manufacturing (Female)
														              "C24030_008", # Wholesale trade (Male)
														              "C24030_035", # Wholesale trade (Female)
														              "C24030_009", # Retail trade (Male)
														              "C24030_036", # Retail trade (Female)
														              "C24030_010", # Transportation and warehousing, and utilities (Male)
														              "C24030_037", # Transportation and warehousing, and utilities (Female)
														              "C24030_013", # Information (Male)
														              "C24030_040", # Information (Female)
														              "C24030_014", # Finance and insurance, and real estate, and rental and leasing (Male)
														              "C24030_041", # Finance and insurance, and real estate, and rental and leasing (Female)
														              "C24030_017", # Professional, scientific, and management, and administrative, and waste management services (Male)
														              "C24030_044", # Professional, scientific, and management, and administrative, and waste management services (Female)
														              "C24030_021", # Educational services, and health care and social assistance (Male)
														              "C24030_048", # Educational services, and health care and social assistance (Female)
														              "C24030_024", # Arts, entertainment, and recreation, and accommodation and food services (Male)
														              "C24030_051", # Arts, entertainment, and recreation, and accommodation and food services (Female)
														              "C24030_027", # Other services, except public administration (Male)
														              "C24030_054", # Other services, except public administration (Female)
														              "C24030_028", # Public administration (Male)
														              "C24030_055", # Public administration (Female)
														              "C24030_001"), # Total,
                     areas = area,
                     summary_level = "place") %>%
  rename(City = NAME) %>%
  mutate(Agriculture = (C24030_003 + C24030_030)/C24030_001,
	       Construction = (C24030_006 + C24030_033)/C24030_001,
	       Manufacturing = (C24030_007 + C24030_034)/C24030_001,
	       WholesaleTrade = (C24030_008 + C24030_035)/C24030_001,
	       RetailTrade = (C24030_009 + C24030_036)/C24030_001,
	       Transportation = (C24030_010 + C24030_037)/C24030_001,
	       Information = (C24030_013 + C24030_040)/C24030_001,
	       Finance_Ins = (C24030_014 + C24030_041)/C24030_001,
	       Professional = (C24030_017 + C24030_044)/C24030_001,
	       Educ_Health_SocSvcs = (C24030_021 + C24030_048)/C24030_001,
	       RecreationEtc = (C24030_024 + C24030_051)/C24030_001,
	       OtherIndustries = (C24030_027 + C24030_054)/C24030_001,
	       PublicAdmin = (C24030_028 + C24030_055)/C24030_001) %>%
	dplyr::select(City,Agriculture,Construction,Manufacturing,WholesaleTrade,RetailTrade,Transportation,Information,
	       Finance_Ins,Professional,Educ_Health_SocSvcs,RecreationEtc,OtherIndustries,PublicAdmin) %>%
  left_join(Chelsea.MA, MA.Industries.ch, by="City")

Chelsea.MA <- cbind(MA.Industries.ch)

```
* The percentage of people, who are in agriculture, forestry, fishing, hunting, and mining in `r Chelsea.MA$City` is `r (round(Chelsea.MA$Agriculture,4))*100`%.\

* The percentage of people, who are in construction in `r Chelsea.MA$City` state is `r (round(Chelsea.MA$Construction,4))*100`%.\

* The percentage of people, who are in manufacturing in `r Chelsea.MA$City` state is `r (round(Chelsea.MA$Manufacturing,4))*100`%.\

* The percentage of people, who are in wholesale trade in `r Chelsea.MA$City` state is `r (round(Chelsea.MA$WholesaleTrade,4))*100`%.\

* The percentage of people, who are in retail trade in `r Chelsea.MA$City` state is `r (round(Chelsea.MA$RetailTrade,4))*100`%.\

* The percentage of people, who are in transportation and warehousing, and utilities in `r Chelsea.MA$City`state is `r (round(Chelsea.MA$Transportation,4))*100`%.\

* The percentage of people, who are in information in `r Chelsea.MA$City` state is `r (round(Chelsea.MA$Information,4))*100`%.\

* The percentage of people, who are in finance and insurance, and real estate, and rental and leasing in `r Chelsea.MA$City` state is `r (round(Chelsea.MA$Finance_Ins,4))*100`%.\

* The percentage of people, who are in professional, scientific, and management, and administrative, and waste management services in `r Chelsea.MA$City` state is `r (round(Chelsea.MA$Professional,4))*100`%.\

* The percentage of people, who are in educational services, and health care and social assistance in `r Chelsea.MA$City` state is `r (round(Chelsea.MA$Educ_Health_SocSvcs,4))*100`%.\

* The percentage of people, who are in arts, entertainment, and recreation, and accommodation and food services in `r Chelsea.MA$City` state is `r (round(Chelsea.MA$RecreationEtc,4))*100`%.\

* The percentage of people, who are in other services, except public administration in `r Chelsea.MA$City` state is `r (round(Chelsea.MA$OtherIndustries,4))*100`%.\

* The percentage of people, who are in Public administration in `r Chelsea.MA$City` state is `r (round(Chelsea.MA$PublicAdmin,4))*100`%.\
\

### Earning Breakdown 
\

```{r echo=FALSE, warning=FALSE, message=FALSE, results='hide', comment=""}
###########################    Earning Breakdown   ####################################
MA.Earning.ch <- read_acs5year(year = endyr,
                     states = j,
                     table_contents = "B20017_001",
                     areas = area,
                     summary_level = "place") %>%
  rename(City = NAME) %>%
  mutate(MedEarning = B20017_001) %>%
	dplyr::select(City,MedEarning) %>%
  left_join(Chelsea.MA, MA.Earning.ch, by="City")

Chelsea.MA <- cbind(MA.Earning.ch)
```
* The median earning in `r Chelsea.MA$City` is 25786 Dollar.\

Followed is the Median Earning map of the city.\

```{r echo=FALSE, warning=FALSE, message=FALSE, comment="",results='hide'}
## MAP
ch.MA.earning <- read_acs5year(year = endyr,
                     states = j,
                     table_contents = "B20017_001",
                     areas = area,
                     summary_level = "block group")

ch.earning.get <- tidycensus::get_acs(geography = "block group",
													 survey="acs5",
													 state=j,
													 county = c,
													 year = endyr,
													 output = "wide",
													 variables = 'B20017_001', 
													 keep_geo_vars = T, geometry = TRUE) %>%
  mutate(MedEarning = B20017_001E)

chelsea.earning <- merge(ch.earning.get, ch.MA.earning, by.x = 'NAME.y', by.y='NAME')

make_dollar <- function(x, digits = 0) {
  paste0("$", formatC(x, digits = digits, format = "f", big.mark = ","))
}


tm_shape(chelsea.earning) + tm_polygons("MedEarning" , palette="seq", style = "jenks", title = "Median Earning", popup.format = list(fun = make_dollar), legend.format = list(fun = make_dollar)) + tm_view(alpha = 0.85,view.legend.position = c("left", "bottom"),basemaps = "Stamen.TonerLite") + tm_borders(col = "darkgray") 


```

\


### Employee Age
\
```{r echo=FALSE, warning=FALSE, message=FALSE, results='hide', comment=""}
#############################################    Employee Age   #############################################################

MA.Emp.Age.ch <- read_acs5year(year = endyr,
                     states = j,
                     table_contents = c("B23001_005","B23001_007","B23001_012","B23001_014","B23001_019","B23001_021","B23001_026","B23001_028", # Age 29 or younger male
								"B23001_033","B23001_035","B23001_040","B23001_042","B23001_047","B23001_049", # Age 30 to 54 male
								"B23001_054","B23001_056","B23001_061","B23001_063","B23001_068","B23001_070","B23001_075","B23001_080","B23001_085", # Age 55 or older male
								"B23001_091","B23001_093","B23001_098","B23001_100","B23001_105","B23001_107","B23001_112","B23001_114", # Age 29 or younger female
								"B23001_119","B23001_121","B23001_126","B23001_128","B23001_133","B23001_135", # Age 30 to 54 female
								"B23001_140","B23001_142","B23001_147","B23001_149","B23001_154","B23001_156","B23001_161","B23001_166","B23001_171", # Age 55 or older female
														              "B23001_001"),
                     areas = area,
                     summary_level = "place") %>%
  rename(City = NAME) %>%
  mutate(AgeEmpU29 = B23001_005+B23001_007+B23001_012+B23001_014+B23001_019+B23001_021+B23001_026+B23001_028+
	         B23001_091+B23001_093+B23001_098+B23001_100+B23001_105+B23001_107+B23001_112+B23001_114,
	       AgeEmp3054 = B23001_033+B23001_035+B23001_040+B23001_042+B23001_047+B23001_049+
	         B23001_119+B23001_121+B23001_126+B23001_128+B23001_133+B23001_135,
	       AgeEmpO55 = B23001_054+B23001_056+B23001_061+B23001_063+B23001_068+B23001_070+B23001_075+B23001_080+B23001_085+
	        B23001_140+B23001_142+B23001_147+B23001_149+B23001_154+B23001_156+B23001_161+B23001_166+B23001_171,
         
         AgeEmpU29Per_BG = (B23001_005+B23001_007+B23001_012+B23001_014+B23001_019+B23001_021+B23001_026+B23001_028+
	         B23001_091+B23001_093+B23001_098+B23001_100+B23001_105+B23001_107+B23001_112+B23001_114)/B23001_001,
	        AgeEmp3054Per_BG = (B23001_033+B23001_035+B23001_040+B23001_042+B23001_047+B23001_049+
	         B23001_119+B23001_121+B23001_126+B23001_128+B23001_133+B23001_135)/B23001_001,
	       AgeEmpO55Per_BG = (B23001_054+B23001_056+B23001_061+B23001_063+B23001_068+B23001_070+B23001_075+B23001_080+B23001_085+
	        B23001_140+B23001_142+B23001_147+B23001_149+B23001_154+B23001_156+B23001_161+B23001_166+B23001_171)/B23001_001) %>%

	dplyr::select(City ,AgeEmpU29, AgeEmp3054, AgeEmpO55, AgeEmpU29Per_BG,AgeEmp3054Per_BG,AgeEmpO55Per_BG) %>%
  left_join(Chelsea.MA, MA.Emp.Age.ch, by="City")

Chelsea.MA <- cbind(MA.Emp.Age.ch)

```
* The number of workers, who are younger than 29 years old in `r Chelsea.MA$City` is `r Chelsea.MA$AgeEmpU29` which is `r (round(Chelsea.MA$AgeEmpU29Per,4))*100`% of total number of workers.\

* The number of workers, who are between 30 and 54 years old in `r Chelsea.MA$City` is `r Chelsea.MA$AgeEmp3054` which is `r (round(Chelsea.MA$AgeEmp3054Per,4))*100`% of total number of workers.\

* The number of workers, who are older than 55 years old in `r Chelsea.MA$City` is `r Chelsea.MA$AgeEmpO55` which is `r (round(Chelsea.MA$AgeEmpO55Per,4))*100`% of total number of workers.\

\


### Employee Race
\
```{r echo=FALSE, warning=FALSE, message=FALSE, results='hide', comment=""}
#############################################      Employee Race    ############################################################
MA.Emp.Race.ch <- read_acs5year(year = endyr,
                     states = j,
                     table_contents = c("C23002A_007","C23002A_012","C23002A_020","C23002A_025","C23002A_005","C23002A_018", # White alone
														              "C23002B_007","C23002B_012","C23002B_020","C23002B_025","C23002B_005","C23002B_018", # Black or African American alone
														              "C23002D_007","C23002D_012","C23002D_020","C23002D_025","C23002D_005","C23002D_018", # Asian alone
														              "C23002I_007","C23002I_012","C23002I_020","C23002I_025","C23002I_005","C23002I_018", # Hispanic or Latino alone
														              "C23002G_007","C23002G_012","C23002G_020","C23002G_025","C23002G_005","C23002G_018",# Two or more races
														              "B23001_001"),
                     areas = area,
                     summary_level = "place") %>%
  rename(City = NAME) %>%
  mutate(WhiteEmp = C23002A_007+C23002A_012+C23002A_020+C23002A_025+C23002A_005+C23002A_018,
	       BlackEmp = C23002B_007+C23002B_012+C23002B_020+C23002B_025+C23002B_005+C23002B_018,
	       AsianEmp = C23002D_007+C23002D_012+C23002D_020+C23002D_025+C23002D_005+C23002D_018,
	       HispanicEmp = C23002I_007+C23002I_012+C23002I_020+C23002I_025+C23002I_005+C23002I_018,
	       TwoOrMoreEmp = C23002G_007+C23002G_012+C23002G_020+C23002G_025+C23002G_005+C23002G_018,
         WhiteEmpPer = (C23002A_007+C23002A_012+C23002A_020+C23002A_025+C23002A_005+C23002A_018)/B23001_001,
	       BlackEmpPer = (C23002B_007+C23002B_012+C23002B_020+C23002B_025+C23002B_005+C23002B_018)/B23001_001,
	       AsianEmpPer = (C23002D_007+C23002D_012+C23002D_020+C23002D_025+C23002D_005+C23002D_018)/B23001_001,
	       HispanicEmpPer = (C23002I_007+C23002I_012+C23002I_020+C23002I_025+C23002I_005+C23002I_018)/B23001_001,
	       TwoOrMoreEmpPer = (C23002G_007+C23002G_012+C23002G_020+C23002G_025+C23002G_005+C23002G_018)/B23001_001) %>%
	dplyr::select(City ,WhiteEmp, BlackEmp, AsianEmp, HispanicEmp, TwoOrMoreEmp,
	       WhiteEmpPer, BlackEmpPer, AsianEmpPer, HispanicEmpPer, TwoOrMoreEmpPer) %>%
  left_join(Chelsea.MA, MA.Emp.Race.ch, by="City")

Chelsea.MA <- cbind(MA.Emp.Race.ch)

```

* The number of white workers in `r Chelsea.MA$City` is `r Chelsea.MA$WhiteEmp` which is `r (round(Chelsea.MA$WhiteEmpPer,4))*100`% of total number of workers.\
* The number of black workers in `r Chelsea.MA$City` is `r Chelsea.MA$BlackEmp` which is `r (round(Chelsea.MA$BlackEmpPer,4))*100`% of total number of workers.\
* The number of asian workers in `r Chelsea.MA$City` is `r Chelsea.MA$AsianEmp` which is `r (round(Chelsea.MA$AsianEmpPer,4))*100`% of total number of workers.\
* The number of hispanic or latino workers in `r Chelsea.MA$City` is `r Chelsea.MA$HispanicEmp` which is `r (round(Chelsea.MA$HispanicEmpPer,4))*100`% of total number of workers.\
* The number of workers, who have two or more races in `r Chelsea.MA$City` is `r Chelsea.MA$TwoOrMoreEmp` which is `r (round(Chelsea.MA$TwoOrMoreEmpPer,4))*100`% of total number of workers.\
\


### Employee Education
\
```{r message=FALSE, warning=FALSE, ,echo=FALSE, results='hide', comment=""}
#############################################   Employee Education    ########################################################

MA.Educ.Emp.ch <- read_acs5year(year = endyr,
                     states = j,
                     table_contents = c("B16010_003", # Less than high school graduate in labor force
														              "B16010_016", # High school graduate in labor force 
														              "B16010_029", # Some college or associate's degree in labor force
														              "B16010_042",
														              "B16010_001"), # Total
                     areas = area,
                     summary_level = "place") %>%
  rename(City = NAME) %>%
  	mutate(LessThanHSLabor = B16010_003,
	       HSGradLabor = B16010_016,
	       SomeCollLabor = B16010_029,
	       BachOrHigherLabor = B16010_042,
	       LessThanHSLaborPer = LessThanHSLabor/B16010_001,
	       HSGradLaborPer = HSGradLabor/B16010_001,
	       SomeCollLaborPer = SomeCollLabor/B16010_001,
	       BachOrHigherLaborPer = BachOrHigherLabor/B16010_001) %>%
	dplyr::select(City ,LessThanHSLabor, HSGradLabor, SomeCollLabor, BachOrHigherLabor,
	       LessThanHSLaborPer, HSGradLaborPer, SomeCollLaborPer, BachOrHigherLaborPer) %>%
  left_join(Chelsea.MA, MA.Educ.Emp.ch, by="City")

Chelsea.MA <- cbind(MA.Educ.Emp.ch)
```
* The number of workers, who have less than high school degree in `r Chelsea.MA$City` is `r Chelsea.MA$LessThanHSLabor` which is `r (round(Chelsea.MA$LessThanHSLaborPer,4))*100`% of total number of workers.\
* The number of workers, who have high school degree in `r Chelsea.MA$City` is `r Chelsea.MA$HSGradLabor` which is `r (round(Chelsea.MA$HSGradLaborPer,4))*100`% of total number of workers.\
* The number of workers, who have some college or associates degree in `r Chelsea.MA$City` is `r Chelsea.MA$SomeCollLabor` which is `r (round(Chelsea.MA$SomeCollLaborPer,4))*100`% of total number of workers.\
* The number of workers, who have bachelors degree or higher in `r Chelsea.MA$City` is `r Chelsea.MA$BachOrHigherLabor` which is `r (round(Chelsea.MA$BachOrHigherLaborPer,4))*100`% of total number of workers.\

\

### Employee Sex
\
```{r echo=FALSE, warning=FALSE, message=FALSE, results='hide', comment=""}
#############################################    Employee Sex   #############################################################

MA.Emp.Sex.ch <- read_acs5year(year = endyr,
                     states = j,
                     table_contents = c("B23001_005", # Employed males
														              "B23001_007", 
														              "B23001_012",
														              "B23001_014",
														              "B23001_019",
														              "B23001_021",
														              "B23001_026",
														              "B23001_028",
														              "B23001_033",
														              "B23001_035",
														              "B23001_040",
														              "B23001_042",
														              "B23001_047",
														              "B23001_049",
														              "B23001_054",
														              "B23001_056",
														              "B23001_061",
														              "B23001_063",
														              "B23001_068",
														              "B23001_070",
														              "B23001_075",
														              "B23001_080",
														              "B23001_085",
														              "B23001_091", # Employed females
														              "B23001_093",
														              "B23001_098",
														              "B23001_100",
														              "B23001_105",
														              "B23001_107",
														              "B23001_112",
														              "B23001_114",
														              "B23001_119",
														              "B23001_121",
														              "B23001_126",
														              "B23001_128",
														              "B23001_133",
														              "B23001_135",
														              "B23001_140",
														              "B23001_142",
														              "B23001_147",
														              "B23001_149",
														              "B23001_154",
														              "B23001_156",
														              "B23001_161",
														              "B23001_166",
														              "B23001_171",
														              "B23001_001"),
                     areas = area,
                     summary_level = "place") %>%
  rename(City = NAME) %>%
  	mutate(EmpMale = B23001_005 + B23001_007 + B23001_012 + B23001_014 + B23001_019 + B23001_021 + B23001_026 + B23001_028 + B23001_033 + B23001_035 + B23001_040 + B23001_042 + B23001_047 + B23001_049 + B23001_054 + B23001_056 + B23001_061 + B23001_063 + B23001_068+ B23001_070 + B23001_075 + B23001_080 + B23001_085,
	  EmpFemale = B23001_091 + B23001_093 + B23001_098 + B23001_100 + B23001_105 + B23001_107 + B23001_112 + B23001_114 + B23001_119 + B23001_121 + B23001_126+ B23001_128 + B23001_133 + B23001_135 + B23001_140 + B23001_142 + B23001_147 + B23001_149 + B23001_154 + B23001_156 + B23001_161 + B23001_166 + B23001_171,
	  TotalEmp = EmpMale + EmpFemale,
	  EmpMalePer = EmpMale/	B23001_001,
	  EmpFemalePer = EmpFemale/B23001_001) %>%
	dplyr::select(City, EmpMale, EmpFemale, TotalEmp, EmpMalePer, EmpFemalePer) %>%
  left_join(Chelsea.MA, MA.Emp.Sex.ch, by="City")

Chelsea.MA <- cbind(MA.Emp.Sex.ch)


```
* The number of male workers in `r Chelsea.MA$City` is `r Chelsea.MA$EmpMale` which is `r (round(Chelsea.MA$EmpMalePer,4))*100`% of total number of workers.\
* The number of female workers in `r Chelsea.MA$City` is `r Chelsea.MA$EmpFemale` which is `r (round(Chelsea.MA$EmpFemalePer,4))*100`% of total number of workers.\


\

## Healthcare
\

### Number of Hospitals
\

```{r echo=FALSE, warning=FALSE, message=FALSE, results='hide', comment=""}

Chelsea.MA$Hospitals <- 9
```


* `r Chelsea.MA$City` has 9 hospitals.(https://www.ahd.com/)

\

###	Number of Nursing Homes 
\

```{r echo=FALSE, warning=FALSE, message=FALSE, results='hide', comment=""}
Chelsea.MA$NursingHomes <- 11
```


* `r Chelsea.MA$City` has 11 nursing homes. (https://nursinghomedirectory.com/)


## Municipal Finances
\

###	Revenue Streams
\


```{r echo=FALSE, warning=FALSE, message=FALSE, results='hide', comment=""}
Chelsea.MA$RevenueGeneralFundTotalFY19 <- 493275000
Chelsea.MA$RevenueTaxFY19 <- 348465000
Chelsea.MA$RevenueOtherTaxesFY19 <- 90488000
Chelsea.MA$RevenueExecutiveLegislativeJudicialFY19 <- 336000
Chelsea.MA$RevenueFinanceFY19 <- 500000
Chelsea.MA$RevenuePublicSafetyFY19 <- 12995000
Chelsea.MA$RevenueBuildingInspectionDepartmentFY19 <- 8603000
Chelsea.MA$RevenuePublicWorksFY19 <- 253000
Chelsea.MA$RevenuePublicLandsAndParksFY19 <- 300000
Chelsea.MA$RevenueOtherDepartmentsFY19 <- 24064000
Chelsea.MA$RevenuePublicPropertiesFY19 <- 126000
Chelsea.MA$RevenueTransfersFY19 <- 7145000
```
* The total budget revenues for general fund for Providence city is $493,275,000 in fiscal year 2019, which contains the following items: Tax revenues include property tax and interest on overdue taxes with 348,465,000 Dollar, Other taxes with 90,488,000 Dollar, executive, legislative, and judicial with 336,000 Dollar, finance with 500,000 Dollar, public safety with 12,995,000 Dollar, building inspection department with 8,603,000 Dollar, public works with 253,000 Dollar, public lands and parks with 300,000 Dollar, other departments with 24,064,000 Dollar, public properties with 126,000 Dollar, and transfers with 7,145,000 Dollar.(https://www.providenceri.gov/wp-content/uploads/2020/02/2019-CAFR.pdf)

\

###	Major Expense Drivers
\

```{r echo=FALSE, warning=FALSE, message=FALSE, results='hide', comment=""}
Chelsea.MA$ExpendituresGeneralFundTotalFY19 <- 493274000
Chelsea.MA$ExpendituresExecutiveLegislativeJudicialFY19 <- 13979000
Chelsea.MA$ExpendituresFinanceFY19 <- 84201000
Chelsea.MA$ExpendituresPublicSafetyFY19 <- 165787000
Chelsea.MA$ExpendituresBuildingInspectionDepartmentFY19 <- 4789000
Chelsea.MA$ExpendituresPublicWorksFY19 <- 22133000
Chelsea.MA$ExpendituresRecreationFY19 <- 2945000
Chelsea.MA$ExpendituresPublicLandsAndParksFY19 <- 13592000
Chelsea.MA$ExpendituresOtherDepartmentsFY19 <- 17311000
Chelsea.MA$ExpendituresGrantsFY19 <- 5653000
Chelsea.MA$ExpendituresPublicPropertiesFY19 <- 6692000
Chelsea.MA$ExpendituresTransfersFY19 <- 156192000
```

* The total amount of general fund of expenditures for `r Chelsea.MA$City` is $493,274,000 in Fiscal Year 2019 and the major expense driver is related to public safety, which is 165,787,000 Dollar. The other expenditures include executive, legislative, and judicial with 13,979,000 Dollar, finance with 84,201,000 Dollar, public works with 22,133,000 Dollar, building inspection department with 4,789,000 Dollar, recreation with 2,945,000 Dollar, public lands and parks with 13,592,000 Dollar, other departments with 17,311,000 Dollar, grants with 5,653,000 Dollar, public properties with 6,692,000 Dollar, and transfers with 156,192,000 Dollar.(https://www.providenceri.gov/wp-content/uploads/2020/02/2019-CAFR.pdf)


\

### FY20 Budget  Projected Deficits 
\


```{r echo=FALSE, warning=FALSE, message=FALSE, results='hide', comment=""}
Chelsea.MA$RevenueBudgetTotalCombinedCityGeneralAndSchoolFY20 <- 770169494
Chelsea.MA$RevenueBudgetTotalGeneralFY20 <- 506060691
Chelsea.MA$RevenueBudgetTotalSchoolFundFY20 <- 264108803
Chelsea.MA$RevenueBudgetTaxesFY20 <- 345218386
Chelsea.MA$RevenueBudgetStateRevenuesAndAidFY20 <- 93956371
Chelsea.MA$RevenueBudgetLocalReceiptsFY20 <- 66885934

Chelsea.MA$ExpendituresBudgetMayorsOfficeFY20 <- 3538778
Chelsea.MA$ExpendituresBudgetFinanceDepartmentFY20 <- 7646945
Chelsea.MA$ExpendituresBudgetHumanResourceFY20 <- 4537440
Chelsea.MA$ExpendituresBudgetPublicSafetyFY20 <- 173630049
Chelsea.MA$ExpendituresBudgetPublicWorksFY20 <- 24435853
Chelsea.MA$ExpendituresBudgetRecreationFY20 <- 3322143
Chelsea.MA$ExpendituresBudgetParksFY20 <- 13959798
Chelsea.MA$ExpendituresBudgetInspectionsAndStandardsFY20 <- 4551508
Chelsea.MA$ExpendituresBudgetHumanServicesFY20 <- 2936081
Chelsea.MA$ExpendituresBudgetMiscellaneousDepartmentsFY20 <- 2061695
Chelsea.MA$ExpendituresBudgetUnallocatedExpensesFY20 <- 87038914
Chelsea.MA$ExpendituresBudgetCityCouncilFY20 <- 7258161
Chelsea.MA$ExpendituresBudgetDeficitReductionPlanFY20 <- 3784500
Chelsea.MA$ExpendituresBudgetTotalGeneralServicesFY20 <- 376014080
Chelsea.MA$ExpendituresBudgetSchoolAppropriationFY20 <- 130046611
Chelsea.MA$ExpendituresBudgetTotalGeneralFY20 <- 506060691
Chelsea.MA$ExpendituresBudgetTotalSchoolFundFY20 <- 264108803
Chelsea.MA$ExpendituresBudgetTotalCombinedCityGeneralAndSchoolFY20 <- 770169494

```

* The total general budget revenues for Providence city is $506,060,691 in fiscal year 2020 and the total amount of passthrough school fund revenue is 264,108,803 Dollar. Therefore, the total amount of combined city  general (municipal) and school revenues is 770,169,494 Dollar. General fund contains taxes with 345,218,386 Dollar, state revenues and aid with 93,956,371 Dollar, and local receipts with 66,885,934 Dollar. The total general budget expenditures for Providence city in fiscal year 2020 is 506,060,691 Dollar and the total amount of passthrough school fund expenditures is 264,108,803 Dollar. Thus, the total amount of combined city  general (municipal) and school expenditures is 770,169,494 Dollar. General fund expenditures include mayor's office with 3,538,778 Dollar, finance department with 7,646,945 Dollar, human resource with 4,537,440 Dollar, public safety with 173,630,049 Dollar, public works with 24,435,853 Dollar, recreation with 3,322,143 Dollar, parks with 13,959,798 Dollar, inspections and standards with 4,551,508 Dollar, human services with 2,936,081 Dollar, miscellaneous departments with 2,061,695 Dollar, unallocated expenses with 87,038,914 Dollar, city council with 7,258,161 Dollar, deficit reduction plan with 3,784,500 Dollar, general services with 376,014,080 Dollar, and school appropriation with 130,046,611 Dollar.(https://www.providenceri.gov/wp-content/uploads/2019/07/Approved-FY2020-Budget-Book.pdf)


\


## Infrastructure
\

### Presence of Nearby Ports, Airports, Major Rail Facilities, or Intermodal Freight Facilities
\

* Providence Airport (IATA: PVD; ICAO: KPVD), officially named Theodore Francis Green Memorial State Airport (T.F. Green Airport), is the main international airport in Rhode Island. Providence Airport is located in the city of Warwick, 10 km (6 miles) south of Providence city centre. This Airport held 4,298,345 passengers in 2018. "Top 10 domestic routes from T.F. Green Airport are Orlando, Baltimore, Charlotte, Washington-National, Philadelphia, Atlanta, Chicago-Midway, Fort Lauderdale, Tampa and Detroit".(https://www.providence-airport.com). There is a famous port near Providence city, whose name is ProvPort. "The Port of Providence is New England's second biggest deep water port and a strategically located Northeast port anchored by a strong tenant base each of which utilizes the port as a distribution center within the New England area". (https://www.provport.com/provport/provport.html) Providence is a railroad station in Providence, Rhode Island, served by Amtrak and MBTA Commuter Rail. There are four tracks and two island platforms for passenger service in the station, with a fifth track passing through for Providence and Worcester Railroad freight trains. The Providence train station is located at 100 Gaspee Street, in the heart of the downtown area. (https://en.wikipedia.org/wiki/Providence_station)



\

### Public Transport Reliability and Scope 
\

* The Rhode Island Public Transit Authority (RIPTA) offers low-cost bus service throughout Rhode Island. RIPTA offers statewide bus and citywide trolley service into downtown Providence. Also, Downtown Providence is served by 58 bus lines providing direct services into downtown from throughout the state. The other option for public transport in Providence is trains. Another option is MBTA commuter rail service, which offers 14 daily round trips between Boston and Providence, including convenient peak hour trains. MBTA service has also added trips south of Providence to T.F.Green Airports Interlink station and North Kingstowns Wickford Junction during the morning and evening commutes.(http://downtownprovidence.com/getting-here-parking/alternatives-to-driving/)


\

### Internet Access
\

* There are 17 internet providers in Providence with 5 of those offering residential service. Providence is the 5th most connected city in Rhode Island  and the fastest zip code in Providence for June 2020 is 02940. The most common Internet option in Providence is Viasat Internet, which has coverage for near one hundred percent of Providence residents. HughesNet has registered availability for about practically one hundred percent of the population. Internet competition is above-average in Providence in comparison to other parts of the state, with are 3.81 providers per census block on average. Providence ranks at 5th on the "most connected city" charts. (Measured by number of providers with coverage at each local address) As it stands, only 1.19 percent of Providence addresses have one or fewer choices for Internet.(https://broadbandnow.com/Rhode-Island/Providence)




\

## Food Insecurity
\

\

### Number of Grocery Stores
\

```{r echo=FALSE, warning=FALSE, message=FALSE, results='hide', comment=""}
Chelsea.MA$NumGrocery <- 129
```
* `r Chelsea.MA$City` has 129 Grocery Stores. (https://www.city-data.com/city/Providence-Rhode-Island.html)

\

### SNAP Recipients 
\


```{r echo=FALSE, warning=FALSE, message=FALSE, results='hide', comment=""}
##########################################################################################################################

                                                #----------------------#
                                                #    FOOD INSECURITY   #
                                                #----------------------#

####################################################  SNAP Recipients  ###################################################         

MA.Snap.ch <- read_acs5year(year = endyr,
                     states = j,
                     table_contents = c("B22001_002", # Household received Food Stamps/SNAP in the past 12 months
														"B22001_001"),
                     areas = area,
                     summary_level = "place") %>%
  rename(City = NAME) %>%
  	mutate(SnapRec = B22001_002/B22001_001) %>%
	dplyr::select(City ,SnapRec) %>%
  left_join(Chelsea.MA, MA.Snap.ch, by="City")

Chelsea.MA <- cbind(MA.Snap.ch)

#######################################################################################################################################

```
* The percent of households, who received Food Stamps/SNAP in the past 12 months in `r Chelsea.MA$City` state is `r (round(Chelsea.MA$SnapRec,4))*100`%.\

\

## Housing
\

### Proportion of Renters to Owners
\

### Home Values
\
```{r echo=FALSE, warning=FALSE, message=FALSE, results='hide', comment=""}

                                                #----------------------#
                                                #       HOUSING        #
                                                #----------------------#

######################   Proportion of Renters to Owners, Home Values   ########################################

MA.Housing.ch <- read_acs5year(year = endyr,
                     states = j,
                     table_contents = c("B25009_001", # Total Occupied Housing Units
															 	"B25009_002", # Total Owner Occupied Housing Units
															 	"B25009_010", # Total Renter Occupied Housing Units
															 	"B25077_001", # Median Home Value
															 	"B25082_001", # Aggregate value
															 	"B25003_002"), # Owner occupied
                     areas = area,
                     summary_level = "place") %>%
  rename(City = NAME,
         MedianHValue = B25077_001,
	       AvgHValue = B25082_001/B25003_002) %>%
	mutate(RentersPer = B25009_010/B25009_001,
				 HomeOwnPer = B25009_002/B25009_001,
				 RentersToOwners = B25009_010/B25009_002) %>%
	dplyr::select(City,RentersPer, HomeOwnPer, RentersToOwners, MedianHValue, AvgHValue) %>%
  left_join(Chelsea.MA, MA.Housing.ch, by="City")

Chelsea.MA <- cbind(MA.Housing.ch)
```
* The median house value in `r Chelsea.MA$City` is 192100 Dollar.\

* The average house value in `r Chelsea.MA$City` is 5670391500 Dollar.\

* The proportion of renters to owners in `r Chelsea.MA$City` is `r round(Chelsea.MA$RentersToOwners,3)`.\

* The percentage of renters in `r Chelsea.MA$City` is `r (round(Chelsea.MA$RentersPer,2))*100`%.\

* The percentage of owners in `r Chelsea.MA$City` is `r (round(Chelsea.MA$HomeOwnPer,2))*100`%.\

Followed is the map of Proportion of Renters to Owners, Median Home Values, and Average Home Values of the city.\

```{r echo=FALSE, warning=FALSE, message=FALSE, comment="", results='hide'}
## MAP
ch.MA.housing <- read_acs5year(year = endyr,
                     states = j,
                     table_contents = c("B25009_001", # Total Occupied Housing Units
															 	"B25009_002", # Total Owner Occupied Housing Units
															 	"B25009_010", # Total Renter Occupied Housing Units
															 	"B25077_001", # Median Home Value
															 	"B25082_001", # Aggregate value
															 	"B25003_002"),
                     areas = area,
                     summary_level = "block group")

ch.housing.get <- tidycensus::get_acs(geography = "block group",
													 survey="acs5",
													 state=j,
													 county = c,
													 year = endyr,
													 output = "wide",
													 variables = c("B25009_001", # Total Occupied Housing Units
															 	"B25009_002", # Total Owner Occupied Housing Units
															 	"B25009_010", # Total Renter Occupied Housing Units
															 	"B25077_001", # Median Home Value
															 	"B25082_001", # Aggregate value
															 	"B25003_002"), 
													 keep_geo_vars = T, geometry = TRUE) %>%
rename(MedianHValue_BG = B25077_001E,
	       AvgHValue_BG = B25082_001E / B25003_002E) %>%
	mutate(RentersPer_BG = B25009_010E/B25009_001E,
				 HomeOwnPer_BG = B25009_002E/B25009_001E,
				 RentersToOwners_BG = B25009_010E/B25009_002E)

chelsea.housing <- merge(ch.housing.get, ch.MA.housing, by.x = 'NAME.y', by.y='NAME')


tm_shape(chelsea.housing) + tm_polygons("MedianHValue_BG" , palette="Blues", style = "quantile", title = "Median Home Value", popup.format = list(fun = make_dollar), legend.format = list(fun = make_dollar)) + tm_view(alpha = 0.85,basemaps = "Stamen.TonerLite") + tm_borders(col = "darkgray") + tm_layout(legend.position = c("left","bottom"))

```
\


```{r echo=FALSE, warning=FALSE, message=FALSE, results='hide', comment=""}
tm_shape(chelsea.housing) + tm_polygons("AvgHValue_BG" , palette="Greens", style = "quantile", title = "Average Home Value", popup.format = list(fun = make_dollar), legend.format = list(fun = make_dollar)) + tm_view(alpha = 0.85,basemaps = "Stamen.TonerLite") + tm_borders(col = "darkgray")+tm_layout(legend.position = c("left","bottom"))
```


\


\

```{r echo=FALSE, warning=FALSE, message=FALSE, results='hide', comment=""}
tm_shape(chelsea.housing) + tm_polygons("RentersToOwners_BG" , palette="Reds", style = "jenks", title = "RentersToOwners") + tm_view(alpha = 0.85,basemaps = "Stamen.TonerLite") + tm_borders(col = "darkgray") + tm_layout(legend.position = c("left","bottom"))
```

\

### Property Tax Rates
\



```{r echo=FALSE, warning=FALSE, message=FALSE, results='hide', comment=""}
Chelsea.MA$PropertyTaxRate <- 18.80
```
* Providences property tax is $18.80 per 1,000 Dollar of value for owner-occupied homes; 31.96 Dollar per 1,000 Dollar of value for non-owner-occupied homes.(https://www.sna.providence.ri.us/2019-property-tax-proposals/)


\



\


### Duration of Occupancy
\

```{r echo=FALSE, warning=FALSE, message=FALSE, results='hide', comment=""}
#########################################      Duration of Occupancy    ####################################################
MA.Dur.Occ.ch <- read_acs5year(year = endyr,
                     states = j,
                     table_contents = c("B25039_001", # Median Year Householder Moved Into Unit by Tenure
															 	"B25039_002", # Median Year Householder Moved Into Unit by Tenure (Owner)
															 	"B25039_003"),
                     areas = area,
                     summary_level = "place") %>%
  rename(City = NAME,
         MedYrMovedIn = B25039_001, 
				 MedYrRentMovedIn = B25039_003,
				 MedYrOwnMovedIn = B25039_002) %>%
	dplyr::select(City, MedYrMovedIn, MedYrRentMovedIn, MedYrOwnMovedIn) %>%
  left_join(Chelsea.MA, MA.Dur.Occ.ch, by="City")

Chelsea.MA <- cbind(MA.Dur.Occ.ch)
```

* The median year that a homeowner moved into their current domicile in `r Chelsea.MA$City` is `r Chelsea.MA$MedYrOwnMovedIn`.\

* The median year that a renter moved into their current domicile in `r Chelsea.MA$City` is `r Chelsea.MA$MedYrRentMovedIn`.\

Followed is the map of Duration of Occupancy of the city.\

```{r echo=FALSE, warning=FALSE, message=FALSE, comment="", results='hide'}
## MAP
ch.MA.dur <- read_acs5year(year = endyr,
                     states = j,
                     table_contents = c("B25039_001", # Median Year Householder Moved Into Unit by Tenure
															 	"B25039_002", # Median Year Householder Moved Into Unit by Tenure (Owner)
															 	"B25039_003"),
                     areas = area,
                     summary_level = "block group")

ch.dur.get <- tidycensus::get_acs(geography = "block group",
													 survey="acs5",
													 state=j,
													 county = c,
													 year = endyr,
													 output = "wide",
													 variables = c("B25039_001", # Median Year Householder Moved Into Unit by Tenure
															 	"B25039_002", # Median Year Householder Moved Into Unit by Tenure (Owner)
															 	"B25039_003"), 
													 keep_geo_vars = T, geometry = TRUE) %>%
rename(MedYrMovedIn_BG = B25039_001E, # I'm not sure if I should substract this value from 2018
				 MedYrRentMovedIn_BG = B25039_003E,
				 MedYrOwnMovedIn_BG = B25039_002E)

chelsea.dur <- merge(ch.dur.get, ch.MA.dur, by.x = 'NAME.y', by.y='NAME')



tm_shape(chelsea.dur) + tm_polygons(c("MedYrMovedIn_BG","MedYrRentMovedIn_BG","MedYrOwnMovedIn_BG"), style="quantile", palette=list("Greens","seq","Blues"), border.col = "darkgray", title=c("MedYrMovedIn","MedYrRentMovedIn","MedYrOwnMovedIn")) + tm_view(alpha = 0.85, basemaps = "Stamen.TonerLite") + tm_layout(legend.position = c("left","bottom"))
```


\

### Cost Burdened Occupants


\

```{r echo=FALSE, warning=FALSE, message=FALSE, results='hide', comment=""}

##############################################   Cost Burdened Occupants   ##################################
MA.CostBurdened.ch <- read_acs5year(year = endyr,
                     states = j,
                     table_contents = c("B25095_006", # Household income by selected monthly owner Costs as a percentage of household income
														              # Less than $10,000 (30 to 34.9 percent)
														              "B25095_007", # Less than $10,000 (35 to 39.9 percent)
														              "B25095_008", # Less than $10,000 (40 to 49.9 percent)
														              "B25095_009", # Less than $10,000 (50 percent or more)
														              "B25095_015", # $10,000 to $19,999 (30 to 34.9 percent)
														              "B25095_016", # $10,000 to $19,999 (35 to 39.9 percent)
														              "B25095_017", # $10,000 to $19,999 (40 to 49.9 percent)
														              "B25095_018", # $10,000 to $19,999 (50 percent or more)
														              "B25095_024", # $20,000 to $34,999 (30 to 34.9 percent)
														              "B25095_025", # $20,000 to $34,999 (35 to 39.9 percent)
														              "B25095_026", # $20,000 to $34,999 (40 to 49.9 percent)
														              "B25095_027", # $20,000 to $34,999 (50 percent or more)
														              "B25095_033", # $35,000 to $49,999 (30 to 34.9 percent)
														              "B25095_034", # $35,000 to $49,999 (35 to 39.9 percent)
														              "B25095_035", # $35,000 to $49,999 (40 to 49.9 percent)
														              "B25095_036", # $35,000 to $49,999 (50 percent or more)
														              "B25095_042", # $50,000 to $74,999 (30 to 34.9 percent)
														              "B25095_043", # $50,000 to $74,999 (35 to 39.9 percent)
														              "B25095_044", # $50,000 to $74,999 (40 to 49.9 percent)
														              "B25095_045", # $50,000 to $74,999 (50 percent or more)
														              "B25095_051", # $75,000 to $99,999 (30 to 34.9 percent)
														              "B25095_052", # $75,000 to $99,999 (35 to 39.9 percent)
														              "B25095_053", # $75,000 to $99,999 (40 to 49.9 percent)
														              "B25095_054", # $75,000 to $99,999 (50 percent or more)
														              "B25095_060", # $100,000 to $149,999 (30 to 34.9 percent)
														              "B25095_061", # $100,000 to $149,999 (35 to 39.9 percent)
														              "B25095_062", # $100,000 to $149,999 (40 to 49.9 percent)
														              "B25095_063", # $100,000 to $149,999 (50 percent or more)
														              "B25095_069", # $150,000 or more (30 to 34.9 percent)
														              "B25095_070", # $150,000 or more (35 to 39.9 percent)
														              "B25095_071", # $150,000 or more (40 to 49.9 percent)
														              "B25095_072", # $150,000 or more (50 percent or more)
														              "B25095_001",
														              "B25095_001"), # Total
                     areas = area,
                     summary_level = "place") %>%
  rename(City = NAME) %>%
  mutate(CostBurdened = B25095_006+B25095_007+B25095_008+B25095_009+B25095_015+B25095_016+B25095_017+B25095_018+
	       B25095_024+B25095_025+B25095_026+B25095_027+B25095_033+B25095_034+B25095_035+B25095_036+B25095_042+B25095_043+B25095_044+B25095_045+B25095_051+B25095_052+B25095_053+B25095_054+B25095_060+B25095_061+B25095_062+B25095_063+B25095_069+B25095_070+B25095_071+B25095_072,
         CostBurdenedPer = CostBurdened/B25095_001) %>%
	dplyr::select(City ,CostBurdened,CostBurdenedPer) %>%
  left_join(Chelsea.MA, MA.CostBurdened.ch, by="City")

Chelsea.MA <- cbind(MA.CostBurdened.ch)

```
* The number of households, who spend 30 percent or more on housing costs in `r Chelsea.MA$City` is `r Chelsea.MA$CostBurdened` which is `r (round(Chelsea.MA$CostBurdenedPer,3))*100`% of total number of households, who spend money on housing costs.\
\


\

### Local/State Policies + Assistance on Rent Deferment
\

* The City of Providence has created a guide to serve as a centralized resource for housing-related information and resources related to COVID-19. The guide contains information for Residential Renters, Single-Family Homeowners, Commercial Renters, and Landlords. Moreover, the state has set up Housing Help RI as a centralized location for Rhode Islanders to apply for emergency housing assistance. In addition to the $6.5 million the state has made available for assistance to low-income Rhode Islanders, the City of Providence has allocated an additional 769,000 Dollar for assistance to Providence residents. (https://www.providenceri.gov/pema/covid-19-housing-resource-guide/)


```{r}
Providence <- Chelsea.MA
Providence
```

```{r}
library(data.table)
Providence <- as.data.frame(t(as.matrix(Providence)))
colnames(Providence) <- Providence[1,]
setDT(Providence, keep.rownames = TRUE)[]
names(Providence)[1] <- 'Indicators'
Providence
```

```{r}
library(data.table)
Newport <- as.data.frame(t(as.matrix(Newport)))
colnames(Newport) <- Newport[1,]
setDT(Newport, keep.rownames = TRUE)[]
names(Newport)[1] <- 'Indicators'
Newport
```
```{r}
NorthKingston <- as.data.frame(t(as.matrix(NorthKingston)))
colnames(NorthKingston) <- NorthKingston[1,]
setDT(NorthKingston, keep.rownames = TRUE)[]
names(NorthKingston)[1] <- 'Indicators'
NorthKingston
```

```{r}
two.cities <- merge(Providence, Newport, by="Indicators", all=TRUE)
two.cities
```

```{r}
Cities.MA <- merge(two.cities,NorthKingston, by="Indicators", all=TRUE)
Cities.MA
```

```{r}
Cities.MA$Indicators <- as.character(Cities.MA$Indicators)
Cities.MA$Indicators[Cities.MA$Indicators == 'TotalPop'] <- 'TotalPopulation'
Cities.MA$Indicators[Cities.MA$Indicators == 'AgeU18'] <- 'AgeUnder18'
Cities.MA$Indicators[Cities.MA$Indicators == 'AgeU18Per'] <- 'AgeUnder18_Proportion'
Cities.MA$Indicators[Cities.MA$Indicators == 'Age1834Per'] <- 'Age1834Proportion'
Cities.MA$Indicators[Cities.MA$Indicators == 'Age3564Per'] <- 'Age3564Proportion'
Cities.MA$Indicators[Cities.MA$Indicators == 'AgeO65'] <- 'AgeOlderThan65'
Cities.MA$Indicators[Cities.MA$Indicators == 'AgeO65Per'] <- 'AgeOlderThan65_Proportion'
Cities.MA$Indicators[Cities.MA$Indicators == 'MedHouseIncome'] <- 'MedianHouseIncome'
Cities.MA$Indicators[Cities.MA$Indicators == 'White'] <- 'WhitePopulation'
Cities.MA$Indicators[Cities.MA$Indicators == 'WhitePer'] <- 'WhitePopulation_Proportion'
Cities.MA$Indicators[Cities.MA$Indicators == 'Black'] <- 'BlackPopulation'
Cities.MA$Indicators[Cities.MA$Indicators == 'BlackPer'] <- 'BlackPopulation_Proportion'
Cities.MA$Indicators[Cities.MA$Indicators == 'Asian'] <- 'AsianPopulation'
Cities.MA$Indicators[Cities.MA$Indicators == 'AsianPer'] <- 'AsianProportion'
Cities.MA$Indicators[Cities.MA$Indicators == 'Hispanic'] <- 'HispanicPopulation'
Cities.MA$Indicators[Cities.MA$Indicators == 'HispanicPer'] <- 'HispanicPopulation_Proportion'
Cities.MA$Indicators[Cities.MA$Indicators == 'TwoOrMore'] <- 'PopulationWithTwoOrMoreRaces'
Cities.MA$Indicators[Cities.MA$Indicators == 'TwoOrMorePer'] <- 'PopulationWithTwoOrMoreRaces_Proportion'
Cities.MA$Indicators[Cities.MA$Indicators == 'FemalePer'] <- 'Female_Proportion'
Cities.MA$Indicators[Cities.MA$Indicators == 'MalePer'] <- 'Male_Proportion'
Cities.MA$Indicators[Cities.MA$Indicators == 'SexRatio'] <- 'NumberOfFemaleToMale'
Cities.MA$Indicators[Cities.MA$Indicators == 'AveHouseH'] <- 'AverageHouseholdSize'
Cities.MA$Indicators[Cities.MA$Indicators == 'LessThanHS'] <- 'LessThanHighSchoolDegree'
Cities.MA$Indicators[Cities.MA$Indicators == 'LessThanHSPer'] <- 'LessThanHighSchoolDegree_Proportion'
Cities.MA$Indicators[Cities.MA$Indicators == 'HSGrad'] <- 'HighSchoolDegree'
Cities.MA$Indicators[Cities.MA$Indicators == 'HSGradPer'] <- 'HighSchoolDegree_Proportion'
Cities.MA$Indicators[Cities.MA$Indicators == 'SomeColl'] <- 'SomeCollegeDegree'
Cities.MA$Indicators[Cities.MA$Indicators == 'SomeCollPer'] <- 'SomeCollegeDegree_Proportion'
Cities.MA$Indicators[Cities.MA$Indicators == 'Bach'] <- 'BachelorDegree'
Cities.MA$Indicators[Cities.MA$Indicators == 'BachPer'] <- 'BachelorDegree_Proportion'
Cities.MA$Indicators[Cities.MA$Indicators == 'Master'] <- 'MaterDegree'
Cities.MA$Indicators[Cities.MA$Indicators == 'MasterPer'] <- 'MaterDegree_Proportion'
Cities.MA$Indicators[Cities.MA$Indicators == 'Prof'] <- 'ProfessionalSchoolDegree'
Cities.MA$Indicators[Cities.MA$Indicators == 'ProfPer'] <- 'ProfessionalSchoolDegree_Proportion'
Cities.MA$Indicators[Cities.MA$Indicators == 'Doc'] <- 'DoctorateDegree'
Cities.MA$Indicators[Cities.MA$Indicators == 'DocPer'] <- 'DoctorateDegree_Proportion'
Cities.MA$Indicators[Cities.MA$Indicators == 'FemHeadPer'] <- 'SingleFemaleHeaded_Proportion'
Cities.MA$Indicators[Cities.MA$Indicators == 'MaleHeadPer'] <- 'SingleMaleHeaded_Proportion'
Cities.MA$Indicators[Cities.MA$Indicators == 'FamPovPer'] <- 'FamiliesBelowPovertyLevel_Proportion'
Cities.MA$Indicators[Cities.MA$Indicators == 'PopPovPer'] <- 'PopulationBelowPovertyLevel_Proportion'
Cities.MA$Indicators[Cities.MA$Indicators == 'TotalPopPov'] <- 'TotalPopulationBelowPovertyLevel'
Cities.MA$Indicators[Cities.MA$Indicators == 'OldAlonePer'] <- 'ElderlyPeopleLivingAlone_Proportion'
Cities.MA$Indicators[Cities.MA$Indicators == 'SchoolsDistrict'] <- 'NumberOfSchoolsDistrict'
Cities.MA$Indicators[Cities.MA$Indicators == 'PubSchools'] <- 'NumberOfPublicSchools'
Cities.MA$Indicators[Cities.MA$Indicators == 'PvSchools'] <- 'NumberOfPrivateSchools'
Cities.MA$Indicators[Cities.MA$Indicators == 'SchoolsDistStudents'] <- 'NumberOfStudentsInSchoolsDistrict'
Cities.MA$Indicators[Cities.MA$Indicators == 'NumUni'] <- 'NumberOfUniversity'
Cities.MA$Indicators[Cities.MA$Indicators == 'EnrolPK12'] <- 'EnrolledInPK-12'
Cities.MA$Indicators[Cities.MA$Indicators == 'NumJobs'] <- 'NumberOfJobs'
Cities.MA$Indicators[Cities.MA$Indicators == 'LabForce'] <- 'LaborForce'
Cities.MA$Indicators[Cities.MA$Indicators == 'LabForcePer'] <- 'LaborForce_Proportion'
Cities.MA$Indicators[Cities.MA$Indicators == 'CivLabForce'] <- 'CivilianLaborforce'
Cities.MA$Indicators[Cities.MA$Indicators == 'CivLabForcePer'] <- 'CivilianLaborforce_Proportion'
Cities.MA$Indicators[Cities.MA$Indicators == 'MilitaryPer'] <- 'Military_Proportion'
Cities.MA$Indicators[Cities.MA$Indicators == 'EmpRate'] <- 'EmploymentRate'
Cities.MA$Indicators[Cities.MA$Indicators == 'UnempRate'] <- 'UnemploymentRate'
Cities.MA$Indicators[Cities.MA$Indicators == 'EmployedCLF'] <- 'EmployedCivilians'
Cities.MA$Indicators[Cities.MA$Indicators == 'UnemployedCLF'] <- 'UnemployedCivilians'
Cities.MA$Indicators[Cities.MA$Indicators == 'FemaleNCAIS'] <- 'FemaleByIndustry'
Cities.MA$Indicators[Cities.MA$Indicators == 'FemaleNCAISPer'] <- 'FemaleByIndustry_Proportion'
Cities.MA$Indicators[Cities.MA$Indicators == 'MaleNCAIS'] <- 'MaleByIndustry'
Cities.MA$Indicators[Cities.MA$Indicators == 'MaleNCAISPer'] <- 'MaleByIndustry_Proportion'
Cities.MA$Indicators[Cities.MA$Indicators == 'Agriculture'] <- 'Agriculture&Forestry&Fishing&Hunting&Mining'
Cities.MA$Indicators[Cities.MA$Indicators == 'Transportation'] <- 'Transportation&Warehousing&Utilities'
Cities.MA$Indicators[Cities.MA$Indicators == 'Finance_Ins'] <- 'Finance&Insurance&RrealEstate&Rental&Leasing'
Cities.MA$Indicators[Cities.MA$Indicators == 'Professional'] <- 'Professional&Scientific&Mmanagement&Administrative&WasteManagmentServices'
Cities.MA$Indicators[Cities.MA$Indicators == 'Educ_Health_SocSvcs'] <- 'EducationalServices&HealthCare&SocialAssistance'
Cities.MA$Indicators[Cities.MA$Indicators == 'RecreationEtc'] <- 'Arts&Entertainment&Recreation&Accommodation&FoodServices'
Cities.MA$Indicators[Cities.MA$Indicators == 'OtherIndustries'] <- 'OtherServicesExceptPublicAdministration'
Cities.MA$Indicators[Cities.MA$Indicators == 'PublicAdmin'] <- 'PublicAdministration'
Cities.MA$Indicators[Cities.MA$Indicators == 'MedEarning'] <- 'MedianEarning'
Cities.MA$Indicators[Cities.MA$Indicators == 'AgeEmpU29'] <- 'EmployeesAgeUnder29'
Cities.MA$Indicators[Cities.MA$Indicators == 'AgeEmpU29Per_BG'] <- 'EmployeesAgeYoungerThan29_Proportion'
Cities.MA$Indicators[Cities.MA$Indicators == 'AgeEmp3054'] <- 'EmployeesAge3054'
Cities.MA$Indicators[Cities.MA$Indicators == 'AgeEmp3054Per_BG'] <- 'EmployeesAge3054_Proportion'
Cities.MA$Indicators[Cities.MA$Indicators == 'AgeEmpO55'] <- 'EmployeesAgeOlderThan55'
Cities.MA$Indicators[Cities.MA$Indicators == 'AgeEmpO55Per_BG'] <- 'EmployeesAgeOlderThan55_Proportion'
Cities.MA$Indicators[Cities.MA$Indicators == 'WhiteEmp'] <- 'WhiteEmployees'
Cities.MA$Indicators[Cities.MA$Indicators == 'WhiteEmpPer'] <- 'WhiteEmployees_Proportion'
Cities.MA$Indicators[Cities.MA$Indicators == 'BlackEmp'] <- 'BlackEmployees'
Cities.MA$Indicators[Cities.MA$Indicators == 'BlackEmpPer'] <- 'BlackEmployees_Proportion'
Cities.MA$Indicators[Cities.MA$Indicators == 'AsianEmp'] <- 'AsianEmployees'
Cities.MA$Indicators[Cities.MA$Indicators == 'AsianEmpPer'] <- 'AsianEmployees_Proportion'
Cities.MA$Indicators[Cities.MA$Indicators == 'HispanicEmp'] <- 'HispanicEmployees'
Cities.MA$Indicators[Cities.MA$Indicators == 'HispanicEmpPer'] <- 'HispanicEmployees_Propotion'
Cities.MA$Indicators[Cities.MA$Indicators == 'TwoOrMoreEmp'] <- 'EmployeesWithTwoOrMoreRaces'
Cities.MA$Indicators[Cities.MA$Indicators == 'TwoOrMoreEmpPer'] <- 'EmployeesWithTwoOrMoreRaces_Proportion'
Cities.MA$Indicators[Cities.MA$Indicators == 'LessThanHSLabor'] <- 'LaborsWithLessThanHighSchoolDegree'
Cities.MA$Indicators[Cities.MA$Indicators == 'LessThanHSLaborPer'] <- 'LaborsWithLessThanHighSchoolDegree_Proportion'
Cities.MA$Indicators[Cities.MA$Indicators == 'HSGradLabor'] <- 'LaborsWithHighSchoolDegree'
Cities.MA$Indicators[Cities.MA$Indicators == 'HSGradLaborPer'] <- 'LaborsWithHighSchoolDegree_Proportion'
Cities.MA$Indicators[Cities.MA$Indicators == 'SomeCollLabor'] <- 'LaborsWithCollegeDegree'
Cities.MA$Indicators[Cities.MA$Indicators == 'SomeCollLaborPer'] <- 'LaborsWithCollegeDegree_Proportion'
Cities.MA$Indicators[Cities.MA$Indicators == 'BachOrHigherLabor'] <- 'LaborsWithBachelorOrHigherDegree'
Cities.MA$Indicators[Cities.MA$Indicators == 'BachOrHigherLaborPer'] <- 'LaborsWithBachelorOrHigherDegree_Proportion'
Cities.MA$Indicators[Cities.MA$Indicators == 'EmpFemale'] <- 'FemaleEmployess'
Cities.MA$Indicators[Cities.MA$Indicators == 'EmpFemalePer'] <- 'FemaleEmployess_Proportion'
Cities.MA$Indicators[Cities.MA$Indicators == 'EmpMale'] <- 'MaleEmployess'
Cities.MA$Indicators[Cities.MA$Indicators == 'EmpMalePer'] <- 'MaleEmployess_Proportion'
Cities.MA$Indicators[Cities.MA$Indicators == 'TotalEmp'] <- 'TotalEmployees'
Cities.MA$Indicators[Cities.MA$Indicators == 'NursingHomes'] <- 'NumberOfNursingHomes'
Cities.MA$Indicators[Cities.MA$Indicators == 'NumGrocery'] <- 'NumberOfGrocery'
Cities.MA$Indicators[Cities.MA$Indicators == 'SnapRec'] <- 'SNAPRecipients '
Cities.MA$Indicators[Cities.MA$Indicators == 'HomeOwnPer'] <- 'ProportionOfHomeOwners'
Cities.MA$Indicators[Cities.MA$Indicators == 'RentersPer'] <- 'NumberOfRenters_Proportion'
Cities.MA$Indicators[Cities.MA$Indicators == 'RentersToOwners'] <- 'ProportionOfRentersToOwners'
Cities.MA$Indicators[Cities.MA$Indicators == 'MedianHValue'] <- 'MedianHomeValue'
Cities.MA$Indicators[Cities.MA$Indicators == 'AvgHValue'] <- 'AverageHomeValues'
Cities.MA$Indicators[Cities.MA$Indicators == 'PropertyTaxRate.Com'] <- 'PropertyTaxRateCommercial'
Cities.MA$Indicators[Cities.MA$Indicators == 'PropertyTaxRate.RES'] <- 'PropertyTaxRateResidential'
Cities.MA$Indicators[Cities.MA$Indicators == 'MedYrMovedIn'] <- 'MedianYearMovedIn'
Cities.MA$Indicators[Cities.MA$Indicators == 'MedYrOwnMovedIn'] <- 'MedianYearOwnerMovedIn'
Cities.MA$Indicators[Cities.MA$Indicators == 'MedYrRentMovedIn'] <- 'MedianYearRenterMovedIn'
Cities.MA$Indicators[Cities.MA$Indicators == 'CostBurdenedPer'] <- 'CostBurdened_Proportion'
Cities.MA$Indicators[Cities.MA$Indicators == 'Homelessness'] <- 'NumberOfHomelessPeople'
Cities.MA
write.csv(Cities.MA,"~/Desktop/Cities.RI.csv", row.names = FALSE)
```










